---
title: "Analysis LST SK Test diff"
author: "Anna Talucci"
date: "2024-10-10"
output: html_document
---


# Overview
Analysis evaluating differences in land surface temperature between reburn classes across the boreal from 2013-2022. 

# Packages
```{r}
library(tidyverse)
library(emmeans) 
library(tweedie)
library(statmod)
library(car)

library(MASS)
library(cplm)
library(aod)
```


# Data

```{r}
alldata = read_csv('../outputs/cleanedSamplePoints/LstBoreal2013-2022.csv') 
```

```{r}
data18 = read_csv('../outputs/cleanedSamplePoints/2018LSTSampleWesternCanadaAlaska.cvs')
```
```{r}
data18
alldata
```

# check factors

```{r}
data18 = data18 %>% mutate(TimeSinceFire= ordered(TimeSinceFire, levels = c("0", "1", "2-5", "6-10", "11-20", "21-31", "Other")),
         burned = ordered(burned, levels = c("Unburned", "One", "Two", "Three", "other")))
```

# Subset
```{r}
unique(data18$TimeSinceFire)
unique(alldata$TimeSinceFire)
```

```{r}
( lstLong = data18 %>% 
    rename(lat = latitude, lon = longitude) %>% 
  filter(burned %in% c("One", "Three", "Two", "Unburned")) %>%
  pivot_longer(
  cols = c(max_2018, mean_2018, median_2018, min_2018),
  names_to = "statYr",
  values_to = "tempK") %>% 
  separate(statYr, c("lstStat", "lstYr")) %>%
  mutate(lstYr = as.numeric(lstYr)) %>%
    mutate(bTSF = paste0(burned,"_",TimeSinceFire))
)
  
```

```{r}
( dataLong = alldata %>% 
    rename(lat = latitude, lon = longitude) %>% 
  filter(burned %in% c("One", "Three", "Two", "Unburned")) %>%
  pivot_longer(
  cols = starts_with("mean"),
  names_to = "statYr",
  values_to = "tempK") %>% 
  separate(statYr, c("lstStat", "lstYr")) %>%
  mutate(lstYr = as.numeric(lstYr))
    
)
  
```

```{r}
( lstStats = lstLong %>% pivot_wider(names_from = lstStat, values_from = tempK) %>% filter(TimeSinceFire %in% c("0", "1", "2-5")))
```
```{r}
lstStats$TimeSinceFire = factor(lstStats$TimeSinceFire, levels = c( "0", "1", "2-5"))
lstStats$burned = factor(lstStats$burned, levels = c( "Unburned", "One", "Two", "Three"))
```

```{r}
unique(lstStats$tsfClass)  
unique(lstStats$burned)  
```
Gaussian

```{r}
( recent = lstLong %>% filter(TimeSinceFire %in% c("0", "1", "2-5")) %>% filter(burned %in% c("Unburned", "One", "Two", "Three")) )
```

```{r}
( recentAllData = dataLong %>% filter(TimeSinceFire %in% c("0", "1", "2-5")) %>% filter(burned %in% c("Unburned", "One", "Two", "Three")) )
```

```{r}
( lateTimeSteps = dataLong %>% filter(TimeSinceFire %in% c("6-10", "10-19", "20-31")) %>% filter(burned %in% c("Unburned", "One", "Two", "Three")) )
```

```{r}
unique(lateTimeSteps$TimeSinceFire)
```
```{r}
lateTimeSteps$TimeSinceFire = factor(lateTimeSteps$TimeSinceFire, levels = c( "6-10", "10-19", "20-31", "other"))
lateTimeSteps$burned = factor(lateTimeSteps$burned, levels = c( "Unburned", "One", "Two", "Three"))
```

# Ecozones
```{r}
unique(recentAllData$ECO_NAME)
```
# Model Recent time since fire effects

```{r}
fit1 = glm(tempK ~ bTSF, data = recentAllData, family = gaussian)
```

```{r}
summary(fit1)
```

### Residuals: 

```{r eval=FALSE, include=FALSE}
recent$fit1.res = resid(fit1, type = "pearson")
recent$fit1.fit = fitted(fit1)

sum(residuals(fit1, type="pearson")^2)/fit1$df.res

qplot(fitted(fit1), resid(fit1)) + theme_bw()

```


```{r}
summary(fit1)
```

```{r}
emmfit1 = emmeans(fit1, specs = ~ bTSF)
emmfit1
```
```{r}
dfemmfit1 = as.data.frame(emmfit1)

dfemmfit1.1 = dfemmfit1 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTSF, levels = c("Unburned_0", 'One_1', 'One_2-5',  'Two_1', 'Two_2-5', 'Three_1','Three_2-5')))

dfemmfit1.1
```

#5495CFFF, #F5AF4DFF, #DB4743FF, #7C873EFF, #FEF4D5FF
```{r}
allDataFit1Plot = ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#7C873EFF','#DB4743FF','#F5AF4DFF',  '#DB4743FF','#F5AF4DFF', '#DB4743FF','#F5AF4DFF'), labels=c("unburned", "one year", "2-5 years")) +
  labs(x="", y="Land surface temperature (Kelvin)") +
  scale_x_discrete(labels= c("Unburned", "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

allDataFit1Plot
```

```{r}
ggsave("../figures/LSTrecentBoreal2013-2022.png", plot = allDataFit1Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```

### Annotate Plot
```{r}
allDataFit1Comp1 = ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#7C873EFF','#DB4743FF','#F5AF4DFF',  '#DB4743FF','#F5AF4DFF', '#DB4743FF','#F5AF4DFF'), labels=c("unburned", "one year", "2-5 years")) +
  labs(x="", y="Land surface temperature (Kelvin)") +
  scale_x_discrete(labels= c("Unburned", "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() +
  annotate("text", x = 2, y = 288.3, label = "*", color='#DB4743FF',  size = 10) +
  annotate("text", x = 4, y = 288.9, label = "*", color='#DB4743FF',  size = 10) +
  annotate("text", x = 6, y = 289.2, label = "*", color='#DB4743FF',  size = 10)
  #theme(legend.position="none")

allDataFit1Comp1
```

```{r}
ggsave("../figures/LSTrecentBoreal2013-2022CompareOneTimeBurned.png", plot = allDataFit1Comp1, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```


```{r}
(recentLSTPlot = ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#1b9e77','#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c("Unburned", "One year", "2-5 years")) +
  labs(x="Times burned", y="Land surface Temperature (Kelvin)") +
  scale_x_discrete(labels= c("Unburned", "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() +
   theme( legend.position="right", 
        legend.text=element_text(size=14), 
        legend.title=element_text(size=16),
        legend.key = element_blank()) +
    theme(axis.title.y = element_text(size = 16, hjust = 0.5, vjust = 1.1),
          axis.title.x = element_text(size = 16, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black")) 
)
```

```{r}
ggsave("../figures/LSTrecentBoreal2013-2022CompareOneTimeBurnedPPTSlide.png", plot = recentLSTPlot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```






## Testing difference in means

```{r}
emmfit1
```

```{r}
One_1 = c(1, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0, 0)
Three_1 = c(0, 0, 1, 0, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 1, 0, 0, 0)
Two_1 = c(0, 0, 0, 0, 1, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1, 0)
Unburned_0 = c(0, 0, 0, 0, 0, 0, 1) 
```

```{r}
compUnburn = contrast(emmfit1, method = list("Unburned_0 - One_1" = Unburned_0 - One_1,
                             "Unburned_0 - One_2_5" = Unburned_0 - One_2_5,
                             "Unburned_0 - Two_1" = Unburned_0 - Two_1,
                             "Unburned_0 - Two_2_5" = Unburned_0 - Two_2_5,
                             "Unburned_0 - Three_1" = Unburned_0 - Three_1,
                             "Unburned_0 - Three_2_5" = Unburned_0 - Three_2_5),
         adjust = "mvt") %>%
     confint()

compUnburn
```

```{r}
contrast(emmfit1, method = list("Unburned_0 - One_1" = Unburned_0 - One_1,
                             "Unburned_0 - One_2_5" = Unburned_0 - One_2_5,
                             "Unburned_0 - Two_1" = Unburned_0 - Two_1,
                             "Unburned_0 - Two_2_5" = Unburned_0 - Two_2_5,
                             "Unburned_0 - Three_1" = Unburned_0 - Three_1,
                             "Unburned_0 - Three_2_5" = Unburned_0 - Three_2_5),
         adjust = "mvt")
```

```{r}
contrast(emmfit1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))
```
```{r}
compTB = contrast(emmfit1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()

compTB
```

# Model All time steps since fire

```{r}
fit2 = glm(tempK ~ bTSF, data = lateTimeSteps, family = gaussian)
```

```{r}
summary(fit2)
```

### Residuals: 

```{r eval=FALSE, include=FALSE}
recent$fit2.res = resid(fit2, type = "pearson")
recent$fit2.fit = fitted(fit2)

sum(residuals(fit2, type="pearson")^2)/fit2$df.res

qplot(fitted(fit2), resid(fit2)) + theme_bw()

```


```{r}
summary(fit2)
```

```{r}
emmfit2 = emmeans(fit2, specs = ~ bTSF)
emmfit2
```

```{r}
unique(lateTimeSteps$bTSF)
```

```{r}
dfemmfit2 = as.data.frame(emmfit2)

dfemmfit2.1 = dfemmfit2 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTSF, levels = c("One_6-10", "One_10-19", "One_20-31", "Two_6-10", "Two_10-19", "Two_20-31", "Three_6-10", "Three_10-19", "Three_20-31")))

dfemmfit2.1
```

#ECB100FF, #A10000FF, #7E605EFF, #FF4500FF, #994823FF, #4B4C4EFF, #572530FF, #000000FF
```{r}
lateDatafit2Plot = ggplot(data = dfemmfit2.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  #scale_color_manual(name= "Time Since Fire", values=c('#7C873EFF','#DB4743FF', '#ECB100FF', '#A10000FF', '#7E605EFF', '#FF4500FF', '#994823FF', '#4B4C4EFF', '#572530FF', '#000000FF')) +
  labs(x="", y="Land surface temperature (Kelvin)") +
  #scale_x_discrete(labels= c("Unburned", "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

lateDatafit2Plot
```











# Model 2018 only DELETE
```{r}
model1 = glm(tempK ~ bTSF, data = recent, family = gaussian)
```

```{r}
summary(model1)
```

### Residuals: 

```{r}
recent$model1.res = resid(model1, type = "pearson")
recent$model1.fit = fitted(model1)

sum(residuals(model1, type="pearson")^2)/model1$df.res

qplot(fitted(model1), resid(model1)) + theme_bw()

```

```{r}

```
```{r}
summary(model1)
```

```{r}
emm1 = emmeans(model1, specs = ~ bTSF)
emm1
```
```{r}
dfemm1 = as.data.frame(emm1)

dfemm1.1 = dfemm1 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTSF, levels = c("Unburned_0", 'One_1', 'One_2-5',  'Two_1', 'Two_2-5', 'Three_1','Three_2-5')))

dfemm1.1
```
```{r}
allDataPlot = ggplot(data = dfemm1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#1b9e77','#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c("unburned", "one year", "2-5 years")) +
  labs(x="", y="Albedo") +
  scale_x_discrete(labels= c("Unburned", "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

allDataPlot
```

```{r}
pairwise.t.test(recent$tempK, recent$bTSF, p.adjust.method="bonferroni")
```

```{r}
emm1
```
```{r}
pairs(emm1)
```
```{r}
One_1 = c(1, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0, 0)
Three_1 = c(0, 0, 1, 0, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 1, 0, 0, 0)
Two_1 = c(0, 0, 0, 0, 1, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1, 0)
Unburned_0 = c(0, 0, 0, 0, 0, 0, 1) 
```

```{r}
compUnburn = contrast(emm1, method = list("Unburned_0 - One_1" = Unburned_0 - One_1,
                             "Unburned_0 - One_2_5" = Unburned_0 - One_2_5,
                             "Unburned_0 - Two_1" = Unburned_0 - Two_1,
                             "Unburned_0 - Two_2_5" = Unburned_0 - Two_2_5,
                             "Unburned_0 - Three_1" = Unburned_0 - Three_1,
                             "Unburned_0 - Three_2_5" = Unburned_0 - Three_2_5),
         adjust = "mvt") %>%
     confint()

compUnburn
```

```{r}
contrast(emm1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))
```
```{r}
compTB = contrast(emm1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()

compTB
```
