---
title: "Analysis LST SK Test diff"
author: "Anna Talucci"
date: "2024-10-10"
output: html_document
---

```{r clear environment, include=FALSE}
rm(list=ls())
```

# Overview

Analysis for both absolute and relative land surface temperature differences between once, twice and three times burned. This script includes some data organization. 

# Packages

```{r}
library(tidyverse)
library(sf)
library(emmeans) 
library(statmod)
library(car)

library(MASS)
library(cplm)
library(aod)
library(cowplot)
```

# Data

## Read in

```{r}
aklst = read_csv('../outputs/cleanedSamplePoints/LstBoreal2013-2022.csv')
```
```{r}
perimeters = st_read("../outputs/boundaries/firePerimeters/akFirePerimeters_1984_2020.shp", "akFirePerimeters_1984_2020")
```

## View

```{r}
aklst
```



## Split data into burned unburned

```{r}
( burned_lst = aklst %>% filter(countYr!=0) )
( unburned_lst = aklst %>% filter(countYr==0) )
```

# Clean and Organize data for model

```{r}
burned_lst
```

```{r}
sort(unique(burned_lst$maxYr))
```

```{r}
burned_lst %>% mutate(tsf2022 = 2022-maxYr) %>% group_by(as.factor(tsf2022)) %>% summarise(n=n())
```

## Clean unburned

```{r}
unburned_lst
```
```{r}
( unburned_data = unburned_lst %>% 
  pivot_longer(cols = starts_with("mean"),
    names_to = "lstSample",
    values_to = "lstMean",
    values_drop_na = TRUE) %>%
  dplyr::select(ID:longitude, burned, maxYr, bTSF, lstSample, lstMean, TimeSinceFire) %>%
  separate(lstSample, into=c("lstFun", "year"), sep="\\_", convert=TRUE, extra = "merge" ) %>%
  mutate(tsf = 0, 
         refLst = lstMean,
         pre_post_Lst = refLst - lstMean,
         post_pre_Lst = lstMean - refLst, 
         tsfStd = "0",
         tsfAlt = "0",
         bTsfStd = paste0(burned, "_", tsfStd),
         bTsfAlt = paste0(burned, "_", tsfAlt),
         sampleYr = year) %>%
  dplyr::select(-lstFun, -year)
    
)   
```
  
## Clean Burned

### ref year for pre-post calculation

```{r}
calculate_refLst <- function(year) {
  unburned_lst %>% 
    group_by(ECO_NAME) %>% 
    summarise(refLst = round(mean(get(paste0("mean_", year))), 1))
}

# Create a list of reference datasets for multiple years
years <- 2014:2022
ref_data <- setNames(lapply(years, calculate_refLst), paste0("ref", years))
ref_data
```

### Function for tsf bins
```{r}
calculate_standardPre <- function(year) {
  ref_year <- ref_data[[paste0("ref", year)]]
  burned_lst %>%
    filter(maxYr < year) %>%
    dplyr::select(ID:maxYr, !!sym(paste0("mean_", year)), burned) %>%
    mutate(refLst = case_when(
      ECO_NAME %in% ref_year$ECO_NAME ~ ref_year$refLst[match(ECO_NAME, ref_year$ECO_NAME)]
    )) %>%
    mutate(
      tsf = year - maxYr,
      pre_post_Lst = refLst - get(paste0("mean_", year)),
      post_pre_Lst =  get(paste0("mean_", year)) - refLst
    ) %>%
    mutate(
      TimeSinceFire = case_when(
        tsf == 1 ~ "1",
        tsf %in% 2:5 ~ "2-5",
        tsf %in% 6:10 ~ "6-10",
        tsf %in% 11:20 ~ "11-20",
        tsf %in% 21:30 ~ "21-30",
        tsf %in% 31:36 ~ "31-36",
        TRUE ~ "Other"
      ),
      tsfStd = case_when(
        tsf %in% 1:5 ~ "1-5",
        tsf %in% 6:10 ~ "6-10",
        tsf %in% 11:20 ~ "11-20",
        tsf %in% 21:36 ~ "21+",
        TRUE ~ "Other"
      ),
      tsfAlt = case_when(
        tsf == 1 ~ "1",
        tsf %in% 2:5 ~ "2-5",
        tsf %in% 6:15 ~ "6-15",
        tsf %in% 16:36 ~ "16+",
        TRUE ~ "Other"
      )
    ) %>%
    mutate(across(
      c(TimeSinceFire),
      ~ ordered(., levels = c("1", "2-5", "6-10", "11-20", "21-30", "31-36"))
    )) %>%
     mutate(across(
      c(tsfStd),
      ~ ordered(., levels = c("1-5", "6-10", "11-20", "21+"))
    )) %>%
     mutate(across(
      c(tsfAlt),
      ~ ordered(., levels = c("1", "2-5", "6-15", "16+"))
    )) %>%
    mutate(bTSF = paste0(burned, "_", TimeSinceFire),
           bTsfStd = paste0(burned, "_", tsfStd),
           bTsfAlt = paste0(burned, "_", tsfAlt),
           sampleYr = year) %>%
    rename(lstMean = !!sym(paste0("mean_", year)))
}

# Generate standard pre-year data for multiple years
standardPre_data <- setNames(lapply(years, calculate_standardPre), paste0("standardPre", years))

```

```{r}
standardPre_data
```

## Creat final DF
```{r}
final_df <- bind_rows(
  standardPre_data$standardPre2022, standardPre_data$standardPre2021, standardPre_data$standardPre2020, standardPre_data$standardPre2019, 
  standardPre_data$standardPre2018, standardPre_data$standardPre2017, standardPre_data$standardPre2016, standardPre_data$standardPre2015, 
  standardPre_data$standardPre2014, unburned_data
) %>%
    mutate(across(
      c(TimeSinceFire),
      ~ ordered(., levels = c("0", "1", "2-5", "6-10", "11-20", "21-30", "31-36"))
    )) %>%
     mutate(across(
      c(tsfStd),
      ~ ordered(., levels = c("0", "1-5", "6-10", "11-20", "21+"))
    )) %>%
     mutate(across(
      c(tsfAlt),
      ~ ordered(., levels = c("0", "1", "2-5", "6-15", "16+"))
    )) %>%
     mutate(across(
      c(bTsfAlt),
      ~ ordered(., levels = c("Unburned_0", "One_1", "One_2-5", "One_6-15", "One_16+",  "Two_1", "Two_2-5", "Two_6-15", "Two_16+",  "Three_1", "Three_2-5", "Three_6-15", "Three_16+"))
    )) 

```

```{r}
final_df
```

## split recent old

```{r}
( recent_burned_2014_2022 = final_df %>% 
    filter(tsf <=5) %>% 
    filter(burned !="Unburned") %>%
     mutate(across(
      c(bTSF),
      ~ ordered(., levels = c("Unburned_0", "One_1", "One_2-5", "Two_1", "Two_2-5", "Three_1", "Three_2-5"))
    )) )
```

```{r}
( old_burned_2014_2022 = final_df %>% filter(tsf >=5) %>% filter(burned !="Unburned"))
```

```{r}
( all_burned_2014_2022 = final_df %>% 
    filter(burned !="Unburned") %>%
     mutate(across(
      c(bTsfAlt),
      ~ ordered(., levels = c("Unburned_0", "One_1", "One_2-5", "One_6-15", "One_16+",  "Two_1", "Two_2-5", "Two_6-15", "Two_16+",  "Three_1", "Three_2-5", "Three_6-15", "Three_16+"))
    )) )
```
# Analysis Burned & Unburned

# Absolute Only

### Model Fit 2: Absolute All time steps since fire

```{r}
fit0 = glm(lstMean ~ bTsfAlt, data =  final_df, family = gaussian)
```


#### Residuals: 

```{r}
summary(fit0)

final_df$fit0.res = resid(fit0, type = "pearson")
final_df$fit0.fit = fitted(fit0)

sum(residuals(fit0, type="pearson")^2)/fit0$df.res

qplot(fitted(fit0), resid(fit0)) + theme_bw()

```

#### Calculate marginal means for difference test

```{r}
emmfit0 = emmeans(fit0, specs = ~ bTsfAlt)
emmfit0
```

```{r}
dfemmfit0 = as.data.frame(emmfit0)

dfemmfit0.1 = dfemmfit0 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTsfAlt, levels = c("Unburned_0", "One_1", "One_2-5", "One_6-15", "One_16+", "Two_1", "Two_2-5", "Two_6-15", "Two_16+", "Three_1", "Three_2-5", "Three_6-15", "Three_16+"))
)
dfemmfit0.1
```

```{r}
Unburned_0 = c(1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)

One_1 =  c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_6_15 = c(0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_16_36 =c(0, 0, 0, 0, 5, 0, 0, 0, 0, 0, 0, 0, 0)

Two_1  = c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
Two_6_15 =  c(0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0) 
Two_16_36 = c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)

Three_1=  c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
Three_2_5 =c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
Three_6_15= c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
Three_16_36= c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)



```

```{r}
contrast(emmfit0, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36))

contrast(emmfit0, method = list("Unburned_0 - Two_1" = Unburned_0 - Two_1,
                             "Unburned_0 - Three_1" = Unburned_0 - Three_1,
                             "Unburned_0 - Two_2_5" = Unburned_0 - Two_2_5,
                             "Unburned_0 - Three_2_5" = Unburned_0 - Three_2_5,
                             "Unburned_0 - Two_6_15" = Unburned_0 - Two_6_15,
                             "Unburned_0 - Three_6_15" = Unburned_0 - Three_6_15,
                             "Unburned_0 - Two_16_36" = Unburned_0 - Two_16_36,
                             "Unburned_0 - Three_16_36" = Unburned_0 - Three_16_36))

contrast(emmfit0, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36),
         adjust = "mvt") %>%
     confint()


```



#### Plots



```{r}
( modelfit0_Plot = ggplot(data = dfemmfit0.1 ) + 
  geom_point(aes(x=bTsfAlt, y=mean, color = bTsfAlt)) +
  geom_errorbar(aes(x=bTsfAlt, ymin = LCL, ymax = UCL, color = bTsfAlt)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#082a54', '#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("Unburned", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years")) +
  labs(x="", y="Land surface temperature\n(Kelvin)")  +
  scale_x_discrete(labels= c("Unburned", "One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

)


```

```{r}
ggsave("../figures/2025-03-18_LSTAbsoluteWithUnburned.png", plot = modelfit0_Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```

```{r}
( modelfit0_PlotAbsolute = ggplot(data = dfemmfit0.1 ) + 
  geom_point(aes(x=bTsfAlt, y=mean, color = bTsfAlt)) +
  geom_errorbar(aes(x=bTsfAlt, ymin = LCL, ymax = UCL, color = bTsfAlt)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years")) +
  labs(x="", y="Land surface temperature\n(Kelvin)")  +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() +
  theme(legend.position="none")

)


```



# Analysis Burned Only

## Absolute

### Model Recent time since fire effects

```{r}
fit1 = glm(lstMean ~ bTSF, data = recent_burned_2014_2022, family = gaussian)
```


#### Residuals: 

```{r}
summary(fit1)

recent_burned_2014_2022$fit1.res = resid(fit1, type = "pearson")
recent_burned_2014_2022$fit1.fit = fitted(fit1)

sum(residuals(fit1, type="pearson")^2)/fit1$df.res

qplot(fitted(fit1), resid(fit1)) + theme_bw()

```

#### Calculate Marginal means for difference test

```{r}
emmfit1 = emmeans(fit1, specs = ~ bTSF)
emmfit1
```

```{r}
dfemmfit1 = as.data.frame(emmfit1)

dfemmfit1.1 = dfemmfit1 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTSF, levels = c("Unburned_0", 'One_1', 'One_2-5',  'Two_1', 'Two_2-5', 'Three_1','Three_2-5')))

dfemmfit1.1
```

```{r}
One_1 = c(1, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0)

Two_1 = c(0, 0, 1, 0, 0, 0)
Two_2_5 = c(0, 0, 0, 1, 0, 0)

Three_1 = c(0, 0, 0, 0, 1, 0)
Three_2_5 = c(0, 0, 0, 0, 0, 0)
```


```{r}
contrast(emmfit1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))

contrast(emmfit1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()

```

#### Plots 
```{r}
ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip()
```

```{r}
allDataFit1Plot = ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#DB4743FF','#F5AF4DFF',  '#DB4743FF','#F5AF4DFF', '#DB4743FF','#F5AF4DFF'), labels=c( "one year", "2-5 years", "one year", "2-5 years", "one year", "2-5 years")) +
  labs(x="", y="Land surface temperature (Kelvin)") +
  scale_x_discrete(labels= c( "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

allDataFit1Plot
```

```{r}
ggsave("../figures/2025-03-18_RecentAbsoluteLST.png", plot = allDataFit1Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```


```{r}
ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#DB4743FF','#F5AF4DFF',  '#DB4743FF','#F5AF4DFF', '#DB4743FF','#F5AF4DFF'), labels=c("one year", "2-5 years", "one year", "2-5 years", "one year", "2-5 years")) +
  labs(x="", y="Land surface temperature (Kelvin)") +
  scale_x_discrete(labels= c("One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() +
  annotate("text", x = 2, y = 288.3, label = "*", color='#DB4743FF',  size = 10) +
  annotate("text", x = 4, y = 288.9, label = "*", color='#DB4743FF',  size = 10) +
  annotate("text", x = 6, y = 289.2, label = "*", color='#DB4743FF',  size = 10)
  #theme(legend.position="none")

```


```{r}
(recentLSTPlot = ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c( "One year", "2-5 years", "One year", "2-5 years", "One year", "2-5 years")) +
  labs(x="Times burned", y="Land surface Temperature (Kelvin)") +
  scale_x_discrete(labels= c( "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() +
   theme( legend.position="right", 
        legend.text=element_text(size=14), 
        legend.title=element_text(size=16),
        legend.key = element_blank()) +
    theme(axis.title.y = element_text(size = 16, hjust = 0.5, vjust = 1.1),
          axis.title.x = element_text(size = 16, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black")) 
)
```

```{r}
ggsave("../figures/2025-03-18_AbsoluteRecentLstPptSlide.png", plot = recentLSTPlot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```

### Model Fit 2: Absolute All time steps since fire

```{r}
fit2 = glm(lstMean ~ bTsfAlt, data =  all_burned_2014_2022, family = gaussian)
```


#### Residuals: 

```{r}
summary(fit2)

 all_burned_2014_2022$fit2.res = resid(fit2, type = "pearson")
 all_burned_2014_2022$fit2.fit = fitted(fit2)

sum(residuals(fit2, type="pearson")^2)/fit2$df.res

qplot(fitted(fit2), resid(fit2)) + theme_bw()

```

#### Calculate marginal means for difference test

```{r}
emmfit2 = emmeans(fit2, specs = ~ bTsfAlt)
emmfit2

dfemmfit2 = as.data.frame(emmfit2)

dfemmfit2.1 = dfemmfit2 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTsfAlt, levels = c("One_1", "One_2-5", "One_6-15", "One_16+", "Two_1", "Two_2-5", "Two_6-15", "Two_16+", "Three_1", "Three_2-5", "Three_6-15", "Three_16+"))
)
dfemmfit2.1
```

```{r}
One_1 =  c(1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_6_15 =c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_16_36 =c(0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0)

Two_1 =  c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
Two_6_15 =c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
Two_16_36 = c(0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)

Three_1 = c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
Three_6_15 =  c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0) 
Three_16_36 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
```

```{r}
contrast(emmfit2, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36))

contrast(emmfit2, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36),
         adjust = "mvt") %>%
     confint()


```



#### Plots



```{r}
( modelfit2_Plot = ggplot(data = dfemmfit2.1 ) + 
  geom_point(aes(x=bTsfAlt, y=mean, color = bTsfAlt)) +
  geom_errorbar(aes(x=bTsfAlt, ymin = LCL, ymax = UCL, color = bTsfAlt)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years")) +
  labs(x="", y="Land surface temperature\n(Kelvin)")  +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

)


```

```{r}
ggsave("../figures/2025-03-18_LstAllYrsBurnedAbsolute.png", plot = modelfit2_Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```

```{r}
( modelfit2_PlotAbsolute = ggplot(data = dfemmfit2.1 ) + 
  geom_point(aes(x=bTsfAlt, y=mean, color = bTsfAlt)) +
  geom_errorbar(aes(x=bTsfAlt, ymin = LCL, ymax = UCL, color = bTsfAlt)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years")) +
  labs(x="", y="Land surface temperature\n(Kelvin)")  +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() +
  theme(legend.position="none")

)


```

## Relative (post-pre) Models
### Model Recent time since fire effects
 Match AGU and ABOVE Analaysis format
```{r}
model1 = glm(post_pre_Lst ~ bTSF, data = recent_burned_2014_2022, family = gaussian)
```

#### Residuals: 

```{r}
recent_burned_2014_2022$model1.res = resid(model1, type = "pearson")
recent_burned_2014_2022$model1.fit = fitted(model1)

sum(residuals(model1, type="pearson")^2)/model1$df.res

qplot(fitted(model1), resid(model1)) + theme_bw()

```

#### Calculate marginal means for difference

```{r}
emm_model1 = emmeans(model1, specs = ~ bTSF)
emm_model1

df_emm_model1 = as.data.frame(emm_model1)

df_emm_model1 = df_emm_model1 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTSF, levels = c( "One_1", "One_2-5",  "Two_1",  "Two_2-5",  "Three_1", "Three_2-5"))
                  )

df_emm_model1
```

```{r}
One_1 = c(1, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0)

Two_1 = c(0, 0, 1, 0, 0, 0)
Two_2_5 = c(0, 0, 0, 1, 0, 0)

Three_1 = c(0, 0, 0, 0, 1, 0)
Three_2_5 = c(0, 0, 0, 0, 0, 1)

contrast(emm_model1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))

contrast(emm_model1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()
```


#### Plots

```{r}
(recent_model1_plot1 = ggplot(data = df_emm_model1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c("One year", "2-5 years")) +
  labs(x="Times burned", y="Land surface Temperature Difference (Kelvin)", subtitle = "Difference = post LST - pre LST") +
  scale_x_discrete(labels= c("One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() +
   theme( legend.position="right", 
        legend.text=element_text(size=14), 
        legend.title=element_text(size=16),
        legend.key = element_blank()) +
    theme(axis.title.y = element_text(size = 16, hjust = 0.5, vjust = 1.1),
          axis.title.x = element_text(size = 16, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black")) 
)
```

```{r}
ggsave("../figures/2025-03-18_LstRecentRelativePPTSlide.png", plot = recent_model1_plot1, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```


## Model 4 with Alternate Time Since Fire groups
Full time for space substitution.

```{r}
model4 = glm(post_pre_Lst ~ bTsfAlt, data =  all_burned_2014_2022, family = gaussian)
```

#### Residuals: 

```{r}
 all_burned_2014_2022$model4.res = resid(model4, type = "pearson")
 all_burned_2014_2022$model4.fit = fitted(model4)

sum(residuals(model4, type="pearson")^2)/model4$df.res

qplot(fitted(model4), resid(model4)) + theme_bw()

summary(model4)
```

#### Calculate marginal means for difference test

```{r}
emm_model4 = emmeans(model4, specs = ~ bTsfAlt)
emm_model4

df_emm_model4 = as.data.frame(emm_model4)

df_emm_model4 = df_emm_model4 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTsfStd = factor(bTsfAlt, levels = c("One_1", "One_2-5", "One_6-15", "One_16+", "Two_1", "Two_2-5", "Two_6-15", "Two_16+", "Three_1", "Three_2-5", "Three_6-15", "Three_16+"))
                  )

df_emm_model4
```

```{r}
One_1 =  c(1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_6_15 =c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_16_36 =c(0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0)

Two_1 =  c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
Two_6_15 =c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
Two_16_36 = c(0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)

Three_1 = c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
Three_6_15 =  c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0) 
Three_16_36 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
```

```{r}
contrast(emm_model4, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36))

compTB = contrast(emm_model4, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36),
         adjust = "mvt") %>%
     confint()

compTB
```

#### Plots

```{r}
( model4_Plot = ggplot(data = df_emm_model4 ) + 
  geom_point(aes(x=bTsfAlt, y=mean, color = bTsfAlt)) +
  geom_errorbar(aes(x=bTsfAlt, ymin = LCL, ymax = UCL, color = bTsfAlt)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years")) +
  labs(x="", y="Land surface temperature differnece (Kelvin)", subtitle = "difference Lst = post LST -pre LST") +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

)


```

```{r}
ggsave("../figures/2025-03-18_LstAllYrsBurnedRelative.png", plot = model4_Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```

```{r}
( model4_PlotRelative = ggplot(data = df_emm_model4 ) + 
  geom_point(aes(x=bTsfAlt, y=mean, color = bTsfAlt)) +
  geom_errorbar(aes(x=bTsfAlt, ymin = LCL, ymax = UCL, color = bTsfAlt)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years")) +
  labs(x="", y="Land surface temperature difference\n(Kelvin)") +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() +
  theme(legend.position="none")

)


```

```{r}
( model4_PlotLegendBottom = ggplot(data = df_emm_model4 ) + 
  geom_point(aes(x=bTsfAlt, y=mean, color = bTsfAlt)) +
  geom_errorbar(aes(x=bTsfAlt, ymin = LCL, ymax = UCL, color = bTsfAlt)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years")) +
  labs(x="", y="Land surface temperature difference\n(Kelvin)") +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() +
  theme(legend.position="bottom") +
    guides(color=guide_legend(nrow=3,byrow=TRUE))

)


```
```{r}
ggsave("../figures/2025-03-18_RelativeLSTAltLegendBottom.png", plot = model4_PlotLegendBottom, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```



## Grid Plots for Manuscript
modelfit2_PlotAbsolute
model4_PlotRelative

```{r}
( gridPlots = cowplot::plot_grid(modelfit2_PlotAbsolute, model4_PlotRelative, ncol=2, labels = c('A', 'B'), label_size = 12)  )
```

 
```{r}
( twoPanelLegend = ggdraw(xlim = c(0, 6), ylim = c(0, 6)) +
    draw_plot(gridPlots, x = 0, y = .5, width = 5.5, height = 5.5) + 
   draw_image("../images/20250313_RelativeLSTAltLegendBottom.PNG", scale = 1, x = 1, y = 0, width = 4, height = .65)  )
```

### Save Image

```{r eval=FALSE, include=FALSE}
ggsave("../figures/2025-03-11_LSTManuscriptFigure.png", plot = twoPanelLegend, dpi = 600, width = 6, height = 4, units = "in", bg='white')
```







