---
title: "Field Data Reburn for dNBR Extract in EE"
author: "Anna Talucci"
date: "2024-05-07"
output: html_document
---


# Overview
Organize field data to select for burn_depth and recent burns from 2016 to present. The goal is to pair these sites with SAR data.

Field data provided by Xanthe Walker from FiSL and Legacy Carbon Project. 

# Packages
```{r}
library(tidyverse)
library(sf)
```

# Data
```{r}
lcsites = read_csv("../data/fieldCsv/LC_SITES_11142023.csv")
lcdata = read_csv("../data/fieldCsv/LC_Data_09052023.csv")
fd = read_csv("../data/fieldCsv/FiSL_sitedata_BB2023.csv")
fp = read_csv("../data/fieldCsv/FINAL_PLOTS_01142024.csv")
```

# Look at data
```{r}
lcsites
lcdata

fp
```

above ground combustion trees + shrubs + cwd
below ground 

```{r}
fp %>% filter(unique_site_id.x==117 | unique_site_id.x ==174)
```

```{r}
lcdata
```
```{r}
glimpse(lcdata)
```

```{r}
lcdata[is.na(lcdata$Burn_Depth),]
```

# Assign unique row ideas

```{r}
( lcdata_id = lcdata %>% mutate(rowId = row_number()) %>% mutate(id = str_c(Researcher,"_",rowId)) )
```
```{r}
lcdata_id$...7[lcdata_id$...7 == "Before 1750"] <- 1750
```

```{r}
( lcsites_id = lcsites %>% mutate(rowId = row_number()) %>% mutate(researcher = "walkerLC") %>% mutate(id = str_c(researcher,"_",rowId)) )
```

```{r}
( fp_id = fp %>% mutate(rowId = row_number()) %>% mutate(researcher = "walker") %>% mutate(id = str_c(researcher,"_",rowId)) )
```


# Filter Data for reburn

## For final plot data
Notes: 
- Projection/Datum NAD83
- by select only fire from 2016 on we only loos 96 sites
- burn_depth has 197 NA


# Clean
```{r}
( fp1 = fp_id %>% 
    dplyr::select(id, unique_site_id.x, date.x, official_fire_scar, site, latitude, longitude, recent_burn, second_burn, triple_burn) %>%
    rename(siteId = unique_site_id.x, date=date.x, location = official_fire_scar, burnYr1=recent_burn, burnYr2=second_burn,  burnYr3=triple_burn, lat=latitude, lon=longitude) %>% 
    mutate(siteId = as.character(siteId),
           site = as.character(site))
  )
```

siteId, date, data, lat, lon, fireYr1, fireYr2, fireYr3
## For Legacy Carbon (LC) data
Notes: 
- Projection/Datum NAD83 
- by select only fire from 2016 on we only loos 96 sites
- burn_depth has 58 NA


```{r}
( lc1 = lcdata_id %>% 
    dplyr::select(id, Researcher, Site_Name, Sampling_Date, StudyArea, Year_of_Burn, `...7`, `...8`, Latitude, Longitude) %>%
    rename(date=Sampling_Date, siteId = Researcher, site= Site_Name, location = StudyArea, burnYr1=Year_of_Burn, burnYr2=`...7`,  burnYr3=`...8`, lat=Latitude, lon=Longitude) %>%
  drop_na(lat) %>%
    mutate(burnYr2 = as.numeric(burnYr2))
    
  )
```


```{r}
( lcsites1 = lcsites_id %>% 
    dplyr::select(id, unique_site_id, date, official_fire_scar, site, latitude, longitude, recent_burn, second_burn, triple_burn) %>%
    rename(siteId = unique_site_id, location = official_fire_scar, burnYr1=recent_burn, burnYr2=second_burn,  burnYr3=triple_burn, lat=latitude, lon=longitude) %>% 
    mutate(siteId = as.character(siteId),
           site = as.character(site))
    
  )
```

# Combine dataframes

```{r}
(df = bind_rows(fp1, lc1, lcsites1))
```

```{r}
unique(sort(df$burnYr1))
```

# To point shp
```{r}
df_pts = st_as_sf(df, coords = c("lon","lat"), crs = 4269, remove = FALSE)
```



# Reproject to WSG 84 
4326
```{r}
pts4326 = st_transform(df_pts, crs=4326)
```

### write to shapefile
```{r eval=FALSE, include=FALSE}
st_write(pts4326, "../outputs/spatialDataCreate/FislLcFieldPoints.shp", driver="ESRI Shapefile")
```



