---
title: "Figure Case study Maps"
author: "Anna Talucci"
date: "2024-11-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Skeleton for developing a map figure for FLARE UAS vegetation paper


# Packages
```{r}

library(tidyverse)
library(maptools)
library(mapproj)
library(sf)
library(terra)
library(ggspatial)
library(cowplot)
library(tidyterra)
library(terrainr)
```

library(rgdal)                                                                                                      
library(raster)
library(rgeos)
library(sp)

# Data

```{r}
eco = st_read("../data/spatialData/dire_ecoregs_caea/dire_ecoregs_caea.shp", "dire_ecoregs_caea")
```
```{r}
fires = st_read("../outputs/boundaries/firePerimeters/AkCaseStudyPerimeters.gpkg")
```

# Raster Data

## Raster Stack for 3-band RGB Image Satellite Image

```{r}
dnbr2015 <- rast("../data/caseStudyData/AK/dnbr2015.tif")
dnbr2018 <- rast("../data/caseStudyData/AK/dnbr2018.tif")
```

```{r}
dlst2015 <- rast("../data/caseStudyData/AK/dlst2015.tif")
dlst2018 <- rast("../data/caseStudyData/AK/dlst2018.tif")
```

```{r}
lst2014 <- rast("../data/caseStudyData/AK/lst2014.tif")
lst2015 <- rast("../data/caseStudyData/AK/lst2015.tif")
lst2016 <- rast("../data/caseStudyData/AK/lst2016.tif")
lst2017 <- rast("../data/caseStudyData/AK/lst2017.tif")
lst2018 <- rast("../data/caseStudyData/AK/lst2018.tif")
lst2019 <- rast("../data/caseStudyData/AK/lst2019.tif")
```

```{r}
color2016 <- rast("../data/caseStudyData/AK/colorImage2016.tif")
```

```{r}
dnbr2015
dlst2015
lst2015
lst2016
color2016
```

## World map
https://datacatalog.worldbank.org/dataset/world-bank-official-boundaries

```{r}
wrld = st_read("../data/spatialData/world/WB_countries_Admin0_10m.shp", "WB_countries_Admin0_10m")
```
### Subset for North America 

```{r}
na = wrld %>% filter(FORMAL_EN %in% c("Canada", "United States of America"))
```

# Quick Plots

```{r}
plot(st_geometry(eco))
```

```{r}
fires_wgs = st_transform(fires, crs=4326)
```

```{r}
fires_wgs
```
```{r}
plot(st_geometry(fires_wgs))
```

# Figures

## Fires 

```{r}

ggplot() + 
   geom_sf(data= fires_wgs, fill="transparent", aes(color=as.factor(year)), lwd=1) +
   scale_color_manual(values = c('#CB6BCEFF', '#468892FF')) +
  coord_sf(crs = 4326) +
  theme_minimal()
```

## dNBR

```{r}

ggplot() + 
  geom_spatraster(data = dnbr2015, aes(fill = dnbr)) +
  geom_spatraster(data = dnbr2018, aes(fill = dnbr)) +
  scale_fill_distiller(palette = "RdYlGn", na.value = "transparent") + 
  geom_sf(data= fires_wgs, fill="transparent", aes(color=as.factor(year)), lwd=1) +
   scale_color_manual(values = c('#CB6BCEFF', '#468892FF')) +
  coord_sf(crs = 4326) +
  theme_minimal()
```

```{r}

ggplot() + 
  geom_spatraster(data = dnbr2018, aes(fill = dnbr)) +
  scale_fill_distiller(palette = "RdYlGn", na.value = "transparent") + 
  geom_sf(data= fires_wgs, fill="transparent", aes(color=as.factor(year)), lwd=1) +
   scale_color_manual(values = c('#CB6BCEFF', '#468892FF')) +
  coord_sf(crs = 4326) +
  theme_minimal()
```

## dLST

```{r}

ggplot() + 
  geom_spatraster(data = dlst2015, aes(fill = dlst)) +
  #scale_fill_grass_c(palette = "celsius") +
   
  scale_fill_distiller(palette = "PuOr", na.value = "transparent") + 
  coord_sf(crs = 4326) +
  theme_minimal()
```

```{r}

ggplot() + 
  geom_spatraster(data = dlst2015, aes(fill = dlst)) +
  #scale_fill_grass_c(palette = "celsius") +
   
  scale_fill_distiller(palette = "PuOr", na.value = "transparent") + 
  coord_sf(crs = 4326) +
  theme_minimal()
```

## LST

```{r}

ggplot() + 
  geom_spatraster(data = lst2014, aes(fill = LST)) +
  #scale_fill_grass_c(palette = "celsius") +
   
  scale_fill_distiller(palette = "PuOr", na.value = "transparent") + 
  coord_sf(crs = 4326) +
  theme_minimal()
```

```{r}

ggplot() + 
  geom_spatraster(data = lst2015, aes(fill = LST)) +
  #scale_fill_grass_c(palette = "celsius") +
   
  scale_fill_distiller(palette = "PuOr", na.value = "transparent") + 
  coord_sf(crs = 4326) +
  theme_minimal()
```

```{r}

ggplot() + 
  geom_spatraster(data = lst2016, aes(fill = LST)) +
  #scale_fill_grass_c(palette = "celsius") +
   
  scale_fill_distiller(palette = "PuOr", na.value = "transparent") + 
  coord_sf(crs = 4326) +
  theme_minimal()
```

```{r}

ggplot() + 
  geom_spatraster(data = lst2017, aes(fill = LST)) +
  #scale_fill_grass_c(palette = "celsius") +
   
  scale_fill_distiller(palette = "PuOr", na.value = "transparent") + 
  coord_sf(crs = 4326) +
  theme_minimal()
```

```{r}

ggplot() + 
  geom_spatraster(data = lst2018, aes(fill = LST)) +
  #scale_fill_grass_c(palette = "celsius") +
   
  scale_fill_distiller(palette = "PuOr", na.value = "transparent") + 
  coord_sf(crs = 4326) +
  theme_minimal()
```

```{r}

ggplot() + 
  geom_spatraster(data = lst2019, aes(fill = LST)) +
  #scale_fill_grass_c(palette = "celsius") +
   
  scale_fill_distiller(palette = "PuOr", na.value = "transparent") + 
  coord_sf(crs = 4326) +
  theme_minimal()
```

## Color
```{r}
ggplot() +
  geom_spatraster_rgb(
  mapping = aes(),
  data=color2016,
  interpolate = TRUE,
  r = B4,
  g = B3,
  b = B2,
  alpha = 1) +
  coord_sf(crs = 4326) +
  theme_minimal()
```


# Reproject

## Define Projection

```{r}
st_crs(eco)
( ecozoneProjection = st_crs(eco)$proj4string )
```



# Maps

## Define zoom level

```{r zoomMap}
#### zoom to kamchatka, different projection ####

zoom_to <- c(-130.46, 64.93)  # ~ center of Kamchatka
zoom_level <- 3

# Lambert azimuthal equal-area projection around center of interest
#target_crs <- polarProj#sprintf('+proj=laea +lon_0=%f +lat_0=%f', zoom_to[1], zoom_to[2])


C <- 40075016.686   # ~ circumference of Earth in meters
x_span <- C / 2^(zoom_level+.1)
y_span <- C / 2^(zoom_level+.5)   # also sets aspect ratio

zoom_to_xy <- st_transform(st_sfc(st_point(zoom_to), crs = 4326), crs = ecozoneProjection)
zoom_to_xy

disp_window <- st_sfc(st_point(st_coordinates(zoom_to_xy - c(x_span / 2, y_span / 2))),
                      st_point(st_coordinates(zoom_to_xy + c(x_span / 2, y_span / 2))),
                      crs = ecozoneProjection)

```

## Create Map

```{r map, echo=FALSE}
ggplot() +
     geom_sf(data = na1, fill = "#F9F6EE", colour="#A9AB9D") +
  
    geom_sf(data = canada_2012_2015, aes(fill='#998ec3', color='#998ec3'), pch=21, stroke =1, alpha=.25) +
 geom_sf(data = eco, fill = NA, colour="black") +
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", linewidth = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank())

```








```{r}
plot_transects5 = ggplot() +
  ggRGB(kr_stack, r=1, g=2, b=3, ggLayer = TRUE, coord_equal = TRUE) +
  geom_sf(data = cherskiy, size = 4, shape = 24, fill = "grey70") +
     annotate(geom = "text", x = 161.76, y = 68.69, label = "Cherskiy",
    fontface = "bold", color = "grey70", size = 4) +
  geom_sf(data = plot100_pt_sf, aes(fill = site, color = site),size = 4, shape = 21, alpha= .9) +
  scale_color_manual(values=linecolors) +
  scale_fill_manual(values=fillcolors) +
  scale_y_continuous(breaks = seq(67.5, 69, by = .5), name = "") +
  scale_x_continuous(breaks = seq(161, 162), name = "") +
 
   annotation_north_arrow(location = "bl", which_north = "true", pad_x = unit(0.0, "cm"), pad_y = unit(0.7, "cm"), style = north_arrow_minimal(line_width = 1, line_col = "white", fill = "white", text_size = 0, text_face = NULL, text_family = "sans", text_col = "white")) +
 
 
  annotation_scale(location = "bl", width_hint = 0.25, text_size = 12, text_face = "bold", text_family = "sans", text_col = "white") +
 
  coord_sf(crs = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0", xlim = c(160.4, 162.4), ylim = c(67.8, 68.8), expand = TRUE) +
 
  theme_classic()+
 
  theme(plot.margin = unit(c(t = 0.2, r = 0.2, b = 0.2, l = 0.2), "cm")) +
    theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "white")) +
  theme(legend.position = "none")
 


plot_transects5

```


# Combining Maps

We combine our study area map with the inset map for a final map figure

```{r}
line1 = data.frame(x1 = 2.86, x2 = 5.2, y1 = 6.6, y2 = 6.25)
line2 = data.frame(x1 = 2.87, x2 = 5.1, y1 = 4.1, y2 = 6.15)
```


```{r fig.height=7, fig.width=6}
fig_maps6 = ggdraw(xlim = c(0, 6), ylim = c(0, 7)) +
    draw_plot(russia_map3, x = 0, y = 3.5, width = 6, height = 3.5) +
  draw_plot(plot_transects5, x = 0.8, y = 3.75, width = 2.2, height = 3.2,scale =0.9) +

  draw_image("../images/2020-05-22_UAV-Veg-RGB-NDVI-CN.PNG", scale = 0.85, x = 0, y = 0.2, width = 6, height = 3.5) +
  draw_label("Russia", x = 3.5, y = 6, fontfamily = "sans", fontface = "bold", color = "black", size = 12) +
  draw_label("ANS", x = 2.43, y = 4.5, fontfamily = "sans", fontface = "bold", color = "white", size = 10) +
  draw_label("Alnus", x = 1.95, y = 5.1, fontfamily = "sans", fontface = "bold", color = "white", size = 10) +
  draw_label("BP", x = 1.4, y = 5.6, fontfamily = "sans", fontface = "bold", color = "white", size = 10) +
  draw_label("CN", x = 2.15, y = 6.5, fontfamily = "sans", fontface = "bold", color = "white", size = 10) +
  draw_label("161째E", x = 1.67, y = 4.09, fontfamily = "sans", fontface = "plain", color = "black", size = 7) +
   draw_label("162째E", x = 2.47, y = 4.09, fontfamily = "sans", fontface = "plain", color = "black", size = 7) +
  draw_label("68째N", x = 0.95, y = 4.69, fontfamily = "sans", fontface = "plain", color = "black", size = 7, angle = 90) +
  draw_label("68.5째N", x = 0.95, y = 5.74, fontfamily = "sans", fontface = "plain", color = "black", size = 7, angle = 90) +

  draw_label("A", x = 0.3, y = 6.8, fontfamily = "sans", fontface = "bold", color = "black", size = 14) +
   draw_label("B", x = 0.3, y = 3.3, fontfamily = "sans", fontface = "bold", color = "black", size = 14) +
   draw_label("C", x = 3.0, y = 3.3, fontfamily = "sans", fontface = "bold", color = "black", size = 14) +
  geom_segment(data =line1, aes(x = x1, y = y1, xend = x2, yend = y2), linetype = "dotted",  colour = "#000080", size= 1, alpha = .5) +
  geom_segment(data =line2, aes(x = x1, y = y1, xend = x2, yend = y2), linetype = "dotted",  colour = "#000080", size = 1, alpha = .5)

fig_maps6
```

```{r}
ggsave("../figures/2020-07-12_UAV-Veg-NDVI-RGB.png", plot = fig_maps6, width = 6, height = 7, dpi = 600)
```