---
title: "Figure Map"
author: "Anna Talucci"
date: "2024-05-15"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```


# Overview
Creat over view map with permafrost extent and summary of site data.
Notes:From PCN-Fire-Sysnthesis "Manuscript Map"
[Better colors for Mapping](https://www.esri.com/arcgis-blog/products/js-api-arcgis/mapping/better-colors-for-better-mapping/)
# Packages
```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(cowplot)
library(ggpubr)
library(ggnewscale)
library(scales)
library(ggrepel)
library(patchwork)
```



# Data
```{r}
above = st_read("../data/spatialData/ABoVE_Study_Domain/ABoVE_Study_Domain.shp", "ABoVE_Study_Domain")
```

```{r}
res = st_read("../data/spatialData/RESOLVE_Ecoregions_and_Biomes/Biomes_and_Ecoregions_2017.shp", "Biomes_and_Ecoregions_2017")
```

```{r}
( na_boreal = res %>% filter(BIOME_NUM==6) %>% filter(REALM=="Nearctic"))
```

```{r}
above
( core = above %>% filter(Region=="Core Region") )
```
    

```{r}
unique(na_boreal$ECO_NAME)
```
"Interior Alaska-Yukon lowland taiga", "Northern Canadian Shield taiga", "Northwest Territories taiga", 

## World map
https://datacatalog.worldbank.org/dataset/world-bank-official-boundaries

```{r}
wrld = st_read("../data/spatialData/world/WB_countries_Admin0_10m.shp", "WB_countries_Admin0_10m")
```


```{r}
na = wrld %>% filter(FORMAL_EN %in% c("Canada", "United States of America"))
```









## Get Eco Projection

```{r}
st_crs(res)
( ecozoneProjection = st_crs(res)$proj4string )
```


```{r}
st_crs(core)

```

```{r}
projection = "+proj=aea +lon_0=-108.984375 +lat_1=48.2911002 +lat_2=69.1799501 +lat_0=58.7355252 +datum=WGS84 +units=m +no_defs"
```

# Reproject

```{r}
na_pp = st_transform(na, crs = projection)
boreal_pp = st_transform(na_boreal, crs = projection)
core_pp = st_transform(core, crs = projection)
```
```{r}
core_pp
```
```{r}
( naBoreal_albers = st_transform(na_boreal, crs = st_crs(core)) )
( na_albers = st_transform(na, crs = st_crs(core)) )
```

```{r}
ecozone_core = st_join(naBoreal_albers, core, largest = TRUE) %>% drop_na(Region)
```

```{r}
ecozone_core
```
# Palettes

```{r}
brltnd_palette = c('#4a6741', '#3d3021')
```

# Maps

## Zoom Level
## Map area zoom
```{r}
#### zoom to kamchatka, different projection ####

zoom_to <- c(-97.93, 63.84)  # ~ center of Kamchatka
zoom_level <- 2.6

# Lambert azimuthal equal-area projection around center of interest
#target_crs <- polarProj#sprintf('+proj=laea +lon_0=%f +lat_0=%f', zoom_to[1], zoom_to[2])


C <- 40075016.686   # ~ circumference of Earth in meters
x_span <- C / 2^(zoom_level+.1)
y_span <- C / 2^(zoom_level+.6)   # also sets aspect ratio

zoom_to_xy <- st_transform(st_sfc(st_point(zoom_to), crs = 4326), crs = projection)
zoom_to_xy

disp_window <- st_sfc(st_point(st_coordinates(zoom_to_xy - c(x_span / 2, y_span / 2))),
                      st_point(st_coordinates(zoom_to_xy + c(x_span / 2, y_span / 2))),
                      crs = projection)

```

```{r}
ggplot() +
     geom_sf(data = na_pp, fill = "#F9F6EE", colour="#F9F6EE") +

    geom_sf(data = core_pp, fill="black", color="black")
```

# Build Overview map by Ecozone

```{r map, echo=FALSE}
( basicBorealMap = ggplot() +
     geom_sf(data = na_pp, fill = "#F9F6EE", colour="#F9F6EE") +
    geom_sf(data = boreal_pp, fill='#4a6741', color='#4a6741') +
    
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", linewidth = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black")) 
)

```


```{r eval=FALSE, include=FALSE}
ggsave("../figures/MapBasicBorealPoster.png", plot = basicBorealMap, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```


```{r map, echo=FALSE}
( coreMap = ggplot() +
     geom_sf(data = na_pp, fill = "#F9F6EE", colour="#F9F6EE") +
    geom_sf(data = boreal_pp, fill='#4a6741', color='gray') +
    geom_sf(data = core_pp, fill=NA, color='black') +
    
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", linewidth = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black")) 
)

```

```{r}
#### zoom to kamchatka, different projection ####

zoom_to <- c(-97.93, 63.84)  # ~ center of Kamchatka
zoom_level <- 2.6

# Lambert azimuthal equal-area projection around center of interest
#target_crs <- polarProj#sprintf('+proj=laea +lon_0=%f +lat_0=%f', zoom_to[1], zoom_to[2])


C <- 40075016.686   # ~ circumference of Earth in meters
x_span <- C / 2^(zoom_level+.1)
y_span <- C / 2^(zoom_level+.6)   # also sets aspect ratio

zoom_to_xy <- st_transform(st_sfc(st_point(zoom_to), crs = 4326), crs = st_crs(core))
zoom_to_xy

disp_window <- st_sfc(st_point(st_coordinates(zoom_to_xy - c(x_span / 2, y_span / 2))),
                      st_point(st_coordinates(zoom_to_xy + c(x_span / 2, y_span / 2))),
                      crs = projection)

```
```{r, echo=FALSE}
( ecozoneCoreMap = ggplot() +
     geom_sf(data = na_albers, fill = "#F9F6EE", colour="#F9F6EE") +
    geom_sf(data = ecozone_core, fill='#4a6741', color='gray') +
    geom_sf(data = above, fill=NA, color="black") +
  
  #coord_sf() +
  
  coord_sf(xlim = st_coordinates(disp_window)[,'X'], ylim = st_coordinates(disp_window)[,'Y']) +
  theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", linewidth = 0.5), 
        panel.background = element_rect(fill = "#C5D8D7")) +
  theme(plot.margin=unit(c(0,0,0,0), "mm")) + 
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black")) 
)

```

```{r eval=FALSE, include=FALSE}
ggsave("../figures/MapBasicBorealAboVECore.png", plot = ecozoneCoreMap, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```





# Raster data Reburns






# Combine Map and Legend

## Resolve

### Grid map with legend Bottom
```{r fig.height=6, fig.width=6}
MapBottomLegendResolve = ggdraw(xlim = c(0, 6), ylim = c(0, 6)) +
    draw_plot(overviewMapRes, x = 0, y = 0, width = 6, height = 6) + 
  draw_image("../images/legendBottomResolve.PNG", scale = 0.99, x = .2, y = -0.1, width = 6, height = 1) 
```

### Save Image
```{r eval=FALSE, include=FALSE}
ggsave("../figures/PCNMapOverviewBottomLegendResolve.png", plot = MapBottomLegendResolve, dpi = 600, width = 6, height = 6, bg='white')
```


