---
title: "Google Drive HDF Images Explore"
author: "Anna Talucci"
date: "2024-12-20"
output: html_document
---
# overview

notes:
Need to move dhf broad albedo image into single google drive folder
Then transfer hdf albedo to cloud bucket
Then use existing code to convert hdf to geotiff
# packages

```{r}
library(googledrive)
library(tidyverse)
```

# Authenticate your Google account

```{r}
drive_deauth()
```


```{r}
drive_auth()
```



```{r}
# Replace with the name or ID of your source folder
source_folder <- "/content/drive/My Drive//Landsat8Collection1ScenesAK/albedo"  # Replace with your source folder name
destination_folder <- "'/content/drive/My Drive//Landsat8Collection1ScenesAK/albedo/broad"  # Replace with your destination folder name

# List files in the source folder
source_files <- drive_ls(path = source_folder)

# Filter files containing "broad" in their names
broad_files <- source_files[grepl("broad", source_files$name, ignore.case = TRUE), ]
```



--------------------

# List files in the specified folder

```{r}
files_in_folder <- drive_ls(path='~/Landsat8Collection1ScenesAK/albedo/', pattern = "albedo_broad")
```
number of broad files = 1039

```{r}
path64 <- drive_ls(path='~/Landsat8Collection1ScenesAK/albedo/byPath/path64/', pattern = "albedo_broad")
path65 <- drive_ls(path='~/Landsat8Collection1ScenesAK/albedo/byPath/path65/', pattern = "albedo_broad")
path66 <- drive_ls(path='~/Landsat8Collection1ScenesAK/albedo/byPath/path66/', pattern = "albedo_broad")
path67 <- drive_ls(path='~/Landsat8Collection1ScenesAK/albedo/byPath/path67/', pattern = "albedo_broad")
path68 <- drive_ls(path='~/Landsat8Collection1ScenesAK/albedo/byPath/path68/', pattern = "albedo_broad")
path69 <- drive_ls(path='~/Landsat8Collection1ScenesAK/albedo/byPath/path69/', pattern = "albedo_broad")
```

```{r}
path64
path65
path66
path67
path68
path69
```


# Extract the file names



```{r}
file_names <- files_in_folder$name
```
1039
```{r}
file_names
```


# New Folder

```{r}
folder <- drive_mkdir(name= 'broad', path="~/Landsat8Collection1ScenesAK/albedo/")
```


# Test Run
```{r}
filesTest1 <- drive_ls(path='~/Landsat8Collection1ScenesAK/albedo/test1/', pattern="albedo_broad")
```

```{r}
# Get folder metadata by name
folder_name <- "~/Landsat8Collection1ScenesAK/albedo/test1"

# Use drive_find to search for the folder by name (it can return files or folders)
folder_metadata <- drive_find(pattern = folder_name, type = "folder")

# View the folder metadata (including the folder ID)
print(folder_metadata)
```

```{r}
# Extract the folder ID from the metadata
folder_id <- folder_metadata$id[1]  # Assuming the first result is the correct folder
print(folder_id)
```

```{r}
parent_folder_id <- file_metadata$parents[[1]]
print(parent_folder_id)
```

```{r}
folderTest1$id[5]
```
```{r}
drive_get(path='~/Landsat8Collection1ScenesAK/albedo/test1')
```

```{r}
files_to_move <- drive_ls(path='~/Landsat8Collection1ScenesAK/albedo/test1/', pattern = "albedo_broad")
```
```{r}
folderTest1
```

```{r}
folderTest1$id
```

```{r}
file_names_test1 <- folderTest1$name
```

```{r}
file_names_test1
```
```{r}
folderTest2 <- drive_mkdir(name= 'test2', path="~/Landsat8Collection1ScenesAK/albedo/test1/")
```

# Get the folder ID of the newly created folder
```{r}
destination_folder_id <- folderTest2$id
destination_folder_id
```
```{r}
source_folder_id <- folderTest1$id[1]
source_folder_id
```

# Move each file to the new folder
```{r}
# Define the source and destination folder IDs (replace with your folder IDs)
source_folder_id <- folderTest1$id[1]  # Folder where the files are currently
destination_folder_id <- folderTest2$id # Folder where you want to move the files

# List all files in the source folder
files_in_source <- drive_ls(as_id(source_folder_id))

# Define a pattern to match filenames (e.g., files with a certain extension or substring)
pattern <- "albedo_broad"  # Replace with the pattern you're searching for

# Filter files based on the pattern (e.g., files containing "report" in their name)
files_to_move <- files_in_source %>%
  filter(grepl(pattern, name))

# Move the files to the destination folder
for (file in files_to_move$id) {
  drive_move(file, path = as_id(destination_folder_id))
  cat("Moved file:", file, "\n")
}
```









# Verify if the files have been moved (optional)
```{r}
moved_files <- drive_find(q = paste0("'", folderTest2_id, "' in parents"))
print(moved_files)
```
