---
title: "LST Pre Post"
author: "Anna Talucci"
date: "2025-03-04"
output: html_document
---

```{r clear environment, include=FALSE}
rm(list=ls())
```

# Overview


# Packages

```{r}
library(tidyverse)
library(sf)
library(emmeans) 
library(statmod)
library(car)

library(MASS)
library(cplm)
library(aod)
```

# Data

```{r}
aklst = read_csv('../outputs/cleanedSamplePoints/LstBoreal2013-2022.csv')
```
```{r}
perimeters = st_read("../outputs/boundaries/firePerimeters/akFirePerimeters_1984_2020.shp", "akFirePerimeters_1984_2020")
```


```{r}
aklst
```


```{r}
( burned_lst = aklst %>% filter(countYr!=0) )
( unburned_lst = aklst %>% filter(countYr==0) )
```


# To Points

```{r eval=FALSE, include=FALSE}
df_to_sf <- function(x){
  st_as_sf(x, coords = c("longitude","latitude"), crs = 4326, remove = FALSE)
}
```

```{r eval=FALSE, include=FALSE}
( burned_pts = burned_lst %>% df_to_sf() )
```

```{r eval=FALSE, include=FALSE}
( perimeters_4 = perimeters %>% select(Event_ID, year, month, day, geometry) )
```

# Transform projections

```{r eval=FALSE, include=FALSE}
burned_pts_NAD83 = st_transform(burned_pts, crs = st_crs(perimeters))
```

```{r eval=FALSE, include=FALSE}
burned_pts_NAD83
```

# Join with fire perimeters for fire years
 **NOT WORKING**
```{r eval=FALSE, include=FALSE}
join = st_join(burned_pts_NAD83, perimeters_4, left=TRUE)
```

```{r eval=FALSE, include=FALSE}
join
```


# Clean and Organize data for model

```{r}
burned_lst
```

```{r}
sort(unique(burned_lst$maxYr))
```

## Clean with pre/post
```{r}

clean_BurnedLst_tsf <- burned_lst %>%
  dplyr::select(ID, ECO_NAME, countYr, maxYr, mean_2013:mean_2022, burned, TSF, TimeSinceFire) %>%
  filter(TSF %in% 1:5) %>%
  mutate(preYear = maxYr - 1,
         fireYear = maxYr,
         postYear = maxYr + TSF) %>%
  pivot_longer(
    cols = starts_with("mean"),
    names_to = "LstYear",
    values_to = "LstValue"
  ) %>%
  separate(LstYear, into = c("LstFun", "LstYear"), sep = "_", convert = TRUE, extra = "merge")

```



```{r}
# Filter and check matches
df_filtered <- clean_BurnedLst_tsf %>%
  #group_by(ID) %>%  # Filter by group
  rowwise() %>%
  mutate(match =  (preYear %in% LstYear) |  (postYear %in% LstYear)) %>%
  filter(match) %>% # Keep only rows where any match occurs
  mutate(match_type = case_when(
    preYear %in% LstYear ~ "pre",
    postYear %in% LstYear ~ "post")) %>% 
  group_by(ID) #%>%
  #pivot_wider(names_from = match_type, values_from = LstValue)

# View results
print(df_filtered)
```



```{r}
( df_wider = df_filtered %>%
  pivot_wider(names_from = match_type, values_from = c(LstYear, LstValue))
)

```


```{r}
pre_post_na = df_wider %>% 
  filter(is.na(LstValue_pre) | is.na(LstValue_post))

print(pre_post_na)
```

```{r}
( data4analysis = df_wider %>% 
  drop_na(LstValue_pre) %>% 
  drop_na(LstValue_post) %>%
    mutate(diffLst = LstValue_pre - LstValue_post)
)
```


```{r}
( recent = data4analysis %>% 
  filter(TimeSinceFire %in% c("0", "1", "2-5")) %>% 
  filter(burned %in% c("Unburned", "One", "Two", "Three")) %>%
  mutate(TimeSinceFire = factor(TimeSinceFire, levels = c( "0", "1", "2-5")),
         burned = factor(burned, levels = c( "Unburned", "One", "Two", "Three")),
         bTSF = paste0(burned,"_", TimeSinceFire)) )
```


# Model Recent time since fire effects

```{r}
fit1 = glm(diffLst ~ bTSF, data = recent, family = gaussian)
```


### Residuals: 

```{r eval=FALSE, include=FALSE}
recent$fit1.res = resid(fit1, type = "pearson")
recent$fit1.fit = fitted(fit1)

sum(residuals(fit1, type="pearson")^2)/fit1$df.res

qplot(fitted(fit1), resid(fit1)) + theme_bw()

```


```{r}
summary(fit1)
```

```{r}
emmfit1 = emmeans(fit1, specs = ~ bTSF)
emmfit1
```
```{r}
dfemmfit1 = as.data.frame(emmfit1)

dfemmfit1.1 = dfemmfit1 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTSF, levels = c("Unburned_0", 'One_1', 'One_2-5',  'Two_1', 'Two_2-5', 'Three_1','Three_2-5')))

dfemmfit1.1
```


```{r}
allDataFit1Plot = ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#DB4743FF','#F5AF4DFF',  '#DB4743FF','#F5AF4DFF', '#DB4743FF','#F5AF4DFF'), labels=c("one year", "2-5 years", "one year", "2-5 years", "one year", "2-5 years")) +
  labs(x="", y="Land surface temperature differnece (Kelvin)", subtitle = "pre LST - post LST") +
  scale_x_discrete(labels= c("One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

allDataFit1Plot
```

```{r}
ggsave("../figures/LSTDifferencerecentBoreal_20250305.png", plot = allDataFit1Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```



```{r}
(recentLSTPlot = ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c("One year", "2-5 years", "One year", "2-5 years", "One year", "2-5 years")) +
  labs(x="Times burned", y="Land surface Temperature Difference (Kelvin)", subtitle = "Difference = pre LST - post LST") +
  scale_x_discrete(labels= c("One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() +
   theme( legend.position="right", 
        legend.text=element_text(size=14), 
        legend.title=element_text(size=16),
        legend.key = element_blank()) +
    theme(axis.title.y = element_text(size = 16, hjust = 0.5, vjust = 1.1),
          axis.title.x = element_text(size = 16, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black")) 
)
```

```{r}
ggsave("../figures/20250306_LSTDifferencerecentBorealCompareOneTimeBurnedPPTSlide.png", plot = recentLSTPlot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```


## Testing difference in means

```{r}
emmfit1
```

```{r}
One_1 = c(1, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0)
Three_1 = c(0, 0, 1, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 1, 0, 0)
Two_1 = c(0, 0, 0, 0, 1, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1)
```

```{r}
contrast(emmfit1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))
```
```{r}
compTB = contrast(emmfit1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()

compTB
```



# Standard unburned LST value for difference
 Standardized chronosequence with 2022 as post fire year
## Calculate prefire LST from unburned

```{r}
unburned_lst
```

```{r}
unique(unburned_lst$ECO_NAME)
```

```{r}
calculate_refLst <- function(year) {
  unburned_lst %>% 
    group_by(ECO_NAME) %>% 
    summarise(refLst = round(mean(get(paste0("mean_", year))), 1))
}

# Create a list of reference datasets for multiple years
years <- 2014:2022
ref_data <- setNames(lapply(years, calculate_refLst), paste0("ref", years))
ref_data
```

## Clean with Standard pre year

```{r}
burned_lst
```

```{r}
sort(unique(burned_lst$maxYr))
```

```{r}
burned_lst %>% mutate(tsf2022 = 2022-maxYr) %>% group_by(as.factor(tsf2022)) %>% summarise(n=n())
```

```{r}
calculate_standardPre <- function(year) {
  ref_year <- ref_data[[paste0("ref", year)]]
  burned_lst %>%
    filter(maxYr < year) %>%
    dplyr::select(ID:maxYr, !!sym(paste0("mean_", year)), burned) %>%
    mutate(refLst = case_when(
      ECO_NAME %in% ref_year$ECO_NAME ~ ref_year$refLst[match(ECO_NAME, ref_year$ECO_NAME)]
    )) %>%
    mutate(
      tsf = year - maxYr,
      diffLst = refLst - get(paste0("mean_", year))
    ) %>%
    mutate(
      TimeSinceFire = case_when(
        tsf == 1 ~ "1",
        tsf %in% 2:5 ~ "2-5",
        tsf %in% 6:10 ~ "6-10",
        tsf %in% 11:20 ~ "11-20",
        tsf %in% 21:30 ~ "21-30",
        tsf %in% 31:36 ~ "31-36",
        TRUE ~ "Other"
      ),
      tsfStd = case_when(
        tsf %in% 1:5 ~ "1-5",
        tsf %in% 6:10 ~ "6-10",
        tsf %in% 11:20 ~ "11-20",
        tsf %in% 21:36 ~ "21+",
        TRUE ~ "Other"
      ),
      tsfAlt = case_when(
        tsf == 1 ~ "1",
        tsf %in% 2:5 ~ "2-5",
        tsf %in% 6:15 ~ "6-15",
        tsf %in% 16:36 ~ "16+",
        TRUE ~ "Other"
      )
    ) %>%
    mutate(across(
      c(TimeSinceFire),
      ~ ordered(., levels = c("1", "2-5", "6-10", "11-20", "21-30", "31-36"))
    )) %>%
     mutate(across(
      c(tsfStd),
      ~ ordered(., levels = c("1-5", "6-10", "11-20", "21+"))
    )) %>%
     mutate(across(
      c(tsfAlt),
      ~ ordered(., levels = c("1", "2-5", "6-15", "16+"))
    )) %>%
    mutate(bTSF = paste0(burned, "_", TimeSinceFire),
           bTsfStd = paste0(burned, "_", tsfStd),
           bTsfAlt = paste0(burned, "_", tsfAlt),
           sampleYr = year) %>%
    rename(lstMean = !!sym(paste0("mean_", year)))
}

# Generate standard pre-year data for multiple years
standardPre_data <- setNames(lapply(years, calculate_standardPre), paste0("standardPre", years))

```

```{r}
standardPre_data
```


```{r}
final_df <- bind_rows(
  standardPre_data$standardPre2022, standardPre_data$standardPre2021, standardPre_data$standardPre2020, standardPre_data$standardPre2019, 
  standardPre_data$standardPre2018, standardPre_data$standardPre2017, standardPre_data$standardPre2016, standardPre_data$standardPre2015, 
  standardPre_data$standardPre2014
)

```

```{r}
final_df
```

```{r}
recent_2014_2022 = final_df %>% filter(TimeSinceFire %in% c("1", "2-5"))
old_2014_2022 = final_df %>% filter(TimeSinceFire %in% c("6-10", "11-20", "21-30", "31-36"))
```

# Model Recent time since fire effects
 Match AGU and ABOVE Analaysis format
```{r}
model1 = glm(diffLst ~ bTSF, data = recent_2014_2022, family = gaussian)
```



### Residuals: 

```{r eval=FALSE, include=FALSE}
recent$model1.res = resid(model1, type = "pearson")
recent$model1.fit = fitted(model1)

sum(residuals(model1, type="pearson")^2)/model1$df.res

qplot(fitted(model1), resid(model1)) + theme_bw()

```

```{r}
emm_model1 = emmeans(model1, specs = ~ bTSF)
emm_model1

df_emm_model1 = as.data.frame(emm_model1)

df_emm_model1 = df_emm_model1 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTSF, levels = c( "One_1", "One_2-5",  "Two_1",  "Two_2-5",  "Three_1", "Three_2-5"))
                  )

df_emm_model1
```

## Plots
```{r}
recent_model1_plot = ggplot(data = df_emm_model1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#DB4743FF','#F5AF4DFF',  '#DB4743FF','#F5AF4DFF', '#DB4743FF','#F5AF4DFF'), labels=c("one year", "2-5 years", "one year", "2-5 years", "one year", "2-5 years")) +
  labs(x="", y="Land surface temperature differnece (Kelvin)", subtitle = "difference Lst = pre LST - post LST") +
  scale_x_discrete(labels= c("One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

recent_model1_plot
```

```{r}
ggsave("../figures/20250306_LSTDifferenceRecentBoreal.png", plot = recent_model1_plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```



```{r}
(recent_model1_plot2 = ggplot(data = df_emm_model1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c("One year", "2-5 years")) +
  labs(x="Times burned", y="Land surface Temperature Difference (Kelvin)", subtitle = "Difference = pre LST - post LST") +
  scale_x_discrete(labels= c("One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() +
   theme( legend.position="right", 
        legend.text=element_text(size=14), 
        legend.title=element_text(size=16),
        legend.key = element_blank()) +
    theme(axis.title.y = element_text(size = 16, hjust = 0.5, vjust = 1.1),
          axis.title.x = element_text(size = 16, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black")) 
)
```

```{r}
ggsave("../figures/20250312_LSTDifferencerecentRecentBorealPPTSlide.png", plot = recent_model1_plot2, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```






## Testing difference in means

```{r}
emm_model1
```

```{r}
One_1 = c(1, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0)
Three_1 = c(0, 0, 1, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 1, 0, 0)
Two_1 = c(0, 0, 0, 0, 1, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1)

contrast(emm_model1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))

contrast(emm_model1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()


```



# Model OLD time since fire effects

```{r}
unique(old$TimeSinceFiretsf)
```

```{r}
model2 = glm(diffLst ~ bTSF, data = old_2014_2022, family = gaussian)
```

```{r}
summary(model2)
```

### Residuals: 

```{r eval=FALSE, include=FALSE}
recent$model2.res = resid(model2, type = "pearson")
recent$model2.fit = fitted(model2)

sum(residuals(model2, type="pearson")^2)/model2$df.res

qplot(fitted(model2), resid(model2)) + theme_bw()

```


```{r}
emm_model2 = emmeans(model2, specs = ~ bTSF)
emm_model2

df_emm_model2 = as.data.frame(emm_model2)

df_emm_model2 = df_emm_model2 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTSF, levels = c( "One_6-10", "One_11-20", "One_21-30", "One_31-36",  "Two_6-10",  "Two_11-20","Two_21-30", "Two_31-36",  "Three_6-10", "Three_11-20", "Three_21-30"))
                  )

df_emm_model2
```

## Plots

```{r}
model2_old_Plot = ggplot(data = df_emm_model2 ) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#02c39a','#00a896',  '#028090','#05668d', '#02c39a','#00a896',  '#028090','#05668d', '#02c39a','#00a896',  '#028090'), labels=c("6-10 year", "11-20 years", "21-30 year", "30-36 years","6-10 year", "11-20 years", "21-30 year", "30-36 years", "6-10 year", "11-20 years", "21-30 year")) +
  labs(x="", y="Land surface temperature differnece (Kelvin)", subtitle = "difference Lst = pre LST - post LST") +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

model2_old_Plot
```

```{r}
ggsave("../figures/20250306_LSTDifferenceOLDBoreal.png", plot = model2_old_Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```



# Model 3 with Standard Time Since Fire groups

```{r}
model3 = glm(diffLst ~ bTsfStd, data = data2014_2022, family = gaussian)
```

```{r}
summary(model3)
```

### Residuals: 

```{r eval=FALSE, include=FALSE}
recent$model3.res = resid(model3, type = "pearson")
recent$model3.fit = fitted(model3)

sum(residuals(model3, type="pearson")^2)/model3$df.res

qplot(fitted(model3), resid(model3)) + theme_bw()

```


```{r}
emm_model3 = emmeans(model3, specs = ~ bTsfStd)
emm_model3

df_emm_model3 = as.data.frame(emm_model3)

df_emm_model3 = df_emm_model3 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTsfStd = factor(bTsfStd, levels = c("One_1-5", "One_6-10", "One_11-20", "One_21+", "Two_1-5", "Two_6-10", "Two_11-20", "Two_21+", "Three_1-5", "Three_6-10", "Three_11-20", "Three_21+"))
                  )

df_emm_model3
```

## Plots

```{r}
( Model3_Plot = ggplot(data = df_emm_model3 ) + 
  geom_point(aes(x=bTsfStd, y=mean, color = bTsfStd)) +
  geom_errorbar(aes(x=bTsfStd, ymin = LCL, ymax = UCL, color = bTsfStd)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("1-5 years", "6-10 years", "11-20 years", "21+ years","1-5 years", "6-10 years", "11-20 years", "21+ years", "1-5 years", "6-10 years", "11-20 years", "21+ years")) +
  labs(x="", y="Land surface temperature differnece (Kelvin)", subtitle = "difference Lst = pre LST - post LST") +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

)


```

```{r}
ggsave("../figures/20250306_LSTDifferenceBorealStdTsfBins.png", plot = Model3_Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```


## Testing difference in means

```{r}
emm_model3
```

```{r}
One_1_5 =  c(1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_6_10 = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_11_20 =c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_21_36 =c(0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0)
Two_1_5 =  c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0)
Two_6_10 = c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
Two_11_20 =c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
Two_21_36 =c(0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
Three_1_5 =c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
Three_6_10 =c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
Three_11_20 =  c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
Three_21_36 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
```




**RECALCULATE**
```{r}
contrast(emm_model3, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))
```
```{r}
compTB = contrast(emm_model3, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()

compTB
```




# Model 4 with Alternate Time Since Fire groups

```{r}
model4 = glm(diffLst ~ bTsfAlt, data = final_df, family = gaussian)
```

```{r}
summary(model4)
```

### Residuals: 

```{r eval=FALSE, include=FALSE}
recent$model4.res = resid(model4, type = "pearson")
recent$model4.fit = fitted(model4)

sum(residuals(model4, type="pearson")^2)/model4$df.res

qplot(fitted(model4), resid(model4)) + theme_bw()

```


```{r}
emm_model4 = emmeans(model4, specs = ~ bTsfAlt)
emm_model4

df_emm_model4 = as.data.frame(emm_model4)

df_emm_model4 = df_emm_model4 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTsfStd = factor(bTsfAlt, levels = c("One_1", "One_2-5", "One_6-15", "One_16+", "Two_1", "Two_2-5", "Two_6-15", "Two_16+", "Three_1", "Three_2-5", "Three_6-15", "Three_16+"))
                  )

df_emm_model4
```

## Plots

```{r}
( model4_Plot = ggplot(data = df_emm_model4 ) + 
  geom_point(aes(x=bTsfAlt, y=mean, color = bTsfAlt)) +
  geom_errorbar(aes(x=bTsfAlt, ymin = LCL, ymax = UCL, color = bTsfAlt)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years")) +
  labs(x="", y="Land surface temperature differnece (Kelvin)", subtitle = "difference Lst = pre LST - post LST") +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

)


```

```{r}
ggsave("../figures/20250312_LSTDifferenceBorealAltTsfBins.png", plot = model4_Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```


## Testing difference in means

```{r}
emm_model4
```

```{r}
One_1 =  c(1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_6_15 =c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_16_36 =c(0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0)
Two_1 =  c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
Two_6_15 =c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0)
Two_16_36 =c(0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
Three_1 =c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
Three_2_5 =c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
Three_6_15 =  c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
Three_16_36 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)
```




**RECALCULATE**
```{r}
contrast(emm_model4, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36))
```
```{r}
compTB = contrast(emm_model4, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()

compTB
```











