---
title: "LST Pre Post"
author: "Anna Talucci"
date: "2025-03-04"
output: html_document
---

```{r clear environment, include=FALSE}
rm(list=ls())
```

# Overview


# Packages

```{r}
library(tidyverse)
library(sf)
library(emmeans) 
library(statmod)
library(car)

library(MASS)
library(cplm)
library(aod)
```

# Data

## Read in

```{r}
aklst = read_csv('../outputs/cleanedSamplePoints/LstBoreal2013-2022.csv')
```
```{r}
perimeters = st_read("../outputs/boundaries/firePerimeters/akFirePerimeters_1984_2020.shp", "akFirePerimeters_1984_2020")
```

## View

```{r}
aklst
```

## Split data into burned unburned

```{r}
( burned_lst = aklst %>% filter(countYr!=0) )
( unburned_lst = aklst %>% filter(countYr==0) )
```

# Clean and Organize data for model

```{r}
burned_lst
```

```{r}
sort(unique(burned_lst$maxYr))
```

```{r}
burned_lst %>% mutate(tsf2022 = 2022-maxYr) %>% group_by(as.factor(tsf2022)) %>% summarise(n=n())
```

## ref year for pre-post calculation

```{r}
calculate_refLst <- function(year) {
  unburned_lst %>% 
    group_by(ECO_NAME) %>% 
    summarise(refLst = round(mean(get(paste0("mean_", year))), 1))
}

# Create a list of reference datasets for multiple years
years <- 2014:2022
ref_data <- setNames(lapply(years, calculate_refLst), paste0("ref", years))
ref_data
```

## Function for tsf bins
```{r}
calculate_standardPre <- function(year) {
  ref_year <- ref_data[[paste0("ref", year)]]
  burned_lst %>%
    filter(maxYr < year) %>%
    dplyr::select(ID:maxYr, !!sym(paste0("mean_", year)), burned) %>%
    mutate(refLst = case_when(
      ECO_NAME %in% ref_year$ECO_NAME ~ ref_year$refLst[match(ECO_NAME, ref_year$ECO_NAME)]
    )) %>%
    mutate(
      tsf = year - maxYr,
      pre_post_Lst = refLst - get(paste0("mean_", year)),
      post_pre_Lst =  get(paste0("mean_", year)) - refLst
    ) %>%
    mutate(
      TimeSinceFire = case_when(
        tsf == 1 ~ "1",
        tsf %in% 2:5 ~ "2-5",
        tsf %in% 6:10 ~ "6-10",
        tsf %in% 11:20 ~ "11-20",
        tsf %in% 21:30 ~ "21-30",
        tsf %in% 31:36 ~ "31-36",
        TRUE ~ "Other"
      ),
      tsfStd = case_when(
        tsf %in% 1:5 ~ "1-5",
        tsf %in% 6:10 ~ "6-10",
        tsf %in% 11:20 ~ "11-20",
        tsf %in% 21:36 ~ "21+",
        TRUE ~ "Other"
      ),
      tsfAlt = case_when(
        tsf == 1 ~ "1",
        tsf %in% 2:5 ~ "2-5",
        tsf %in% 6:15 ~ "6-15",
        tsf %in% 16:36 ~ "16+",
        TRUE ~ "Other"
      )
    ) %>%
    mutate(across(
      c(TimeSinceFire),
      ~ ordered(., levels = c("1", "2-5", "6-10", "11-20", "21-30", "31-36"))
    )) %>%
     mutate(across(
      c(tsfStd),
      ~ ordered(., levels = c("1-5", "6-10", "11-20", "21+"))
    )) %>%
     mutate(across(
      c(tsfAlt),
      ~ ordered(., levels = c("1", "2-5", "6-15", "16+"))
    )) %>%
    mutate(bTSF = paste0(burned, "_", TimeSinceFire),
           bTsfStd = paste0(burned, "_", tsfStd),
           bTsfAlt = paste0(burned, "_", tsfAlt),
           sampleYr = year) %>%
    rename(lstMean = !!sym(paste0("mean_", year)))
}

# Generate standard pre-year data for multiple years
standardPre_data <- setNames(lapply(years, calculate_standardPre), paste0("standardPre", years))

```

```{r}
standardPre_data
```

## Creat final DF
```{r}
final_df <- bind_rows(
  standardPre_data$standardPre2022, standardPre_data$standardPre2021, standardPre_data$standardPre2020, standardPre_data$standardPre2019, 
  standardPre_data$standardPre2018, standardPre_data$standardPre2017, standardPre_data$standardPre2016, standardPre_data$standardPre2015, 
  standardPre_data$standardPre2014
)

```

```{r}
final_df
```

```{r}
( recent_2014_2022 = final_df %>% filter(tsf <=5) )
( old_2014_2022 = final_df %>% filter(tsf >=5) )
```

# Absolute

## Model Recent time since fire effects

```{r}
fit1 = glm(lstMean ~ bTSF, data = recent_2014_2022, family = gaussian)
```


### Residuals: 

```{r}
summary(fit1)

recent_2014_2022$fit1.res = resid(fit1, type = "pearson")
recent_2014_2022$fit1.fit = fitted(fit1)

sum(residuals(fit1, type="pearson")^2)/fit1$df.res

qplot(fitted(fit1), resid(fit1)) + theme_bw()

```

### Calculate Marginal means for difference test

```{r}
emmfit1 = emmeans(fit1, specs = ~ bTSF)
emmfit1

dfemmfit1 = as.data.frame(emmfit1)

dfemmfit1.1 = dfemmfit1 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTSF, levels = c("Unburned_0", 'One_1', 'One_2-5',  'Two_1', 'Two_2-5', 'Three_1','Three_2-5')))

dfemmfit1.1
```

```{r}
One_1 = c(1, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0)
Three_1 = c(0, 0, 1, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 1, 0, 0)
Two_1 = c(0, 0, 0, 0, 1, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1)

```


```{r}
contrast(emmfit1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))

contrast(emmfit1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()

```

### Plots 
```{r}
ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip()
```

```{r}
allDataFit1Plot = ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#DB4743FF','#F5AF4DFF',  '#DB4743FF','#F5AF4DFF', '#DB4743FF','#F5AF4DFF'), labels=c( "one year", "2-5 years")) +
  labs(x="", y="Land surface temperature (Kelvin)") +
  scale_x_discrete(labels= c( "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

allDataFit1Plot
```

```{r}
ggsave("../figures/20250313_RecentAbsoluteLST.png", plot = allDataFit1Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```


```{r}
ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#DB4743FF','#F5AF4DFF',  '#DB4743FF','#F5AF4DFF', '#DB4743FF','#F5AF4DFF'), labels=c("one year", "2-5 years")) +
  labs(x="", y="Land surface temperature (Kelvin)") +
  scale_x_discrete(labels= c("One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() +
  annotate("text", x = 2, y = 288.3, label = "*", color='#DB4743FF',  size = 10) +
  annotate("text", x = 4, y = 288.9, label = "*", color='#DB4743FF',  size = 10) +
  annotate("text", x = 6, y = 289.2, label = "*", color='#DB4743FF',  size = 10)
  #theme(legend.position="none")

```


```{r}
(recentLSTPlot = ggplot(data = dfemmfit1.1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c( "One year", "2-5 years")) +
  labs(x="Times burned", y="Land surface Temperature (Kelvin)") +
  scale_x_discrete(labels= c( "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() +
   theme( legend.position="right", 
        legend.text=element_text(size=14), 
        legend.title=element_text(size=16),
        legend.key = element_blank()) +
    theme(axis.title.y = element_text(size = 16, hjust = 0.5, vjust = 1.1),
          axis.title.x = element_text(size = 16, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black")) 
)
```

```{r}
ggsave("../figures/20250313_AbsoluteRecentLstPptSlide.png", plot = recentLSTPlot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```

## Model Fit 2: Absolute All time steps since fire

```{r}
fit2 = glm(lstMean ~ bTsfAlt, data = final_df, family = gaussian)
```


### Residuals: 

```{r}
summary(fit2)

final_df$fit2.res = resid(fit2, type = "pearson")
final_df$fit2.fit = fitted(fit2)

sum(residuals(fit2, type="pearson")^2)/fit2$df.res

qplot(fitted(fit2), resid(fit2)) + theme_bw()

```

### Calculate marginal means for difference test

```{r}
emmfit2 = emmeans(fit2, specs = ~ bTsfAlt)
emmfit2

dfemmfit2 = as.data.frame(emmfit2)

dfemmfit2.1 = dfemmfit2 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTsfAlt, levels = c("One_1", "One_2-5", "One_6-15", "One_16+", "Two_1", "Two_2-5", "Two_6-15", "Two_16+", "Three_1", "Three_2-5", "Three_6-15", "Three_16+"))
)
dfemmfit2.1
```

```{r}
One_1 =  c(1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_6_15 =c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_16_36 =c(0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0)

Two_1 =  c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
Two_6_15 =c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
Two_16_36 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)

Three_1 = c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
Three_6_15 =  c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0) 
Three_16_36 = c(0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
```

```{r}
contrast(emmfit2, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36))

contrast(emmfit2, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36),
         adjust = "mvt") %>%
     confint()


```



### Plots



```{r}
( modelfit2_Plot = ggplot(data = dfemmfit2.1 ) + 
  geom_point(aes(x=bTsfAlt, y=mean, color = bTsfAlt)) +
  geom_errorbar(aes(x=bTsfAlt, ymin = LCL, ymax = UCL, color = bTsfAlt)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years")) +
  labs(x="", y="Land surface temperature (Kelvin)") +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

)


```

```{r}
ggsave("../figures/20250313_AbsoluteLSTDifferenceBorealAltTsfBins.png", plot = modelfit2_Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```



# Relative (post-pre) Models
## Model Recent time since fire effects
 Match AGU and ABOVE Analaysis format
```{r}
model1 = glm(post_pre_Lst ~ bTSF, data = recent_2014_2022, family = gaussian)
```

### Residuals: 

```{r}
recent_2014_2022$model1.res = resid(model1, type = "pearson")
recent_2014_2022$model1.fit = fitted(model1)

sum(residuals(model1, type="pearson")^2)/model1$df.res

qplot(fitted(model1), resid(model1)) + theme_bw()

```

### Calculate marginal means for difference

```{r}
emm_model1 = emmeans(model1, specs = ~ bTSF)
emm_model1

df_emm_model1 = as.data.frame(emm_model1)

df_emm_model1 = df_emm_model1 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTSF = factor(bTSF, levels = c( "One_1", "One_2-5",  "Two_1",  "Two_2-5",  "Three_1", "Three_2-5"))
                  )

df_emm_model1
```

```{r}
One_1 = c(1, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0)
Three_1 = c(0, 0, 1, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 1, 0, 0)
Two_1 = c(0, 0, 0, 0, 1, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1)

contrast(emm_model1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))

contrast(emm_model1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()
```


### Plots

```{r}
(recent_model1_plot1 = ggplot(data = df_emm_model1) + 
  geom_point(aes(x=bTSF, y=mean, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c("One year", "2-5 years")) +
  labs(x="Times burned", y="Land surface Temperature Difference (Kelvin)", subtitle = "Difference = post LST - pre LST") +
  scale_x_discrete(labels= c("One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() +
   theme( legend.position="right", 
        legend.text=element_text(size=14), 
        legend.title=element_text(size=16),
        legend.key = element_blank()) +
    theme(axis.title.y = element_text(size = 16, hjust = 0.5, vjust = 1.1),
          axis.title.x = element_text(size = 16, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 14, color = "black")) 
)
```

```{r}
ggsave("../figures/20250313_LSTDifferencerecentRecentBorealPPTSlide.png", plot = recent_model1_plot1, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```


## Model 4 with Alternate Time Since Fire groups
Full time for space substitution.

```{r}
model4 = glm(post_pre_Lst ~ bTsfAlt, data = final_df, family = gaussian)
```

### Residuals: 

```{r}
final_df$model4.res = resid(model4, type = "pearson")
final_df$model4.fit = fitted(model4)

sum(residuals(model4, type="pearson")^2)/model4$df.res

qplot(fitted(model4), resid(model4)) + theme_bw()

summary(model4)
```

### Calculate marginal means for difference test

```{r}
emm_model4 = emmeans(model4, specs = ~ bTsfAlt)
emm_model4

df_emm_model4 = as.data.frame(emm_model4)

df_emm_model4 = df_emm_model4 %>% mutate(mean=(emmean),
                  SE=(SE),
                  LCL=(lower.CL),
                  UCL=(upper.CL),
                  bTsfStd = factor(bTsfAlt, levels = c("One_1", "One_2-5", "One_6-15", "One_16+", "Two_1", "Two_2-5", "Two_6-15", "Two_16+", "Three_1", "Three_2-5", "Three_6-15", "Three_16+"))
                  )

df_emm_model4
```

```{r}
One_1 =  c(1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_6_15 =c(0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0)
One_16_36 =c(0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0)

Two_1 =  c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0)
Two_6_15 =c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0)
Two_16_36 = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1)

Three_1 = c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0)
Three_6_15 =  c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0) 
Three_16_36 = c(0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0)
```

```{r}
contrast(emm_model4, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36))

compTB = contrast(emm_model4, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5,
                             "One_6_15 - Two_6_15" = One_6_15 - Two_6_15,
                             "One_6_15 - Three_6_15" = One_6_15 - Three_6_15,
                             "One_16_36 - Two_16_36" = One_16_36 - Two_16_36,
                             "One__16_36 - Three_16_36" = One_16_36 - Three_16_36),
         adjust = "mvt") %>%
     confint()

compTB
```

### Plots

```{r}
( model4_Plot = ggplot(data = df_emm_model4 ) + 
  geom_point(aes(x=bTsfAlt, y=mean, color = bTsfAlt)) +
  geom_errorbar(aes(x=bTsfAlt, ymin = LCL, ymax = UCL, color = bTsfAlt)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca','#f79256','#fbd1a2',  '#7dcfb6','#00b2ca'), labels=c("1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years", "1 year", "2-5 years", "6-15 years", "16+ years")) +
  labs(x="", y="Land surface temperature differnece (Kelvin)", subtitle = "difference Lst = post LST -pre LST") +
  scale_x_discrete(labels= c("One", "One", "One", "One", "Two", "Two", "Two", "Two", "Three", "Three","Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

)


```

```{r}
ggsave("../figures/20250313_RelativeLSTDifferenceBorealAltTsfBins.png", plot = model4_Plot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```











