---
title: "OrganizeSHPData"
author: "Anna Talucci"
date: "2023-07-20"
output: html_document
---

# Overview
Organize shapefile data that feeds into other scripts

Combine Canadian fire perimetes

join Ecoregions to perimeters
simplify perimeters
# Packages
```{r}
library(sf)
library(tidyverse)
library(magrittr)
library(pacman)
```

# Data 
## AK Fires
```{r}
AKFires = st_read("../data/boundaries/firePerimeters/mtbs_perimeter_data/mtbs_perims_DD.shp", "mtbs_perims_DD")

```

## Canada Fire perimetes
```{r}
CanFires = st_read("../data/boundaries/firePerimeters/nbac_1986_to_2020/nbac_1986_to_2020_20210810.shp", "nbac_1986_to_2020_20210810")
CanFires2021 = st_read("../data/boundaries/firePerimeters/nbac_2021/nbac_2021_r9_20220624.shp", "nbac_2021_r9_20220624")
```

## Canada Proviences
```{r}
prov = st_read("../data/boundaries/lpr_000b16a_e/lpr_000b16a_e.shp", "lpr_000b16a_e")
```
## Political Bounds
```{r}
political = st_read("../data/boundaries/NorthAmerica/boundaries_p_2021_v3.shp", "boundaries_p_2021_v3")
```

## Ecoregions
```{r}
ecor = st_read("../data/boundaries/ecozones/ecoLevel2/NA_CEC_Eco_Level2.shp", "NA_CEC_Eco_Level2")
```

# Select 
##Poloitical Polygons of interest

```{r}
( polBor = political %>% filter(NAME_En %in% c("Alberta", "Alaska", "British Columbia", "Yukoon", "Northwest Territories", "Saskatchewan", "Manitoba", "New Brunswick", "Newfoundland and Labrador", "Nova Scotia", "Nunavut", "Ontario", "Prince Edward Island", "Quebec")) )
```

```{r eval=FALSE, include=FALSE}
st_write(polBor, "../outputs/StudyAreaPoliticalBounds.shp", driver="ESRI Shapefile")
```

## Eco regions of interest
```{r}
ecostudy = ecor %>% filter(NA_L2NAME %in% c("TAIGA CORDILLERA", "TAIGA PLAIN", "TAIGA SHIELD", "ALASKA BOREAL INTERIOR", "BOREAL PLAIN", "BOREAL CORDILLERA", "SOFTWOOD SHIELD"))
```

```{r eval=FALSE, include=FALSE}
st_write(ecostudy, "../outputs/StudyAreaEco.shp", driver="ESRI Shapefile")
```


# Projection
```{r}
projEA = st_crs(ecor)
```

# Ak 
## Find Munson creek fire
```{r}
munson = AKFires %>% filter(BurnBndLat>50) %>% 
  dplyr::mutate(year = lubridate::year(Ig_Date), 
                month = lubridate::month(Ig_Date), 
                day = lubridate::day(Ig_Date)) %>%
  filter(year==2021) %>%
  filter(Incid_Name =="MUNSON CREEK")
```

```{r}
unique(munson$Incid_Name)
```
```{r eval=FALSE, include=FALSE}
st_write(munson, "../outputs/MunsonCreek2021.shp", driver="ESRI Shapefile")
```
## Select Case study fires

var fireIDList = ["AK6586915468120150622", 'AK6594715517520180604']
var fireAk = ee.FeatureCollection('users/FiSL-Reburn/AK4326Fires1986-2021').filter(ee.Filter.inList("Year",  yrList))
var reburnEx = fireAk.filter(ee.Filter.inList('FireID', fireIDList ))

```{r}
akCaseStudyID = c("AK6586915468120150622", 'AK6594715517520180604')
manleyHS = c('AK6481415114120180607', 'AK6490715062520210615')
```
```{r}
AKFires
max(AKFires$Ig_Date)
```

```{r}
( akCaseStudy = AKFires %>% filter(BurnBndLat>50) %>% 
  dplyr::mutate(year = lubridate::year(Ig_Date), 
                month = lubridate::month(Ig_Date), 
                day = lubridate::day(Ig_Date)) %>%
  filter(Event_ID %in% akCaseStudyID) 
  )

```

```{r}
AKFires %>% filter(BurnBndLat>50) %>% 
  dplyr::mutate(year = lubridate::year(Ig_Date), 
                month = lubridate::month(Ig_Date), 
                day = lubridate::day(Ig_Date)) %>%
  filter(Event_ID=='AK6490715062520210615') 
```

```{r}
( manleyHSCaseStudy = AKFires %>% filter(BurnBndLat>50) %>% 
  dplyr::mutate(year = lubridate::year(Ig_Date), 
                month = lubridate::month(Ig_Date), 
                day = lubridate::day(Ig_Date)) %>%
  filter(Event_ID %in% manleyHS) 
  )

```

```{r include=FALSE}
st_write(akCaseStudy, "../outputs/boundaries/firePerimeters/AkCaseStudyPerimeters.shp", driver="ESRI Shapefile")
```

```{r include=FALSE}
st_write(akCaseStudy, "../outputs/boundaries/firePerimeters/AkCaseStudyPerimeters.gpkg")
```

```{r include=FALSE}
st_write(manleyHSCaseStudy, "../outputs/boundaries/firePerimeters/ManleyHSPerimeters.gpkg")
```

# Combine Canadian Fires
##Select columns and bind all fires canadian
```{r}
( CanFires1 = CanFires %>% dplyr::select(-BASRC_1, -new_flg) )
```

```{r}
( allFires = bind_rows(CanFires1, CanFires2021) %>% dplyr::select(-COMMENTS) )
```

```{r include=FALSE}
st_write(allFires, "../outputs/firePerimeters/Canada_1986_2021.shp", driver="ESRI Shapefile")
```

# Reproject
```{r}
prov = st_transform(prov, projEA)
firesCA = st_transform(allFires, projEA)
firesAK = st_transform(AKFires, projEA)
```

```{r}
eco4326 = st_transform(ecor, 4326)
```
```{r include=FALSE}
st_write(eco4326, "../outputs/NAEcoregionsII4326.shp", driver="ESRI Shapefile")
```

# Join provinces to fires


```{r}
(
  fire = st_join(firesCA, prov) %>% 
    st_join(., ecor)
   )
```
```{r}
( firev1 = fire %>% dplyr::select(YEAR:PRUID, PRENAME, NA_L2CODE, NA_L2NAME, NA_L1CODE, NA_L1NAME, geometry))
```




# Write to shp
```{r eval=FALSE, include=FALSE}
st_write(firev1, "../outputs/CanadaFires1986-2021.shp", driver="ESRI Shapefile")
```
```{r}
firev1 = st_read("../outputs/CanadaFires1986-2021.shp", "CanadaFires1986-2021")
```
# Simplify
```{r}
mv_simpl <- st_simplify(firev1, preserveTopology = TRUE,
                        dTolerance = 500)
```


```{r}
simpFire4326 = st_transform(mv_simpl, crs=4326)

```

```{r eval=FALSE, include=FALSE}
st_write(simpFire4326, "../outputs/CanadaFires4326Simp1986-2021.shp", driver="ESRI Shapefile")
```
# Add Alaska

# Simplify polygons
Need to reduce some of the vertices of the polygons
# Filter for SK
```{r}
(skFires = firev1 %>% filter(PRENAME=="Saskatchewan"))
```

```{r eval=FALSE, include=FALSE}
st_write(skFires, "../outputs/SKFires1986-2021.shp", driver="ESRI Shapefile")
```

#  Alaska Fires

```{r}
( akFIRES = AKFires %>% filter(BurnBndLat > 50) %>%
  dplyr::mutate(fireYr = lubridate::year(Ig_Date), 
                month = lubridate::month(Ig_Date), 
                day = lubridate::day(Ig_Date)) %>%
    filter(fireYr>= 1986) )
```


```{r eval=FALSE, include=FALSE}
st_write(akFIRES, "../outputs/Ak_1986_2021.shp", driver="ESRI Shapefile")
```

```{r}
(
  AKFeco = st_join(akFIRES, prov) %>% 
    st_join(., ecor, largest = TRUE)
   )
```
## Simplify AK
```{r eval=FALSE, include=FALSE}
ak_simpl <- st_simplify(akFires, preserveTopology = TRUE,
                        dTolerance = 500)
```
Error in wk_handle.wk_wkb(wkb, s2_geography_writer(oriented = oriented, :
Loop 0 is not valid: Edge 12580 has duplicate vertex with edge 12653


## Transform AK
```{r}
simpAK4326 = st_transform(AKFeco, crs=4326)

```

# Combine AK and Canada perimeters
```{r}
names(simpFire4326)
```

```{r}
( CA = simpFire4326 %>% 
  dplyr::select("YEAR", "NFIREID", "PRUID", "PRENAME", "NA_L2NAME", "geometry") %>%
  rename(Year = YEAR, FireID = NFIREID) %>%
    mutate(FireID = as.character(FireID))
)
```
```{r}
str(CA)
```

```{r}
names(simpAK4326)
```

```{r}
( AK = simpAK4326 %>% 
  dplyr::select("fireYr", "Event_ID", "PRUID", "PRENAME", "NA_L2NAME", "geometry") %>%
  rename(Year = fireYr, FireID = Event_ID) 
)
```
```{r eval=FALSE, include=FALSE}
st_write(AK, "../outputs/AK4326Fires1986-2021.shp", driver="ESRI Shapefile")
```
```{r}
str(AK)
```

```{r}
( borealFires = bind_rows(CA, AK) )
```

```{r eval=FALSE, include=FALSE}
st_write(borealFires, "../outputs/Boreal4326Fires1986-2021.shp", driver="ESRI Shapefile")
```