---
title: "OrganizeSHPData"
author: "Anna Talucci"
date: "2023-07-20"
output: html_document
---

# Overview
Organize shapefile data that feeds into other scripts

Combine Canadian fire perimetes

join Ecoregions to perimeters
simplify perimeters
# Packages
```{r}
library(sf)
library(tidyverse)
library(magrittr)
library(pacman)
```

# Data 
## AK Fires
```{r}
AKFires = st_read("../data/mtbs_perims_DD.shp", "mtbs_perims_DD")

```

## Canada Fire perimetes
```{r}
CanFires = st_read("../data/nbac_1986_to_2020/nbac_1986_to_2020_20210810.shp", "nbac_1986_to_2020_20210810")
CanFires2021 = st_read("../data/nbac_2021/nbac_2021_r9_20220624.shp", "nbac_2021_r9_20220624")
```

## Canada Proviences
```{r}
prov = st_read("../data/lpr_000b16a_e/lpr_000b16a_e.shp", "lpr_000b16a_e")
```

## Ecoregions
```{r}
ecor = st_read("../data/ecoLevel2/NA_CEC_Eco_Level2.shp", "NA_CEC_Eco_Level2")
```


# Projection
```{r}
projEA = st_crs(ecor)
```

# Ak 
## Find Munson creek fire
```{r}
munson = AKFires %>% filter(BurnBndLat>50) %>% 
  dplyr::mutate(year = lubridate::year(Ig_Date), 
                month = lubridate::month(Ig_Date), 
                day = lubridate::day(Ig_Date)) %>%
  filter(year==2021) %>%
  filter(Incid_Name =="MUNSON CREEK")
```

```{r}
unique(munson$Incid_Name)
```
```{r }
st_write(munson, "../outputs/MunsonCreek2021.shp", driver="ESRI Shapefile")
```

# Combine Canadian Fires
##Select columns and bind all fires canadian
```{r}
( CanFires1 = CanFires %>% dplyr::select(-BASRC_1, -new_flg) )
```

```{r}
( allFires = bind_rows(CanFires1, CanFires2021) %>% dplyr::select(-COMMENTS) )
```


# Reproject
```{r}
prov = st_transform(prov, projEA)
firesCA = st_transform(allFires, projEA)
```

# Join provinces to fires


```{r}
(
  fire = st_join(firesCA, prov) %>% 
    st_join(., ecor)
   )
```
```{r}
( firev1 = fire %>% dplyr::select(YEAR:PRUID, PRENAME, NA_L2CODE, NA_L2NAME, NA_L1CODE, NA_L1NAME, geometry))
```




# Write to shp
```{r }
st_write(firev1, "../outputs/CanadaFires1986-2021.shp", driver="ESRI Shapefile")
```
```{r}
firev1 = st_read("../outputs/CanadaFires1986-2021.shp", "CanadaFires1986-2021")
```
# Simplify
```{r}
mv_simpl <- st_simplify(firev1, preserveTopology = TRUE,
                        dTolerance = 500)
```


```{r}
simpFire4326 = st_transform(mv_simpl, crs=4326)

```

```{r }
st_write(simpFire4326, "../outputs/CanadaFires4326Simp1986-2021.shp", driver="ESRI Shapefile")
```
# Add Alaska

# Simplify polygons
Need to reduce some of the vertices of the polygons
# Filter for SK
```{r}
(skFires = firev1 %>% filter(PRENAME=="Saskatchewan"))
```

```{r }
st_write(skFires, "../outputs/SKFires1986-2021.shp", driver="ESRI Shapefile")
```
##  Alaska Fires

```{r}
( akFires = UsaFires %>% filter(BurnBndLat > 50) %>%
  dplyr::mutate(fireYr = lubridate::year(Ig_Date), 
                month = lubridate::month(Ig_Date), 
                day = lubridate::day(Ig_Date)) %>%
    filter(fireYr>= 1986) )
```


```{r eval=FALSE, include=FALSE}
st_write(akFires, "../outputs/AkFires1986-2021.shp", driver="ESRI Shapefile")
```