---
title: "Organize Fire perimeters with too many vertices"
author: "Anna Talucci"
date: "2025-02-27"
output: html_document
---

# Overview 

In Earth engine the following fire years will not upload because there is a feature that has too many vertices

1995 feature '77' has 1150448 vertices
2016 feature '93' has 1345805 vertices, 
2019 feature '32' has 1106788 vertices,


# Packages
```{r}
library(sf)
library(lwgeom)
```

# Data
```{r}
can1995 = st_read("../outputs/firesByYearCAN/can1995.shp", "can1995")
```
```{r}
can2016 = st_read("../outputs/firesByYearCAN/can2016.shp", "can2016")
```

```{r}
can2019 = st_read("../outputs/firesByYearCAN/can2019.shp", "can2019")
```

```{r}
can1995
```



```{r}
can1995_cast <- st_cast(can1995, "POLYGON")
```
```{r}
can2016_cast <- st_cast(can2016, "POLYGON")
```

```{r}
can2019_cast <- st_cast(can2019, "POLYGON")
```
 feature '2118' has 1087929 vertices
```{r}
( f2118 = can2019_cast %>% slice(2118) )
( f2118_nfireid = can2019_cast %>% filter(NFIREID==274) )
```
```{r}
( f274 = can2019 %>% filter(NFIREID==274) )
```
```{r}
plot(st_geometry(f2118))
plot(st_geometry(f274))

```

# Get polygon centroid and bounding box
```{r}
library(sf)
library(lwgeom)

centroid <- st_coordinates(st_centroid(f2118))
bbox <- st_bbox(f2118)

```

# Create a vertical line through the centroid to split the polygon
```{r}
split_line <- st_sfc(st_linestring(matrix(c(bbox["xmin"], centroid[2],  # Start point (left)
                                            bbox["xmax"], centroid[2]), # End point (right)
                                          ncol=2, byrow=TRUE)), 
                     crs = st_crs(f2118))
```

# Convert line to sf object
```{r}
# Convert to simple feature geometry (sfg) and then to sfc
split_line_sfc <- st_sfc(split_line, crs = st_crs(polygon))

# Split the polygon using st_split() (NOTE: lwgeom must be installed)
split_result <- st_split(polygon, split_line_sfc)

# Extract polygons from the result
split_polygons <- st_collection_extract(split_result, "POLYGON")
```



# Plot result
```{r}
plot(st_geometry(f2118), col = 'lightblue', border = 'black')
plot(st_geometry(split_line_sf), col = 'red', add = TRUE, lwd = 2)
#plot(st_geometry(split_polygons), col = c('blue', 'green'), add = TRUE, border = 'black')
```



```{r}
can1995_cast
```

```{r}
( Can2019_Not274 = can2019_cast %>% filter(NFIREID!=274) )
```

```{r}
st_write(can1995_cast , "../outputs/firesByYearCAN/castAll_can1995.shp", driver = "ESRI Shapefile", append = FALSE)
```

```{r}
st_write(can2016_cast , "../outputs/firesByYearCAN/castAll_can2016.shp", driver = "ESRI Shapefile", append = FALSE)
```

```{r}
st_write(can2019_cast , "../outputs/firesByYearCAN/castAll_can2019.shp", driver = "ESRI Shapefile", append = FALSE)
```

```{r}
st_write(Can2019_Not274 , "../outputs/firesByYearCAN/castAllNot274_can2019.shp", driver = "ESRI Shapefile", append = FALSE)
```
