---
title: "Organize Multi Index Data"
author: "Anna Talucci"
date: "2025-03-31"
output: html_document
---

# Packages

```{r}
library(tidyverse)

```



# Data

```{r}
lst = read_csv("../outputs/cleanedSamplePoints/LstBoreal2013-2022.csv")
```

```{r}
csvList = list.files("../data/multiIndex/multiIndexSampled", pattern=".csv$", full.names=TRUE)
```

```{r}

sampledData = csvList %>% 
                map_df(~ read_csv(.x) %>% 
                mutate(filename = tools::file_path_sans_ext(basename(.x)))) %>%
                separate(filename, into = c("code", "file", "sample", "dataYear"), sep = "_", convert = TRUE, extra = "merge")
```



# View

```{r}
sampledData
```

```{r}
names(sampledData)
```


# Organize

```{r}
( burnedData = sampledData %>% filter(maxYr !=-999999) )

( unburnedData = sampledData %>% filter(maxYr ==-999999) )
```
## Unburned sampled

```{r}
unburnedData
```

```{r}
set.seed(23)

df_unburned <-  unburnedData %>%
  filter(dataYear !=2023) %>%
    dplyr::select(-`system:index`, -.geo) %>%
    relocate(ID, row_id) %>%
  group_by(ID, row_id) %>%
  slice_sample(n=1) %>%  
  ungroup() %>%
  filter(blue!=-9999.0) %>%
  filter(nbrt1!=-9999.0) %>%
  slice_sample(prop = .05) %>%
  mutate(tx="unburned") %>%
    rename_with(~ paste0(.,"_post"), -row_id) %>%
    mutate(preYear = dataYear_post - 1) 

df_unburned

```
```{r}
( unburned_delta_post = unburnedData %>%
  filter(dataYear !=2023) %>%
    dplyr::select(-`system:index`, -.geo) %>%
    relocate(ID, row_id) %>%
  rename(preYear = dataYear) %>%
  rename_with(~ paste0(.,"_pre"), -row_id) %>%
  rename(preYear = preYear_pre) %>%
  full_join(., df_unburned, by= c("row_id", "preYear")) %>%
  drop_na(tx_post) %>%
  select(row_id:blue_pre, csi_pre:green_pre, mirbi_pre:nir_pre, red_pre:vi6t_pre, ba_post:blue_post, csi_post:green_post, mirbi_post:nir_post, red_post:vi6t_post) %>%
  mutate(across(ends_with("_pre"), .names = "{col}_diff") - across(ends_with("_post"))) %>%
 rename_with(~ sub("_\\d+", "", .), ends_with("_diff")) %>%
  select(row_id, contains("_post"), contains("_diff")) %>%
   rename_with(~ gsub("_post$", "", .x)) %>%
   rename_with(~ gsub("_pre", "", .x)) %>%
   mutate(tx = "unburned")
)
```
rename(dataYear = dataYear_post, maxYr = maxYr_post, countYr = countYr_post) %>%
  
  mutate(across(ends_with("_pre"), .names = "{col}_diff") - across(ends_with("_post"))) %>%
 rename_with(~ sub("_\\d+", "", .), ends_with("_diff")) %>%
  select(row_id, contains("_post"), contains("_diff")) %>%
   rename_with(~ gsub("_post$", "", .x))
## Burned postfire

```{r}
( df_burned = burnedData %>% 
    mutate(postYear = maxYr +1) %>%
    dplyr::select(row_id, dataYear, maxYr, postYear, countYr, ba:blue, csi:green, mirbi:nir, red, savi:vi6t) %>%
    rowwise() %>%
  mutate(match =  (dataYear %in% postYear)) %>%
  filter(match) %>% # Keep only rows where any match occurs
  mutate(match_type = case_when(
    dataYear %in% postYear ~ "post")) %>%
    filter(blue!=-9999.0	) %>%
    filter(nbrt1 !=-9999.0) %>%
  mutate(tx="burned") %>%
    rename_with(~ paste0(.,"_post"), -row_id)
)
```

```{r}

( df_prefire = burnedData %>% 
    mutate(preYear = maxYr - 1) %>%
    dplyr::select(row_id, dataYear, maxYr, preYear, countYr, ba:blue, csi:green, mirbi:nir, red, savi:vi6t) %>%
    rowwise() %>%
  mutate(match =  (dataYear %in% preYear)) %>%
  filter(match) %>% # Keep only rows where any match occurs
  mutate(match_type = case_when(
    dataYear %in% preYear ~ "pre")) %>%
    filter(blue!=-9999.0	) %>%
    filter(nbrt1 !=-9999.0) %>%
  mutate(tx="prefire") %>%
    rename_with(~ paste0(.,"_pre"), -row_id)
)

```

```{r}
unique(df_burned$maxYr)
unique(unburnedData$dataYear)
n_distinct(unburnedData$row_id)
n_distinct(unburnedData$ID)
```

# Calculate delta varaibles
prefire - postfire
```{r}
( delta_post_variables = df_burned %>% 
  full_join(., df_prefire, by=c("row_id")) %>%
  rename(dataYear = dataYear_post, maxYr = maxYr_post, countYr = countYr_post) %>%
  select(row_id:maxYr, countYr:vi6t_post, ba_pre:vi6t_pre) %>%
  mutate(across(ends_with("_pre"), .names = "{col}_diff") - across(ends_with("_post"))) %>%
 rename_with(~ sub("_\\d+", "", .), ends_with("_diff")) %>%
  select(row_id, contains("_post"), contains("_diff")) %>%
   rename_with(~ gsub("_post$", "", .x)) %>%
   rename_with(~ gsub("_pre", "", .x)) %>%
   mutate(tx = "burned")
)
 
```


 



# Combine burned and unburned  into df

```{r}
( combine = bind_rows(df_burned, df_unburned) %>%
    dplyr::select(row_id, tx, dataYear, maxYr, countYr, ba:vi6t))
```


```{r}
( delta_post_combine = bind_rows(delta_post_variables, unburned_delta_post) )
```
# Write to csv

```{r}
write.csv( combine, '../outputs/cleanedSamplePoints/2025_05_16sampledMultiIndexAK.csv', row.names=F)
```

```{r}
write.csv(delta_post_combine, '../outputs/cleanedSamplePoints/2025-05-17sampledMultiIndexAK_delta_post.csv', row.names=F)
```