---
title: "Reburn Field point dNBR"
author: "Anna Talucci"
date: "2024-03-19"
output: html_document
---

# Overview

Evaluate the dNBR values form Landsat for field sites that experienced reburn


# Packages
```{r}
library(tidyverse)
library(sf)
```

# Data

```{r}
nbr = read.csv("../data/eeSampledData/field/fislLcNBRsample.csv")
```

```{r}
field = read_csv("../outputs/field/combinedFieldData.csv")
```

```{r}
lcsites = read_csv("../data/fieldCsv/LC_SITES_11142023.csv")
lcdata = read_csv("../data/fieldCsv/LC_Data_09052023.csv")
fd = read_csv("../data/fieldCsv/FiSL_sitedata_BB2023.csv")
fp = read_csv("../data/fieldCsv/FINAL_PLOTS_01142024.csv")
```

# Dataframe
```{r}
field
```
# Field organize and clean to point for ee

```{r}
( fieldPlots = field %>% 
  group_by(id, fire_scar, site, lat, lon) %>%
  summarize(n= n(),
            recent_burn = toString(unique(recent_burn)),
            second_burn = toString(unique(second_burn)),
            triple_burn = toString(unique(triple_burn)),
            reburn_age = toString(unique(reburn_age)),
            triple_age = toString(unique(triple_age))) %>%
    mutate(recent_burn = as.numeric(recent_burn),
           second_burn = as.numeric(second_burn),
           triple_burn = as.numeric(triple_burn),
           reburn_age = as.numeric(reburn_age),
           triple_age = as.numeric(triple_age))
)          

```

```{r}
fieldPlots %>% filter(lon>0)
```
```{r}
fieldPlots$lon[fieldPlots$id=="MmBonanza_4"]<- -150.6562
fieldPlots$lon[fieldPlots$id=="SnohomishEagle_61"]<- -152.7834
```

```{r}
( fieldPlotsRecent = fieldPlots %>% 
    filter(reburn_age <=30)
)
```

## Function to make points

```{r}
df_to_sf <- function(x){
  st_as_sf(x, coords = c("lon","lat"), crs = 4326, remove = FALSE)
}
```

```{r}
( fieldPlotsPoints = fieldPlots %>% df_to_sf()
)
```

```{r eval=FALSE, include=FALSE}
st_write(fieldPlotsPoints, "../outputs/spatialDataCreate/shapefiles/field/fieldPlotPointsReburn.shp", driver="ESRI Shapefile")
```

# NBR From ee

```{r}
nbr
```
```{r}
( nbr_clean = nbr %>%
  rename(fire_scar = fir_scr, recent_burn=rcnt_br, reburn_age = rebrn_g, second_burn = scnd_br) %>%
  dplyr::select(-.geo, -system.index)
)
```

```{r}
( join = fieldPlots 
  %>% left_join(., nbr_clean, by = c("id", "lat", "lon", "fire_scar", "n", "recent_burn", "reburn_age", "second_burn", "site")) #%>% dplyr::select(id:file, nbr2003:nbr2023)
)

```

```{r}
(recent = join %>% 
    filter(reburn_age <=30)
  
)
```

```{r}
unique(sort(recent$recent_burn))
```

```{r}
( recent2015 = recent %>%
    filter(recent_burn==2015) %>%
    dplyr::select(id:second_burn, reburn_age, nbr2003, nbr2004, nbr2005, nbr2014, nbr2015, nbr2016) %>%
    mutate(dNBR2015 = nbr2014-nbr2016,
           dNBR2004 = nbr2003-nbr2005,
           dNBRTotalChange = nbr2003-nbr2015
           )
    
)
```


```{r}
( recent2018 = recent %>%
    filter(recent_burn==2018) %>%
    dplyr::select(id:second_burn, reburn_age, nbr2017, nbr2018, nbr2019) %>%
    mutate(dNBR2018 = nbr2017-nbr2019)
    
)
```

```{r}
( recent2019 = recent %>%
    filter(recent_burn==2019) %>%
    dplyr::select(id:second_burn, reburn_age, nbr2018, nbr2019, nbr2020) %>%
    mutate(dNBR2019 = nbr2018-nbr2020)
    
)
```

```{r}
( recent2021 = recent %>%
    filter(recent_burn==2021) 
)
```

```{r}
( recent2022 = recent %>%
    filter(recent_burn==2022) 
)
```

```{r}
unique(sort(recent2022$second_burn))
```

```{r}
( recent2022_2005 = recent %>%
    filter(recent_burn==2022) %>%
    filter(second_burn == 2005) %>%
    dplyr::select(id:second_burn, reburn_age, nbr2004, nbr2005, nbr2006, nbr2021, nbr2022, nbr2023) %>%
    mutate(dNBR2022 = nbr2021-nbr2023,
           dNBR2005 = nbr2004-nbr2006,
           dNBRTotalChange = nbr2004-nbr2023
           ) %>%
  drop_na(dNBRTotalChange)
)
```

```{r}
( recent2022_2007 = recent %>%
    filter(recent_burn==2022) %>%
    filter(second_burn == 2007) %>%
    dplyr::select(id:second_burn, reburn_age, nbr2006, nbr2007, nbr2008, nbr2021, nbr2022, nbr2023) %>%
    mutate(dNBR2022 = nbr2021-nbr2023,
           dNBR2005 = nbr2006-nbr2008,
           dNBRTotalChange = nbr2006-nbr2023
           )  
)
```


```{r}
( recent2022_2011 = recent %>%
    filter(recent_burn==2022) %>%
    filter(second_burn == 2011) %>%
    dplyr::select(id:second_burn, reburn_age, nbr2010, nbr2011, nbr2012, nbr2021, nbr2022, nbr2023) %>%
    mutate(dNBR2022 = nbr2021-nbr2023,
           dNBR2011 = nbr2010-nbr2012,
           dNBRTotalChange = nbr2010-nbr2023
           ) %>%
  drop_na(dNBRTotalChange)
)
```













# OLD CODE


```{r}
( fpSubset = fp %>% 
    dplyr::select(fire_scar, site, latitude, longitude, recent_burn, second_burn, triple_burn, reburn_age, triple_age) %>%
    rename(lat=latitude, lon = longitude))
```

```{r}
unique(fpSubset$site)
```

# Join

```{r}
( nbr1 = nbr %>% 
    mutate(idRow = row_number()) )
```


```{r}
nbr
```

```{r}
( field = fpSubset %>% unite("id", fire_scar, site, sep = "_", remove = FALSE, na.rm = FALSE) )
```






```{r}
join = field %>% left_join(., nbr1, by = c("id", "lat", "lon")) #%>% dplyr::select(id:file, nbr2003:nbr2023)

join
```

```{r}

( reburns = join %>% drop_na(burnYr2) %>%
    mutate(lag1 = (burnYr1-burnYr2), 
           lag2 = (burnYr2 - burnYr3)) %>% 
  filter(lag1 <= 30) %>%
    filter(burnYr2 >= 1984) %>%
    mutate()
  ) 
```
```{r}
unique(reburns$lag1)
unique(reburns$burnYr1)
unique(reburns$burnYr2)
```
# Organize Data

```{r}
(nbr_long = nbr1 %>% dplyr::select(-.geo, -system.index) %>%
  pivot_longer(
    cols = starts_with("nbr"),
    names_to = "year",
    names_prefix = "nbr",
    values_to = "nbr")
)
```
```{r}
unique(nbr_long$site)
unique(nbr_long$siteId)
unique(nbr_long$id)
unique(nbr_long$location)
unique(nbr_long$burnYr1)
```


```{r}
nbr_long %>% filter_all(any_vars(is.na(.))) 
```

```{r}
nbr_long  %>% 
  group_by(location, burnYr1) %>%
  summarise(n_site = n_distinct(site),
            n_siteId = n_distinct(siteId),
            n_id = n_distinct(id),
            n_idRow = n_distinct(idRow))
```

```{r}
nbr_long
```

# Figures

## Trajectories
facet_grid(vars(location), vars(burnYr1),  scales = "free") +
```{r}
nbr_long %>% 
  filter(location == "Alaska") %>% 
  filter(burnYr1 == 2004) %>%
  ggplot(., aes(x=year, y=nbr, color=as.factor(idRow))) +
  geom_point() +
  theme_minimal() +
  labs(x="Years", y="NBR", subtitle="burn year 1 2004")
```











## Boxplot dNBR for field sites split by most recent burn year

```{r}
ggplot(dnbr, aes(x=as.factor(rcnt_br), y=nbr)) +
  geom_boxplot(color="#91bfdb") +
  theme_minimal() +
  labs(x="Year of most recent burn", y="dNBR")
```

## Boxplot dNBR for field sites split by time between burns

```{r}
ggplot(dnbr, aes(x=as.factor(rebrn_g), y=dnbr)) +
  geom_boxplot(color="#91bfdb") +
  theme_minimal() +
  labs(x="Time between fire events", y="dNBR")
```

## Boxplot dNBR for field sites split by time between burns with points

```{r}
ggplot(dnbr, aes(x=as.factor(rebrn_g), y=dnbr)) +
  geom_boxplot(color="#91bfdb") +
  geom_point(aes(position = "dodge"), shape=1, color="#fc8d59", size=2) +
  theme_minimal() +
  labs(x="Time between fire events", y="dNBR")
```