---
title: "Reburn Field point dNBR"
author: "Anna Talucci"
date: "2024-03-19"
output: html_document
---

# Overview

Evaluate the dNBR values form Landsat for field sites that experienced reburn


# Packages
```{r}
library(tidyverse)
library(sf)
```

# Data
```{r}
file_list <- list.files(path = "../data/fieldDnbr/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
file_list
```

```{r}
li <- purrr::map(file_list,
function(file){
df <- read.csv(file)

current_names <- names(df)

# find what names are missing
# do mutate(missing_names = NA)

return(df)
}
)
```

```{r}
dnbr = bind_rows(li) 
```

# Dataframe

```{r}
head(dnbr)
```

## Number of sites with dNBR above or below 100

n=306 greater than 100

n=90 less than 100

```{r eval=FALSE, include=FALSE}
dnbr %>% filter(dnbr >= 100)
```
# Figures

## Boxplot dNBR for field sites split by most recent burn year

```{r}
ggplot(dnbr, aes(x=as.factor(rcnt_br), y=dnbr)) +
  geom_boxplot(color="#91bfdb") +
  theme_minimal() +
  labs(x="Year of most recent burn", y="dNBR")
```

## Boxplot dNBR for field sites split by time between burns

```{r}
ggplot(dnbr, aes(x=as.factor(rebrn_g), y=dnbr)) +
  geom_boxplot(color="#91bfdb") +
  theme_minimal() +
  labs(x="Time between fire events", y="dNBR")
```

## Boxplot dNBR for field sites split by time between burns with points

```{r}
ggplot(dnbr, aes(x=as.factor(rebrn_g), y=dnbr)) +
  geom_boxplot(color="#91bfdb") +
  geom_point(aes(position = "dodge"), shape=1, color="#fc8d59", size=2) +
  theme_minimal() +
  labs(x="Time between fire events", y="dNBR")
```