---
title: "Sampled Points Summary"
author: "Anna Talucci"
date: "2023-11-30"
output: html_document
---

# Overview

# Packages
```{r}
library(tidyverse)
library(sf)
```

# Data

```{r}
all = read_csv("../data/samplePoints/sk/skdNBRLstAlb.csv") %>% dplyr::select(-.geo)
```

```{r eval=FALSE, include=FALSE}
all
```


```{r}
sklstalb = read_csv("../data/samplePoints/sk/skLstAlb.csv") %>% dplyr::select(-.geo)
skreburn = read_csv("../data/samplePoints/sk/skreburn.csv") %>% dplyr::select(-.geo)
```

```{r}
sklst500 = read_csv("../data/samplePoints/raw/sk/skLst500.csv") %>% dplyr::select(-.geo)
skreburn500 = read_csv("../data/samplePoints/raw/sk/skReburn500.csv") %>% dplyr::select(-.geo)
skalb500 = read_csv("../data/samplePoints/raw/sk/skAlb500.csv") %>% dplyr::select(-.geo)
```

```{r}
sklst1k = read_csv("../data/samplePoints/raw/sk/skLst1000v2.csv") %>% dplyr::select(-.geo)
skreburn1k = read_csv("../data/samplePoints/raw/sk/skReburn1000v2.csv") %>% dplyr::select(-.geo)
skalb1k = read_csv("../data/samplePoints/raw/sk/skAlb1000v2.csv") %>% dplyr::select(-.geo)
```





# Organize Sample points
cols = new_sp_m014:newrel_f65,
## Split by Canada and Alaksa an then recombine
```{r eval=FALSE, include=FALSE}
( can = all %>% 
  rename(ID='system:index') %>%
  dplyr::select(ID, ST_L8, albedo, countCA, dnbr1986:dnbr2017, dnbrCt, maxCA, medianCA, minCA, sumCA) %>%
    rename(countYr = countCA, maxYr= maxCA, medianYr = medianCA, minYr = minCA, sumYr = sumCA)
  
)

```

```{r eval=FALSE, include=FALSE}
( ak = all %>% 
  rename(ID='system:index') %>%
  dplyr::select(ID, ST_L8, albedo, countAK, dnbr1986:dnbr2017, dnbrCt, maxAK, medianAK, minAK, sumAK) %>%
    rename(countYr = countAK, maxYr= maxAK, medianYr = medianAK, minYr = minAK, sumYr = sumAK)
  
)

```

```{r eval=FALSE, include=FALSE}

( df = bind_rows(can, ak) %>%
    dplyr::select(ID:countYr, dnbrCt:sumYr, dnbr1986:dnbr2017) %>%
  pivot_longer(cols = dnbr1986:dnbr2017, names_to = "dnbrYr", values_to = "dnbrValue", values_drop_na = TRUE) %>%
  mutate(new = readr::parse_number(as.character(dnbrYr))) %>%
  dplyr::select(-dnbrYr) %>%
  rename(dnbrYr=new) 
    
)

```

mutate(burned = ifelse(countYr==0, "Unburned",
                                     ifelse(countYr==1, "One",
                                            ifelse(countYr==2, "Two",
                                                   ifelse(countYr==3, "Three", "other"))))
Colapse dnbr values and years



```{r eval=FALSE, include=FALSE}
(
dnbrValues = df %>% 
  group_by(ID) %>% 
  summarise(dnbrValue = paste0(unique(dnbrValue), collapse = ",")) 
)
```

```{r eval=FALSE, include=FALSE}
( dnbrYrs = df %>% 
  group_by(ID) %>% 
  summarise(dnbrYr = paste0(unique(dnbrYr), collapse = ",")) %>%
    right_join(., dnbrValues, by='ID')
)
```
```{r eval=FALSE, include=FALSE}
(
  df1 = df %>% 
    pivot_wider(names_from = dnbrYr,names_glue = "yr_{dnbrYr}",values_from = dnbrValue) %>%
    dplyr::select(ID:sumYr) %>%
    right_join(., dnbrYrs, by = 'ID') %>%
    mutate(burned = ifelse(countYr==0, "Unburned",
                                     ifelse(countYr==1, "One",
                                            ifelse(countYr==2, "Two",
                                                   ifelse(countYr==3, "Three", "other"))))) %>%
    mutate(TSF = 2018-maxYr) %>%
    filter(countYr<=3) %>%
    mutate(TimeSinceFire = ifelse(countYr==0, "unburned",
                                     ifelse(TSF==1, "One",
                                            ifelse(TSF %in% c(2, 3, 4, 5), "2-5",
                                                   ifelse(TSF %in% c(6, 7, 8, 9, 10), "6-10", 
                                                          ifelse(TSF %in% c(10:19), "10-19", 
                                                                 ifelse(TSF %in% c(20:31), "20-31", "other")))))))  %>% 
    filter(maxYr <=2017 | sumYr==0) %>%
    drop_na("albedo", "ST_L8")  #%>% 
  #mutate_at("TimeSinceFire", ~replace_na(., "Unburned")) %>%
  #dplyr::mutate(
    #TimeSinceFire = factor(TimeSinceFire, levels = c('Unburned', 'One', '2-5', '6-10', '10-19', '20-31')),
    #burned = factor(burned, labels = c("Unburned", "One", "Two", "Three"))) 
    
)
```
```{r eval=FALSE, include=FALSE}
unique(sort(df1$TSF))
```

```{r eval=FALSE, include=FALSE}
na_rows <- df1[!complete.cases(df1), ]
na_rows
```


# Organize Sample Region Points

```{r}
sklstalb
skreburn
```

```{r}
(
skSamples = sklstalb %>% 
  full_join(., skreburn, by=c("system:index", "ECO_NAME", "lat", "lon", "rowId")) %>%
  rename(ID = "system:index", countYr = count.x, maxYr=maxCA, medianYr=medianCA, minYr=minCA, sumYr=sumCA) %>%
  dplyr::select(ID:rowId, maxYr:sumYr) %>%
  mutate(burned = ifelse(countYr==0, "Unburned",
                  ifelse(countYr==1, "One",
                  ifelse(countYr==2, "Two",
                  ifelse(countYr==3, "Three", "other"))))) %>%
  mutate(TSF = 2018-maxYr) %>%
    filter(countYr<=3) %>%
    mutate(TimeSinceFire = ifelse(countYr==0, "0",
                                     ifelse(TSF==1, "1",
                                            ifelse(TSF %in% c(2, 3, 4, 5), "2-5",
                                                   ifelse(TSF %in% c(6, 7, 8, 9, 10), "6-10", 
                                                          ifelse(TSF %in% c(10:19), "10-19", 
                                                                 ifelse(TSF %in% c(20:31), "20-31", "Other")))))))  %>% 
  mutate(TimeSinceFire= ordered(TimeSinceFire, levels = c("0", "1", "2-5", "6-10", "10-19", "20-31", "Other")),
         burned = ordered(burned, levels = c("Unburned", "One", "Two", "Three", "other"))) %>%
  mutate_at(c('maxYr', 'medianYr', 'minYr', 'sumYr', 'TSF'), ~replace_na(.,0)) %>%

  drop_na("ST_L8") %>% 
  drop_na("albedo") %>%
  filter(maxYr <=2017) # takes n from 590 to 510
)

```
 
```{r}
write.csv(skSamples, '../outputs/cleanedSamplePoints/skStratifiedSample.cvs', row.names=F)
```

# Organize Sample Region Points 500

```{r}
skalb500
skreburn500
sklst500
```

```{r}
(
skSamples500 = skalb500 %>% 
  full_join(., skreburn500, by=c("system:index", "ECO_NAME", "lat", "lon", "rowId")) %>%
  full_join(., sklst500, by=c("system:index", "ECO_NAME", "lat", "lon", "rowId")) %>%
  rename(ID = "system:index", countYr = count.x, maxYr=maxCA, medianYr=medianCA, minYr=minCA, sumYr=sumCA) %>%
  
  mutate_at(c('maxYr', 'medianYr', 'minYr', 'sumYr'), ~replace_na(.,0)) %>%
  dplyr::select(ID:rowId, maxYr:sumYr, ST_L8, water) %>%
  mutate(burned = ifelse(countYr==0, "Unburned",
                  ifelse(countYr==1, "One",
                  ifelse(countYr==2, "Two",
                  ifelse(countYr==3, "Three", "other"))))) %>%
  mutate(TSF2018 = 2018-maxYr,
         TSF2017 = 2017-maxYr,
         TSF2016 = 2016-maxYr,
         TSF2015 = 2015-maxYr,
         TSF2014 = 2014-maxYr,
         TSF2013 = 2013-maxYr) %>%
    mutate(TimeSinceFire2018 = ifelse(countYr==0, "0",
                               ifelse(TSF2018==1, "1",
                               ifelse(TSF2018 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2018 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2018 %in% c(10:19), "10-19", 
                               ifelse(TSF2018 %in% c(20:31), "20-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2017 = ifelse(countYr==0, "0",
                               ifelse(TSF2017==1, "1",
                               ifelse(TSF2017 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2017 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2017 %in% c(10:19), "10-19", 
                               ifelse(TSF2017 %in% c(20:31), "20-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2016 = ifelse(countYr==0, "0",
                               ifelse(TSF2016==1, "1",
                               ifelse(TSF2016 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2016 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2016 %in% c(10:19), "10-19", 
                               ifelse(TSF2016 %in% c(20:31), "20-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2015 = ifelse(countYr==0, "0",
                               ifelse(TSF2015==1, "1",
                               ifelse(TSF2015 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2015 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2015 %in% c(10:19), "10-19", 
                               ifelse(TSF2015 %in% c(20:31), "20-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2014 = ifelse(countYr==0, "0",
                               ifelse(TSF2014==1, "1",
                               ifelse(TSF2014 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2014 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2014 %in% c(10:19), "10-19", 
                               ifelse(TSF2014 %in% c(20:31), "20-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2013 = ifelse(countYr==0, "0",
                               ifelse(TSF2013==1, "1",
                               ifelse(TSF2013 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2013 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2013 %in% c(10:19), "10-19", 
                               ifelse(TSF2013 %in% c(20:31), "20-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2018= ordered(TimeSinceFire2018, levels = c("0", "1", "2-5", "6-10", "10-19", "20-31", "Other")),
         TimeSinceFire2017= ordered(TimeSinceFire2017, levels = c("0", "1", "2-5", "6-10", "10-19", "20-31", "Other")),
         TimeSinceFire2016= ordered(TimeSinceFire2016, levels = c("0", "1", "2-5", "6-10", "10-19", "20-31", "Other")),
         TimeSinceFire2015= ordered(TimeSinceFire2015, levels = c("0", "1", "2-5", "6-10", "10-19", "20-31", "Other")),
         TimeSinceFire2014= ordered(TimeSinceFire2014, levels = c("0", "1", "2-5", "6-10", "10-19", "20-31", "Other")),
         TimeSinceFire2013= ordered(TimeSinceFire2013, levels = c("0", "1", "2-5", "6-10", "10-19", "20-31", "Other")),
         burned = ordered(burned, levels = c("Unburned", "One", "Two", "Three", "other"))) %>%
  filter(water==1)

)

```
 
```{r}
skSamples500 %>% filter(water !=1)
```


```{r}
write.csv(skSamples500, '../outputs/cleanedSamplePoints/skStratifiedSample500.cvs', row.names=F)
```

## Reorganize 500
```{r}
( 
  samp2018 = 
    skSamples500 %>% 
    dplyr::select(ID, ECO_NAME, alb2018, countYr:sumYr, burned, TSF2018, TimeSinceFire2018) %>%
    rename(albedo=alb2018, TSF = TSF2018, TimeSinceFire=TimeSinceFire2018) %>%
    mutate(albedoYr=2018,
           tsfYr=2018)
)
```

```{r}
( 
  samp2017 = 
    skSamples500 %>% 
    dplyr::select(ID, ECO_NAME, alb2017, countYr:sumYr, burned, TSF2017, TimeSinceFire2017) %>%
    rename(albedo=alb2017, TSF = TSF2017, TimeSinceFire=TimeSinceFire2017) %>%
    mutate(albedoYr=2017,
           tsfYr=2017)
)
```

```{r}
( 
  samp2016 = 
    skSamples500 %>% 
    dplyr::select(ID, ECO_NAME, alb2016, countYr:sumYr, burned, TSF2016, TimeSinceFire2016) %>%
    rename(albedo=alb2016, TSF = TSF2016, TimeSinceFire=TimeSinceFire2016) %>%
    mutate(albedoYr=2016,
           tsfYr=2016)
)
```

```{r}
( 
  samp2015 = 
    skSamples500 %>% 
    dplyr::select(ID, ECO_NAME, alb2015, countYr:sumYr, burned, TSF2015, TimeSinceFire2015) %>%
    rename(albedo=alb2015, TSF = TSF2015, TimeSinceFire=TimeSinceFire2015) %>%
    mutate(albedoYr=2015,
           tsfYr=2015)
)
```

```{r}
( 
  samp2014 = 
    skSamples500 %>% 
    dplyr::select(ID, ECO_NAME, alb2014, countYr:sumYr, burned, TSF2014, TimeSinceFire2014) %>%
    rename(albedo=alb2014, TSF = TSF2014, TimeSinceFire=TimeSinceFire2014) %>%
    mutate(albedoYr=2014,
           tsfYr=2014)
)
```

```{r}
( 
  samp2013 = 
    skSamples500 %>% 
    dplyr::select(ID, ECO_NAME, alb2013, countYr:sumYr, burned, TSF2013, TimeSinceFire2013) %>%
    rename(albedo=alb2013, TSF = TSF2013, TimeSinceFire=TimeSinceFire2013) %>%
    mutate(albedoYr=2013,
           tsfYr=2013)
)
```

```{r}

(
  skdata =
    bind_rows(samp2018, samp2017, samp2016, samp2015, samp2014, samp2013)
)
```


```{r}
write.csv(skdata, '../outputs/cleanedSamplePoints/skStratifiedSample50020180213.cvs', row.names=F)
```


# Organize Sample Region Points 1k v2

```{r}
skalb1k
skreburn1k
sklst1k
```

```{r}
(
skSamples1k = skalb1k %>% 
  full_join(., skreburn1k, by=c("system:index", "ECO_NAME", "lat", "lon", "rowId")) %>%
  full_join(., sklst1k, by=c("system:index", "ECO_NAME", "lat", "lon", "rowId")) %>%
  rename(ID = "system:index", countYr = count.x, maxYr=maxCA, medianYr=medianCA, minYr=minCA, sumYr=sumCA) %>%
  
  mutate_at(c('maxYr', 'medianYr', 'minYr', 'sumYr'), ~replace_na(.,0)) %>%
  dplyr::select(ID:rowId, maxYr:sumYr, ST_L8, water) %>%
  mutate(burned = ifelse(countYr==0, "Unburned",
                  ifelse(countYr==1, "One",
                  ifelse(countYr==2, "Two",
                  ifelse(countYr==3, "Three", "other"))))) %>%
  mutate(TSF2018 = 2018-maxYr,
         TSF2017 = 2017-maxYr,
         TSF2016 = 2016-maxYr,
         TSF2015 = 2015-maxYr,
         TSF2014 = 2014-maxYr,
         TSF2013 = 2013-maxYr) %>%
    mutate(TimeSinceFire2018 = ifelse(countYr==0, "0",
                               ifelse(TSF2018==1, "1",
                               ifelse(TSF2018 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2018 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2018 %in% c(10:19), "11-20", 
                               ifelse(TSF2018 %in% c(20:31), "21-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2017 = ifelse(countYr==0, "0",
                               ifelse(TSF2017==1, "1",
                               ifelse(TSF2017 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2017 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2017 %in% c(10:19), "11-20", 
                               ifelse(TSF2017 %in% c(20:31), "21-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2016 = ifelse(countYr==0, "0",
                               ifelse(TSF2016==1, "1",
                               ifelse(TSF2016 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2016 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2016 %in% c(10:19), "11-20", 
                               ifelse(TSF2016 %in% c(20:31), "21-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2015 = ifelse(countYr==0, "0",
                               ifelse(TSF2015==1, "1",
                               ifelse(TSF2015 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2015 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2015 %in% c(10:19), "11-20", 
                               ifelse(TSF2015 %in% c(20:31), "21-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2014 = ifelse(countYr==0, "0",
                               ifelse(TSF2014==1, "1",
                               ifelse(TSF2014 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2014 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2014 %in% c(10:19), "11-20", 
                               ifelse(TSF2014 %in% c(20:31), "21-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2013 = ifelse(countYr==0, "0",
                               ifelse(TSF2013==1, "1",
                               ifelse(TSF2013 %in% c(2, 3, 4, 5), "2-5",
                              ifelse(TSF2013 %in% c(6, 7, 8, 9, 10), "6-10", 
                              ifelse(TSF2013 %in% c(10:19), "11-20", 
                               ifelse(TSF2013 %in% c(20:31), "21-31", "Other")))))))  %>% 
  mutate(TimeSinceFire2018= ordered(TimeSinceFire2018, levels = c("0", "1", "2-5", "6-10", "11-20", "21-31", "Other")),
         TimeSinceFire2017= ordered(TimeSinceFire2017, levels = c("0", "1", "2-5", "6-10", "11-20", "21-31", "Other")),
         TimeSinceFire2016= ordered(TimeSinceFire2016, levels = c("0", "1", "2-5", "6-10", "11-20", "21-31", "Other")),
         TimeSinceFire2015= ordered(TimeSinceFire2015, levels = c("0", "1", "2-5", "6-10", "11-20", "21-31", "Other")),
         TimeSinceFire2014= ordered(TimeSinceFire2014, levels = c("0", "1", "2-5", "6-10", "11-20", "21-31", "Other")),
         TimeSinceFire2013= ordered(TimeSinceFire2013, levels = c("0", "1", "2-5", "6-10", "11-20", "21-31", "Other")),
         burned = ordered(burned, levels = c("Unburned", "One", "Two", "Three", "other"))) %>%
  filter(water==1) %>%
  filter(lat >=55)

)

```
 
```{r}
skSamples1k %>% filter(water !=1)
skSamples1k %>% filter(lat >=55)

unique(skSamples1k$ECO_NAME)
```


```{r}
write.csv(skSamples1k, '../outputs/cleanedSamplePoints/skStratifiedSample1kv2.cvs', row.names=F)
```

## Reorganize 
```{r}
( 
  samp2018k = 
    skSamples1k %>% 
    dplyr::select(ID, ECO_NAME, alb2018, countYr:sumYr, burned, TSF2018, TimeSinceFire2018) %>%
    rename(albedo=alb2018, TSF = TSF2018, TimeSinceFire=TimeSinceFire2018) %>%
    mutate(albedoYr=2018,
           tsfYr=2018)
)
```

```{r}
( 
  samp2017k = 
    skSamples1k %>% 
    dplyr::select(ID, ECO_NAME, alb2017, countYr:sumYr, burned, TSF2017, TimeSinceFire2017) %>%
    rename(albedo=alb2017, TSF = TSF2017, TimeSinceFire=TimeSinceFire2017) %>%
    mutate(albedoYr=2017,
           tsfYr=2017)
)
```

```{r}
( 
  samp2016k = 
    skSamples1k %>% 
    dplyr::select(ID, ECO_NAME, alb2016, countYr:sumYr, burned, TSF2016, TimeSinceFire2016) %>%
    rename(albedo=alb2016, TSF = TSF2016, TimeSinceFire=TimeSinceFire2016) %>%
    mutate(albedoYr=2016,
           tsfYr=2016)
)
```

```{r}
( 
  samp2015k = 
    skSamples1k %>% 
    dplyr::select(ID, ECO_NAME, alb2015, countYr:sumYr, burned, TSF2015, TimeSinceFire2015) %>%
    rename(albedo=alb2015, TSF = TSF2015, TimeSinceFire=TimeSinceFire2015) %>%
    mutate(albedoYr=2015,
           tsfYr=2015)
)
```

```{r}
( 
  samp2014k = 
    skSamples1k %>% 
    dplyr::select(ID, ECO_NAME, alb2014, countYr:sumYr, burned, TSF2014, TimeSinceFire2014) %>%
    rename(albedo=alb2014, TSF = TSF2014, TimeSinceFire=TimeSinceFire2014) %>%
    mutate(albedoYr=2014,
           tsfYr=2014)
)
```

```{r}
( 
  samp2013k = 
    skSamples1k %>% 
    dplyr::select(ID, ECO_NAME, alb2013, countYr:sumYr, burned, TSF2013, TimeSinceFire2013) %>%
    rename(albedo=alb2013, TSF = TSF2013, TimeSinceFire=TimeSinceFire2013) %>%
    mutate(albedoYr=2013,
           tsfYr=2013)
)
```

```{r}

(
  skdata1k =
    bind_rows(samp2018k, samp2017k, samp2016k, samp2015k, samp2014k, samp2013k)
)
```
```{r}
unique(skdata1k$ECO_NAME)
skdata1k %>% filter(ECO_NAME=="Northern Canadian Shield taiga" )
```

```{r}
write.csv(skdata1k, '../outputs/cleanedSamplePoints/skStratifiedSample1k20180213v2.cvs', row.names=F)
```










