---
title: "Sampled Points"
author: "Anna Talucci"
date: "2023-07-31"
output: html_document
---

# Overview

# Packages
```{r}
library(tidyverse)
```

# Data
```{r}
temp = read.csv("../data/sampledPts/temp/meanSampledST2013.csv")
dnbr = read.csv("../data/sampledPts/dnbr/dNBR2013.csv")
alb = read.csv("../data/sampledPts/albedo/medianAlbedo2013.csv")
```

```{r}
templist <- list.files(path="../data/sampledPts/temp",pattern='csv$', full.names = TRUE)
dnbrlist <- list.files(path="../data/sampledPts/dnbr",pattern='csv$', full.names = TRUE)
alblist <- list.files(path="../data/sampledPts/albedo",pattern='csv$', full.names = TRUE)
```

```{r}
templist1 <- list.files(path="../data/sampledPts/once/temp",pattern='csv$', full.names = TRUE)
dnbrlist1 <- list.files(path="../data/sampledPts/once/dnbr",pattern='csv$', full.names = TRUE)
alblist1 <- list.files(path="../data/sampledPts/once/albedo",pattern='csv$', full.names = TRUE)
```
### Name items in a list with file name
```{r}
names(templist) <- tools::file_path_sans_ext(basename(templist))
names(dnbrlist) <- tools::file_path_sans_ext(basename(dnbrlist))
names(alblist) <- tools::file_path_sans_ext(basename(alblist))
```

```{r}
names(templist1) <- tools::file_path_sans_ext(basename(templist1))
names(dnbrlist1) <- tools::file_path_sans_ext(basename(dnbrlist1))
names(alblist1) <- tools::file_path_sans_ext(basename(alblist1))
```

# View DF
```{r}
temp
```

```{r}
dnbr
```

```{r}
alb
```

# Select columns and combine data frames

## Functions for lists

    map2(.,  ~mutate(ptID = stri_rand_strings(n(), 8)))
```{r}
orgTemp <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo, -count, -longitude, -latitude, -max, -min, -median,-sum) 
}

```

```{r}
orgDnbr <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo, -count, -longitude, -latitude, -max, -min, -median,-sum) 
}

```
   map(., ~ .x %>% mutate(rowID = row_number())) %>%
```{r}
orgAlb <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo)
}

```

## Apply function to list 
```{r}
templist
```
```{r}
allTemp = orgTemp(templist)
```

```{r}
allTemp
```
```{r}
allDnbr = orgDnbr(dnbrlist)
```

```{r}
allDnbr
```

```{r}
allAlb = orgAlb(alblist)
```

```{r}
allAlb
```
## Once
```{r}
orgTemp1 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo,  -longitude, -latitude, -min) 
}

```

```{r}
orgDnbr1 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo,  -longitude, -latitude,  -min ) 
}

```
   map(., ~ .x %>% mutate(rowID = row_number())) %>%
```{r}
orgAlb1 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo)
}

```
```{r}
allTemp1 = orgTemp1(templist1)
allDnbr1 = orgDnbr1(dnbrlist1)
allAlb1 = orgAlb1(alblist1)
```

## Individual Dataframe
```{r}
temp1 = temp%>% dplyr::select(system.index, ST_L8)
dnbr1 = dnbr%>% dplyr::select(system.index, dNBR)
```

# Combine all Dataframes
```{r}
allData = allAlb %>% 
  group_by(YrSamp) %>%
  left_join(., allTemp, by = ("system:index")) %>%
  left_join(., allDnbr, by = ("system:index")) %>%
  dplyr::select(-FXN.x, -FXN.y, -FXN, -YrSamp.x, -YrSamp.y)
  
```

## Once
```{r}
allDataOne = allAlb1 %>% 
  group_by(YrSamp) %>%
  left_join(., allTemp1, by = ("system:index")) %>%
  left_join(., allDnbr1, by = ("system:index")) %>%
  dplyr::select(-FXN.x, -FXN.y, -FXN, -YrSamp.x, -YrSamp.y)
  
```
# Add Burned unburned class


```{r}
allData
```
```{r}
min(allData$dNBR)
max(allData$dNBR)
min(allDataOne$dNBR)
max(allDataOne$dNBR)
```



```{r}
allData1 = allData %>% 
  mutate(albedo = albedo/10000) %>%
  mutate( burned = ifelse(between(dNBR,-283,-0),"unburned",
            ifelse(between(dNBR,0,100),"unburned",
            ifelse(between(dNBR,101,495),"burned", "no")))) %>%
  mutate(xBurns = sum-max-min) %>%
  mutate(reburn = ifelse(xBurns ==0, "twice",
                         ifelse(xBurns %in% c(1988, 1992, 1993, 1994, 1995, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015), "three","other" ))) %>%
  mutate(lag1 = max-min) 
  


```

```{r}
allData1
```
```{r}
min(allData1$xBurns)
max(allData1$xBurns)
```
```{r}
reburnValues = allData1 %>% filter(reburn=="other")
sort(unique(reburnValues$xBurns))
```

```{r}
sort(c(1988, 1994, 2002, 2001, 2006, 1995, 1998, 2008, 2010, 2000, 2012, 2011, 1992,  1993,  1999, 2003,  2004,  2005,  2007,  2009,  2013,  2014,  2015 ))
```

```{r}
sort(unique(allData1$lag1))
```

```{r}
allDataOnce1 = allDataOne %>% 
  mutate(albedo = albedo/10000) %>%
  mutate( burned = ifelse(between(dNBR,-283,-0),"unburned",
            ifelse(between(dNBR,0,100),"unburned",
            ifelse(between(dNBR,101,495),"burned", "no")))) 



```

# Summary stats

```{r}
allData1 %>% 
  filter(burned=='burned') %>%
  filter(reburn!="other") %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  filter(lag1>0) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            meanTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

```{r}
allData1 %>% 
  filter(burned=='unburned') %>%
  filter(reburn!="other") %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  filter(lag1>0) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            meanTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

```{r}
allDataOnce1 %>% 
  filter(burned=='burned') %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            menaTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

```{r}
allDataOnce1 %>% 
  filter(burned=='unburned') %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            menaTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

# Summarize

```{r}
allData1 %>% group_by(burned) %>%
  summarize(ct = n())
```

```{r}
allData1 %>% group_by(reburn) %>%
  summarize(ct = n())
```
# Filter out odd values
```{r}
allData2 = allData1 %>% 
  filter(burned!='no') %>%
  filter(reburn!="other") %>%
  filter(lag1>0)
```

```{r}
allData2 %>% group_by(burned) %>%
  summarize(ct = n())
```

```{r}
allData2 %>% group_by(reburn) %>%
  summarize(ct = n())
```

```{r}
allData2
```


# Explore data
```{r}


ggplot(allData2, aes(x = burned, y=albedo)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="albedo")
  

```

```{r}


ggplot(allData2, aes(x = burned, y=ST_L8)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Temperature")
  

```

```{r}


ggplot(allData2, aes(x = reburn, y=ST_L8)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Temperature")
  

```

```{r}


ggplot(allData2, aes(x = reburn, y=albedo)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Temperature")
  

```

```{r}
allData2 %>% 
  filter(reburn=="twice") %>%
  filter(burned=="burned") %>%
ggplot(., aes(x = dNBR, y=albedo)) +
  geom_point() + 
  labs(x="dNBR", y="albedo")
```

```{r}
allData2 %>% 
  filter(reburn=="twice") %>%
  filter(burned=="burned") %>%
ggplot(., aes(x = dNBR, y=ST_L8)) +
  geom_point() + 
  labs(x="dNBR", y="temp")
```

```{r}
allData2 %>% 
  filter(reburn=="twice") %>%
  filter(burned=="burned") %>%
ggplot(., aes(x = as.character(lag1), y=albedo)) +
  geom_boxplot() + 
  labs(x="lag between 1st and 2nd fire", y="albedo")
```