---
title: "Sampled Points AK Canada clean"
author: "Anna Talucci"
date: "2024-05-16"
output: html_document
---

```{r clearenvironment, include=FALSE}
rm(list=ls())
```

# Overview

Organize Albedo points from GEE.

# Packages
```{r}
library(tidyverse)
library(sf)
```

# Data

```{r}

caAlb = read_csv("../data/eeSampledData/sampleCAalbedo2018.csv") %>% dplyr::select(-.geo)
caLst =  read_csv("../data/eeSampledData/sampleCAalbedo2018.csv") %>% dplyr::select(-.geo) 
cafire = read_csv("../data/eeSampledData/sampleFireCA2018.csv") %>% dplyr::select(-.geo)
```
```{r}
akAlb = read_csv("../data/eeSampledData/sampleAKalbedo2018.csv") %>% dplyr::select(-.geo)
akfire = read_csv("../data/eeSampledData/sampleFireAK2018.csv") %>% dplyr::select(-.geo)
```

```{r}
fires = st_read("../data/spatialDataCreate/combineFires/Boreal4326Fires1986-2021.shp", "Boreal4326Fires1986-2021")
```



# Dataframes
```{r}
caAlb
cafire
```

```{r}
akAlb
akfire
```


# Join



```{r}
( caJoined = caAlb %>% 
    full_join(., cafire, by=c('system:index', 'latitude', 'longitude', 'countCA', "ECO_NAME")) %>% 
    filter(albedo !=-99) %>%
  rename(ID = "system:index", CtDnbr = dnbrCt, countYr = countCA, maxYr=maxCA, medianYr=medianCA, minYr=minCA, sumYr=sumCA) %>%
  dplyr::select(ID:longitude, dnbr1986:CtDnbr, maxYr, medianYr, minYr, sumYr) %>%
    filter(countYr <=3) %>%
  mutate(burned = ifelse(countYr==0, "Unburned",
                  ifelse(countYr==1, "One",
                  ifelse(countYr==2, "Two",
                  ifelse(countYr==3, "Three", "other"))))) %>%
  mutate(TSF = 2018-maxYr) %>%
    
    mutate(TimeSinceFire = ifelse(countYr==0, "0",
                                     ifelse(TSF==1, "1",
                                            ifelse(TSF %in% c(2, 3, 4, 5), "2-5",
                                                   ifelse(TSF %in% c(6, 7, 8, 9, 10), "6-10", 
                                                          ifelse(TSF %in% c(10:19), "10-19", 
                                                                 ifelse(TSF %in% c(20:31), "20-31", "Other")))))))  %>% 
  mutate(TimeSinceFire= ordered(TimeSinceFire, levels = c("0", "1", "2-5", "6-10", "10-19", "20-31", "Other")),
         burned = ordered(burned, levels = c("Unburned", "One", "Two", "Three", "other"))) %>%
  mutate_at(c('maxYr', 'medianYr', 'minYr', 'sumYr', 'TSF'), ~replace_na(.,0)) %>%
    pivot_longer(cols = starts_with("dnbr"), names_to = "dnbrYr", values_to = "dnbr") %>%
    mutate(dnbrYr = str_remove_all(dnbrYr, "[dnbr]")) 
  
)
```

```{r}
( akJoined = akAlb %>% 
    full_join(., akfire, by=c('system:index', 'latitude', 'longitude', 'countAK', "ECO_NAME")) %>% 
    filter(albedo !=-99) %>%
  rename(ID = "system:index", CtDnbr = dnbrCt, countYr = countAK, maxYr=maxAK, medianYr=medianAK, minYr=minAK, sumYr=sumAK) %>%
  dplyr::select(ID:longitude, dnbr1986:CtDnbr, maxYr, medianYr, minYr, sumYr) %>%
    filter(countYr <=3) %>%
  mutate(burned = ifelse(countYr==0, "Unburned",
                  ifelse(countYr==1, "One",
                  ifelse(countYr==2, "Two",
                  ifelse(countYr==3, "Three", "other"))))) %>%
  mutate(TSF = 2018-maxYr) %>%
    
    mutate(TimeSinceFire = ifelse(countYr==0, "0",
                                     ifelse(TSF==1, "1",
                                            ifelse(TSF %in% c(2, 3, 4, 5), "2-5",
                                                   ifelse(TSF %in% c(6, 7, 8, 9, 10), "6-10", 
                                                          ifelse(TSF %in% c(10:19), "10-19", 
                                                                 ifelse(TSF %in% c(20:31), "20-31", "Other")))))))  %>% 
  mutate(TimeSinceFire= ordered(TimeSinceFire, levels = c("0", "1", "2-5", "6-10", "10-19", "20-31", "Other")),
         burned = ordered(burned, levels = c("Unburned", "One", "Two", "Three", "other"))) %>%
  mutate_at(c('maxYr', 'medianYr', 'minYr', 'sumYr', 'TSF'), ~replace_na(.,0)) %>%
    pivot_longer(cols = starts_with("dnbr"), names_to = "dnbrYr", values_to = "dnbr") %>%
    mutate(dnbrYr = str_remove_all(dnbrYr, "[dnbr]")) 
  
)
```

# Filter burned unburned

```{r}

( unburnedCA = caJoined %>% 
  filter(countYr==0)
)
```

```{r}

( unburnedAK = akJoined %>% 
  filter(countYr==0)
)
```

```{r}

( burnedCA = caJoined %>% 
  filter(countYr != 0) 
)
```


```{r}

( burnedAK = akJoined %>% 
  filter(countYr != 0) 
)
```

# Add fires and calculate Lag
## Function to make points

```{r}
df_to_sf <- function(x){
  st_as_sf(x, coords = c("longitude","latitude"), crs = 4326, remove = FALSE)
}
```

## Drop Lat Lon NA and convert to shapefile 

```{r}
( burnedCApts = burnedCA %>% 
    df_to_sf() )
```

```{r}
( burnedAKpts = burnedAK %>% 
    df_to_sf() )
```

# Reproject

```{r}
fireProj = st_crs(fires)
```

```{r}
( burnedCAptsProj = st_transform(burnedCApts, fireProj) )
( burnedAKptsProj = st_transform(burnedAKpts, fireProj) )
```

# join

```{r}
sf_use_s2(FALSE) # need for error in next code chunk
```
```{r}
joinToPointsCA = st_join(burnedCAptsProj, fires) 
joinToPointsAK = st_join(burnedAKptsProj, fires) 
```

```{r}
joinToPointsCA
joinToPointsAK
```



# Oragnaize joined data

```{r}
( dfOrgCA = joinToPointsCA %>% 
  st_drop_geometry() %>%
  dplyr::select(ID, Year, FireID) %>% 
  group_by(ID) %>%
  distinct() %>%
  arrange(Year) %>%
  mutate(groupId = row_number()) %>%
  mutate(fireNum = ifelse(groupId == 1, "fire_1",
                      ifelse(groupId == 2, "fire_2",
                          ifelse(groupId == 3, "fire_3", "F")))) %>%
    dplyr::select(!groupId) %>%
    pivot_wider(names_from = fireNum, values_from = c(Year, FireID))
)
```

```{r}
( dfOrgAK = joinToPointsAK %>% 
  st_drop_geometry() %>%
  dplyr::select(ID, Year, FireID) %>% 
  group_by(ID) %>%
  distinct() %>%
  arrange(Year) %>%
  mutate(groupId = row_number()) %>%
  mutate(fireNum = ifelse(groupId == 1, "fire_1",
                      ifelse(groupId == 2, "fire_2",
                          ifelse(groupId == 3, "fire_3", "F")))) %>%
    dplyr::select(!groupId) %>%
    pivot_wider(names_from = fireNum, values_from = c(Year, FireID))
)
```


# Rejoin to dataframe

```{r}
( dfLagCA = burnedCA %>% full_join(., dfOrgCA, by="ID") %>% 
  mutate(lag1 = Year_fire_2- Year_fire_1) %>%
  mutate(lag2 = Year_fire_3- Year_fire_2) %>%
  rename(fireYr1 = Year_fire_1, fireYr2 = Year_fire_2, fireYr3 = Year_fire_3, FireId1 = FireID_fire_1, FireId2 = FireID_fire_2, FireId3 = FireID_fire_3) 
  
)
```

```{r}
( dfLagAK = burnedCA %>% full_join(., dfOrgAK, by="ID") %>% 
  mutate(lag1 = Year_fire_2- Year_fire_1) %>%
  mutate(lag2 = Year_fire_3- Year_fire_2) %>%
  rename(fireYr1 = Year_fire_1, fireYr2 = Year_fire_2, fireYr3 = Year_fire_3, FireId1 = FireID_fire_1, FireId2 = FireID_fire_2, FireId3 = FireID_fire_3) 
  
)
```

## Split one, two, three times burned

```{r}
oneCA = dfLagCA %>% filter(burned == "One") %>% mutate(reburnCl = "once")  %>%
    distinct(ID, .keep_all = TRUE)
twoCA = dfLagCA %>% filter(burned == "Two")
threeCA = dfLagCA %>% filter(burned == "Three")
```

```{r}
oneAK = dfLagAK %>% filter(burned == "One") %>% mutate(reburnCl = "once")  %>%
    distinct(ID, .keep_all = TRUE)
twoAK = dfLagAK %>% filter(burned == "Two")
threeAK = dfLagAK %>% filter(burned == "Three")
```

### Check lags

```{r}
unique(sort(twoCA$lag1))
unique(sort(threeCA$lag1))
unique(sort(threeCA$lag2))
unique(dfLagCA$burned)
```

```{r}
unique(sort(twoAK$lag1))
unique(sort(threeAK$lag1))
unique(sort(threeAK$lag2))
unique(dfLagAK$burned)
```

# Add Reburn Class

```{r}
( dfLagBurnedCA = dfLagCA %>% 
    filter(burned %in% c("Two", "Three")) %>%
    #filter(dnbrYr == Year_fire_1 | dnbrYr == Year_fire_2 | dnbrYr == Year_fire_3 ) %>%
    dplyr::select(!c(Year_F, FireID_F, CtDnbr)) %>%
    filter(maxYr <=2017) %>%
    #filter(lag1 <= 30 | lag2 <= 30) %>%
  mutate(reburnCl = ifelse( (lag1 %in% 1:30) | (lag2 %in% 1:30), "short",
                      ifelse( (lag1 > 30) | (lag2 > 30), "medium", "other"))) %>%
    filter(reburnCl !="medium") %>%
    distinct(ID, .keep_all = TRUE)
    
    
    )
```
 
```{r}
( dfLagBurnedAK = dfLagAK %>% 
    filter(burned %in% c("Two", "Three")) %>%
    #filter(dnbrYr == Year_fire_1 | dnbrYr == Year_fire_2 | dnbrYr == Year_fire_3 ) %>%
    dplyr::select(!c(Year_F, FireID_F, CtDnbr)) %>%
    filter(maxYr <=2017) %>%
    #filter(lag1 <= 30 | lag2 <= 30) %>%
  mutate(reburnCl = ifelse( (lag1 %in% 1:30) | (lag2 %in% 1:30), "short",
                      ifelse( (lag1 > 30) | (lag2 > 30), "medium", "other"))) %>%
    filter(reburnCl !="medium") %>%
    distinct(ID, .keep_all = TRUE)
    
    
    )
```
                              
```{r}
n_distinct(dfLagBurnedCA$ID)
unique(dfLagBurnedCA$burned)
```

# Add Unburned back in

```{r}
names(dfLagBurnedCA)
dfLagBurnedCA
```
```{r}
names(unburnedCA)
```

```{r}
( unburnedAddColumnsCA = unburnedCA %>% mutate(fireYr1 = 300, 
                    fireYr2 = 300, 
                    fireYr3 = 300, 
                    FireId1 = "x", 
                    FireId2 = "x",
                    FireId3 = "x",
                    lag1 = 300,
                    lag2 = 300,
                    reburnCl="unburned") %>%
  dplyr::select(!c(CtDnbr)) %>%
    distinct(ID, .keep_all = TRUE)
  
)
```

```{r}
( unburnedAddColumnsAK = unburnedAK %>% mutate(fireYr1 = 300, 
                    fireYr2 = 300, 
                    fireYr3 = 300, 
                    FireId1 = "x", 
                    FireId2 = "x",
                    FireId3 = "x",
                    lag1 = 300,
                    lag2 = 300,
                    reburnCl="unburned") %>%
  dplyr::select(!c(CtDnbr)) %>%
    distinct(ID, .keep_all = TRUE)
  
)
```

# Combine burned unburned
```{r}
(dfcombineCA = bind_rows(unburnedAddColumnsCA, oneCA, dfLagBurnedCA) %>%
 mutate(bTSF = paste0(burned,"_",TimeSinceFire),
        sample = "CAN") %>%
   dplyr::select(!c(CtDnbr, Year_F, FireID_F ))) 
```


```{r}
(dfcombineAK = bind_rows(unburnedAddColumnsAK, oneAK, dfLagBurnedAK) %>%
 mutate(bTSF = paste0(burned,"_",TimeSinceFire),
        sample="AK") %>%
   dplyr::select(!c(CtDnbr, Year_F, FireID_F ))) 
```

# Combine AK and CA

```{r}
( akCancombined = bind_rows(dfcombineAK, dfcombineCA) %>%
    distinct(ID, .keep_all = TRUE)
)
```


# Write to csv
```{r}
write.csv( akCancombined, '../outputs/cleanedSamplePoints/2018AlbedoSampleWesternCanadaAlaska.cvs', row.names=F)
```






