---
title: "SK 2013-2017"
author: "Anna Talucci"
date: "2023-09-29"
output: html_document
---



# Overview

# Packages
```{r}
library(tidyverse)
```

# Data

dnbrlist <- list.files(path="../data/sampledPts/dnbr",pattern='csv$', full.names = TRUE)
```{r}
dnbrlist1 <- list.files(path="../data/sampledPts/once/dnbr",pattern='csv$', full.names = TRUE)
dnbrlist2 <- list.files(path="../data/sampledPts/dnbr",pattern='csv$', full.names = TRUE)
```

```{r}

albspringlist1 <- list.files(path="../data/skAlbedo/onceBurned/spring",pattern='csv$', full.names = TRUE)
albspringlist2 <- list.files(path="../data/skAlbedo/twiceBurned/spring",pattern='csv$', full.names = TRUE)
```

```{r}
tempspringlist2 <- list.files(path="../data/skLST/twiceBurned/spring", pattern='csv$', full.names = TRUE)
tempsummerlist2 <- list.files(path="../data/skLST/twiceBurned/summer", pattern='csv$', full.names = TRUE)
```

```{r}
lst2Sp2013 = read_csv("../data/skLST/LSTReburnSpringMean2013.csv") %>% mutate(YrSamp = 2013) %>% dplyr::select(-.geo)
lst2Sp2014 = read_csv("../data/skLST/LSTReburnSpringMean2014.csv") %>% mutate(YrSamp = 2014) %>% dplyr::select(-.geo)
lst2Sp2015 = read_csv("../data/skLST/LSTReburnSpringMean2015.csv") %>% mutate(YrSamp = 2015) %>% dplyr::select(-.geo)
lst2Sp2016 = read_csv("../data/skLST/LSTReburnSpringMean2016.csv") %>% mutate(YrSamp = 2016) %>% dplyr::select(-.geo)
```
```{r}
lst1Sp2013 = read_csv("../data/skLST/LSTOnceSpringMean2013.csv") %>% mutate(YrSamp = 2013) %>% dplyr::select(-.geo)
lst1Sp2014 = read_csv("../data/skLST/LSTOnceSpringMean2014.csv") %>% mutate(YrSamp = 2014) %>% dplyr::select(-.geo)
lst1Sp2015 = read_csv("../data/skLST/LSTOnceSpringMean2015.csv") %>% mutate(YrSamp = 2015) %>% dplyr::select(-.geo)
lst1Sp2016 = read_csv("../data/skLST/LSTOnceSpringMean2016.csv") %>% mutate(YrSamp = 2016) %>% dplyr::select(-.geo)
```
### Name items in a list with file name
names(dnbrlist) <- tools::file_path_sans_ext(basename(dnbrlist))
```{r}
names(albspringlist1) <- tools::file_path_sans_ext(basename(albspringlist1))
names(albspringlist2) <- tools::file_path_sans_ext(basename(albspringlist2))
```

```{r}
names(dnbrlist1) <- tools::file_path_sans_ext(basename(dnbrlist1))
names(dnbrlist2) <- tools::file_path_sans_ext(basename(dnbrlist2))
```
# Combine Temp

```{r}
lst1 = bind_rows(lst1Sp2013, lst1Sp2014, lst1Sp2015, lst1Sp2016)
lst2 = bind_rows(lst2Sp2013, lst2Sp2014, lst2Sp2015, lst2Sp2016)
```
# Select columns and combine data frames
## dnbr
```{r}
orgDnbr1 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo,  -longitude, -latitude,  -min ) 
}

dnbr1 = orgDnbr1(dnbrlist1)
```

```{r}
orgDnbr2 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo, -count, -longitude, -latitude, -max, -min, -median,-sum) 
}

dnbr2 = orgDnbr2(dnbrlist2)
```

## New data
### Temp Functions for list
```{r eval=FALSE, include=FALSE}
orgTempSpring1 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo) 
}

allTempSpring1 = orgTempSpring1(tempspringlist1)
```

```{r eval=FALSE, include=FALSE}
orgTempSpring2 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  dplyr::mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo) 
}


allTempSpring2 = orgTempSpring2(tempspringlist2)
```

## Albedo Functions
```{r}
orgAlbSpring1 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo, -FXN)
}

allAlbSpring1 = orgAlbSpring1(albspringlist1)
```

```{r}
orgAlbSpring2 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo, -FXN)
}


allAlbSpring2 = orgAlbSpring2(albspringlist2)
```



# Dataframes

## Albedo
```{r}
allAlbSpring1 
allAlbSpring2
lst1
lst2
dnbr1
dnbr2
```

```{r}
( data1 = allAlbSpring1 %>% 
  mutate(burn = "once") %>% 
  mutate(season='spring') %>% 
  mutate(albedo = albedo/10000) %>% 
  group_by(YrSamp) %>%
  left_join(., dnbr1, by = ("system:index")) %>%
  left_join(., lst1, by= ("system:index")) %>%
  dplyr::select(-FXN, -YrSamp.x, -latitude.x,-longitude.x, -min.x, -YrSamp.y, -latitude.y,-longitude.y, -min.y) %>%
  mutate( burned = ifelse(between(dNBR,-283,-0),"unburned",
            ifelse(between(dNBR,0,100),"unburned",
            ifelse(between(dNBR,101,495),"burned", "no")))) 
)
  
```

```{r}
( data2 = allAlbSpring2 %>% 
  mutate(burn = "twice") %>% 
  mutate(season='spring') %>% 
  mutate(albedo = albedo/10000) %>% 
  mutate(xBurns = sum-max-min) %>%
  mutate(reburn = ifelse(xBurns ==0, "Two",
                         ifelse(xBurns %in% c(1988, 1992, 1993, 1994, 1995, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015), "Three","other" ))) %>%
  mutate(lag1 = max-min) %>% 
  group_by(YrSamp) %>%
  left_join(., dnbr2, by = ("system:index")) %>%
  left_join(., lst2, by= ("system:index")) %>%
  dplyr::select(-FXN, -YrSamp.x, -count.x, -latitude.x, -longitude.x, -max.x, -median.x, -min.x, -sum.x, -count.y, -latitude.y, -longitude.y, -max.y, -median.y, -min.y, -sum.y) %>%
  mutate( burned = ifelse(between(dNBR,-283,-0),"unburned",
            ifelse(between(dNBR,0,100),"unburned",
            ifelse(between(dNBR,101,495),"burned", "no")))) 
  )
```

```{r}
unique(data2$lag1)

```









# Summary stats

```{r}
allData1 %>% 
  filter(burned=='burned') %>%
  filter(reburn!="other") %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  filter(lag1>0) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            meanTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

```{r}
allData1 %>% 
  filter(burned=='unburned') %>%
  filter(reburn!="other") %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  filter(lag1>0) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            meanTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

```{r}
allDataOnce1 %>% 
  filter(burned=='burned') %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            menaTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

```{r}
allDataOnce1 %>% 
  filter(burned=='unburned') %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            menaTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

# Summarize

```{r}
allData1 %>% group_by(burned) %>%
  summarize(ct = n())
```

```{r}
allData1 %>% group_by(reburn) %>%
  summarize(ct = n())
```
# Filter out odd values
```{r}
filtData2 = data2 %>% 
  filter(burned!='no') %>%
  filter(burned!='unburned') %>%
  filter(reburn!="other") %>%
  filter(lag1>0) %>%
  mutate( tsf = ifelse(between(lag1,1,10),"Short",
            ifelse(between(lag1,11,20),"Moderate",
            ifelse(between(lag1,21,30),"Long", "other")))) %>%
  filter(tsf!="other")
```

```{r}
allData2 %>% group_by(burned) %>%
  summarize(ct = n())
```

```{r}
allData2 %>% group_by(reburn) %>%
  summarize(ct = n())
```

```{r}
allData2
```


# Explore data
```{r}
filtData2
```

```{r}
(
  plot1 = ggplot(filtData2, aes(x = reburn, y=albedo)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="Number of reburns", y="Spring albedo") +
    theme_bw()
  
)
```
```{r}
ggsave("../figures/Plot1.png", plot = plot1, dpi = 600, bg='white')
```
```{r}
(
  plot2 = ggplot(filtData2, aes(x = reburn, y=ST_L8)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Mean Spring Temperature degrees C") +
    theme_bw()
)
```
```{r}
ggsave("../figures/Plot2.png", plot = plot2, dpi = 600, bg='white')
```

```{r}
( 
  plot3 = ggplot(filtData2, aes(x = tsf, y=albedo)) +
  geom_violin() + 
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="albedo") +
  theme_bw()
)
```

```{r}
ggsave("../figures/Plot3.png", plot = plot3, dpi = 600, bg='white')
```
```{r}
( 
  plot4 = ggplot(filtData2, aes(x = tsf, y=ST_L8)) +
  geom_violin() + 
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Mean Spring Temperature degrees C") +
  theme_bw()
)
```

```{r}
ggsave("../figures/Plot4.png", plot = plot4, dpi = 600, bg='white')
```







```{r}


ggplot(allData2, aes(x = burned, y=ST_L8)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Temperature")
  

```

```{r}


ggplot(allData2, aes(x = reburn, y=ST_L8)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Temperature")
  

```

```{r}


ggplot(allData2, aes(x = reburn, y=albedo)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Temperature")
  

```

```{r}
allData2 %>% 
  filter(reburn=="twice") %>%
  filter(burned=="burned") %>%
ggplot(., aes(x = dNBR, y=albedo)) +
  geom_point() + 
  labs(x="dNBR", y="albedo")
```

```{r}
allData2 %>% 
  filter(reburn=="twice") %>%
  filter(burned=="burned") %>%
ggplot(., aes(x = dNBR, y=ST_L8)) +
  geom_point() + 
  labs(x="dNBR", y="temp")
```

```{r}
allData2 %>% 
  filter(reburn=="twice") %>%
  filter(burned=="burned") %>%
ggplot(., aes(x = as.character(lag1), y=albedo)) +
  geom_boxplot() + 
  labs(x="lag between 1st and 2nd fire", y="albedo")
```