---
title: "SK 2013-2017"
author: "Anna Talucci"
date: "2023-09-29"
output: html_document
---



# Overview

# Packages
```{r}
library(tidyverse)
```

# Data

dnbrlist <- list.files(path="../data/sampledPts/dnbr",pattern='csv$', full.names = TRUE)
```{r}
dnbrlist1 <- list.files(path="../data/sampledPts/once/dnbr",pattern='csv$', full.names = TRUE)
dnbrlist2 <- list.files(path="../data/sampledPts/dnbr",pattern='csv$', full.names = TRUE)
```

```{r}
tempspringlist1 <- list.files(path="../data/skLST/onceBurned/spring",pattern='csv$', full.names = TRUE)
tempsummerlist1 <- list.files(path="../data/skLST/onceBurned/summer",pattern='csv$', full.names = TRUE)
albspringlist1 <- list.files(path="../data/skAlbedo/onceBurned/spring",pattern='csv$', full.names = TRUE)
albsummerlist1 <- list.files(path="../data/skAlbedo/onceBurned/summer",pattern='csv$', full.names = TRUE)
```

```{r}
tempspringlist2 <- list.files(path="../data/skLST/twiceBurned/spring", pattern='csv$', full.names = TRUE)
tempsummerlist2 <- list.files(path="../data/skLST/twiceBurned/summer", pattern='csv$', full.names = TRUE)
```

```{r}
lstSp2013 = read_csv("../data/skLST/twiceBurned/spring/LSTReburnSpringMean2013.csv")
lstSp2014 = read_csv("../data/skLST/twiceBurned/spring/LSTReburnSpringMean2014.csv")
lstSp2015 = read_csv("../data/skLST/twiceBurned/spring/LSTReburnSpringMean2015.csv")
lstSp2016 = read_csv("../data/skLST/twiceBurned/spring/LSTReburnSpringMean2016.csv")
```
```{r}
albspringlist2 <- list.files(path="../data/skAlbedo/twiceBurned/spring",pattern='csv$', full.names = TRUE)
albsummerlist2 <- list.files(path="../data/skAlbedo/twiceBurned/summer",pattern='csv$', full.names = TRUE)
```
### Name items in a list with file name
names(dnbrlist) <- tools::file_path_sans_ext(basename(dnbrlist))
```{r}
names(tempspringlist1) <- tools::file_path_sans_ext(basename(tempspringlist1))
names(tempsummerlist1) <- tools::file_path_sans_ext(basename(tempsummerlist1))
names(albspringlist1) <- tools::file_path_sans_ext(basename(albspringlist1))
names(albsummerlist1) <- tools::file_path_sans_ext(basename(albsummerlist1))
```

```{r}
names(tempspringlist2) <- tools::file_path_sans_ext(basename(tempspringlist2))
names(tempsummerlist2) <- tools::file_path_sans_ext(basename(tempsummerlist2))
names(albspringlist2) <- tools::file_path_sans_ext(basename(albspringlist2))
names(albsummerlist2) <- tools::file_path_sans_ext(basename(albsummerlist2))
```

```{r}
tempspringlist2
```

```{r}
names(dnbrlist1) <- tools::file_path_sans_ext(basename(dnbrlist1))
names(dnbrlist2) <- tools::file_path_sans_ext(basename(dnbrlist2))
```

# Select columns and combine data frames
## dnbr
```{r}
orgDnbr1 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo,  -longitude, -latitude,  -min ) 
}

dnbr1 = orgDnbr1(dnbrlist1)
```

```{r}
orgDnbr2 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo, -count, -longitude, -latitude, -max, -min, -median,-sum) 
}

dnbr2 = orgDnbr2(dnbrlist2)
```

## New data

```{r}
orgTempSpring1 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo) 
}

allTempSpring1 = orgTempSpring1(tempspringlist1)
```

```{r}
orgTempSpring2 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  dplyr::mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo) 
}


allTempSpring2 = orgTempSpring2(tempspringlist2)
```

```{r}
orgAlbSpring1 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo, -FXN)
}

allAlbSpring1 = orgAlbSpring1(albspringlist1)
```

```{r}
orgAlbSpring2 <- function(list){ 
  purrr::map(list, read_csv) %>%
  map2(names(.), ~mutate(.x, FXN = .y)) %>%
  bind_rows() %>%
  mutate(YrSamp = as.numeric(str_extract(FXN, "[0-9]+"))) %>%
    dplyr::select(-.geo, -FXN)
}


allAlbSpring2 = orgAlbSpring2(albspringlist2)
```



# Dataframes

## Albedo
```{r}
allAlbSpring1 
allAlbSpring2
```

```{r}
aspo = allAlbSpring1 %>% 
  mutate(burn = "once") %>% 
  mutate(season='spring') %>% 
  mutate(albedo = albedo/10000) %>% 
  group_by(YrSamp) %>%
  left_join(., dnbr1, by = ("system:index")) %>%
  dplyr::select(-FXN, -YrSamp.x) %>%
  mutate( burned = ifelse(between(dNBR,-283,-0),"unburned",
            ifelse(between(dNBR,0,100),"unburned",
            ifelse(between(dNBR,101,495),"burned", "no"))))
```

```{r}
aspr = allAlbSpring2 %>% 
  mutate(burn = "twice") %>% 
  mutate(season='spring') %>% 
  mutate(albedo = albedo/10000) %>% 
  mutate(xBurns = sum-max-min) %>%
  mutate(reburn = ifelse(xBurns ==0, "twice",
                         ifelse(xBurns %in% c(1988, 1992, 1993, 1994, 1995, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015), "three","other" ))) %>%
  mutate(lag1 = max-min) %>% 
  group_by(YrSamp) %>%
  left_join(., dnbr2, by = ("system:index")) %>%
  dplyr::select(-FXN, -YrSamp.x) %>%
  mutate( burned = ifelse(between(dNBR,-283,-0),"unburned",
            ifelse(between(dNBR,0,100),"unburned",
            ifelse(between(dNBR,101,495),"burned", "no"))))
```

```{r}
aspo
aspr
```
```{r}
asp = rbind(aspo, aspr)
```



## dnbr
```{r}
dnbr = rbind(dnbr1, dnbr2)
```

```{r}
dnbr
```
# Combine Dataframes
```{r}
albedoDnbr = albedoSpring %>% 
  group_by(YrSamp) %>%
  left_join(., dnbr, by = ("system:index")) %>%
  dplyr::select(-FXN, -YrSamp.x)
  
```
```{r}
albedoDnbr
```
## Once

allDataOne = allAlb1 %>% 
  group_by(YrSamp) %>%
  left_join(., allTemp1, by = ("system:index")) %>%
  left_join(., allDnbr1, by = ("system:index")) %>%
  dplyr::select(-FXN.x, -FXN.y, -FXN, -YrSamp.x, -YrSamp.y)
  

# Add Burned unburned class


```{r}
allData
```
```{r}
min(allData$dNBR)
max(allData$dNBR)
min(allDataOne$dNBR)
max(allDataOne$dNBR)
```



```{r}
albedoDnbr1 = albedoDnbr %>% 
  mutate( burned = ifelse(between(dNBR,-283,-0),"unburned",
            ifelse(between(dNBR,0,100),"unburned",
            ifelse(between(dNBR,101,495),"burned", "no")))) %>%
  mutate(xBurns = sum-max-min) %>%
  mutate(reburn = ifelse(xBurns ==0, "twice",
                         ifelse(xBurns %in% c(1988, 1992, 1993, 1994, 1995, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015), "three","other" ))) %>%
  mutate(lag1 = max-min) 
  


```

```{r}
allData1
```
```{r}
min(allData1$xBurns)
max(allData1$xBurns)
```
```{r}
reburnValues = allData1 %>% filter(reburn=="other")
sort(unique(reburnValues$xBurns))
```

```{r}
sort(c(1988, 1994, 2002, 2001, 2006, 1995, 1998, 2008, 2010, 2000, 2012, 2011, 1992,  1993,  1999, 2003,  2004,  2005,  2007,  2009,  2013,  2014,  2015 ))
```

```{r}
sort(unique(allData1$lag1))
```

```{r}
allDataOnce1 = allDataOne %>% 
  mutate(albedo = albedo/10000) %>%
  mutate( burned = ifelse(between(dNBR,-283,-0),"unburned",
            ifelse(between(dNBR,0,100),"unburned",
            ifelse(between(dNBR,101,495),"burned", "no")))) 



```

# Summary stats

```{r}
allData1 %>% 
  filter(burned=='burned') %>%
  filter(reburn!="other") %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  filter(lag1>0) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            meanTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

```{r}
allData1 %>% 
  filter(burned=='unburned') %>%
  filter(reburn!="other") %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  filter(lag1>0) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            meanTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

```{r}
allDataOnce1 %>% 
  filter(burned=='burned') %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            menaTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

```{r}
allDataOnce1 %>% 
  filter(burned=='unburned') %>%
  filter(YrSamp %in% c(2014, 2015, 2016, 2017)) %>%
  group_by(YrSamp) %>%
  summarize(meanAlbedo= mean(albedo), 
            menaTemp = mean(ST_L8), 
            meandNBR = mean(dNBR))
```

# Summarize

```{r}
allData1 %>% group_by(burned) %>%
  summarize(ct = n())
```

```{r}
allData1 %>% group_by(reburn) %>%
  summarize(ct = n())
```
# Filter out odd values
```{r}
allData2 = allData1 %>% 
  filter(burned!='no') %>%
  filter(reburn!="other") %>%
  filter(lag1>0)
```

```{r}
allData2 %>% group_by(burned) %>%
  summarize(ct = n())
```

```{r}
allData2 %>% group_by(reburn) %>%
  summarize(ct = n())
```

```{r}
allData2
```


# Explore data
```{r}
aspo
aspr
```

```{r}

aspr %>% 
  filter(burned %in% c('burned', 'unburned')) %>%
ggplot(., aes(x = burned, y=albedo)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="albedo")
  

```
```{r}


ggplot(aspr, aes(x = reburn, y=albedo)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="albedo")
  

```

```{r}
ggplot(aspr, aes(x = dNBR, y=albedo)) +
  geom_point() + 
  labs(x="dnbr", y="albedo")
```

```{r}


ggplot(aspo, aes(x = burn, y=albedo)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="albedo")
  

```
```{r}


ggplot(allData2, aes(x = burned, y=ST_L8)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Temperature")
  

```

```{r}


ggplot(allData2, aes(x = reburn, y=ST_L8)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Temperature")
  

```

```{r}


ggplot(allData2, aes(x = reburn, y=albedo)) +
  geom_violin() + 
  
  geom_boxplot(width=0.1) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, color="red") +
  labs(x="", y="Temperature")
  

```

```{r}
allData2 %>% 
  filter(reburn=="twice") %>%
  filter(burned=="burned") %>%
ggplot(., aes(x = dNBR, y=albedo)) +
  geom_point() + 
  labs(x="dNBR", y="albedo")
```

```{r}
allData2 %>% 
  filter(reburn=="twice") %>%
  filter(burned=="burned") %>%
ggplot(., aes(x = dNBR, y=ST_L8)) +
  geom_point() + 
  labs(x="dNBR", y="temp")
```

```{r}
allData2 %>% 
  filter(reburn=="twice") %>%
  filter(burned=="burned") %>%
ggplot(., aes(x = as.character(lag1), y=albedo)) +
  geom_boxplot() + 
  labs(x="lag between 1st and 2nd fire", y="albedo")
```