---
title: "Test for differences"
author: "Anna Talucci"
date: "2023-12-05"
output: html_document
---

# Overview


# Packages
```{r}
library(tidyverse)
library(emmeans) 
library(tweedie)
library(statmod)
library(car)

library(MASS)
library(cplm)
library(aod)
```


# Data
```{r}
skdf1k = read_csv("../data/samplePoints/cleanedSamplePoints/skStratifiedSample1k20180213v2.cvs")
```
```{r}
skdf1k
```
# Lag

```{r}

```

# check factors

```{r}
skdf1k = skdf1k %>% mutate(TimeSinceFire= ordered(TimeSinceFire, levels = c("0", "1", "2-5", "6-10", "11-20", "21-31", "Other")),
         burned = ordered(burned, levels = c("Unburned", "One", "Two", "Three", "other")))
```

# Subset
```{r}
unique(skdf1k$TimeSinceFire)
```

```{r}
(
  skdf1krecent = skdf1k %>% filter(TimeSinceFire %in% c("0", "1", "2-5")) 
    )

(
  skdf1krecentskonly = skdf1k %>% 
    filter(TimeSinceFire %in% c("0", "1", "2-5")) %>% 
    filter(ECO_NAME %in% c("Mid-Canada Boreal Plains forests", "Midwest Canadian Shield forests")) %>%
    mutate(bTSF = paste(burned,"_",TimeSinceFire))
    )
```

```{r}
(
  skdf1kold = skdf1k %>% filter(TimeSinceFire %in% c("0", "6-10", "11-20", "21-31")) 
    )

skdf1koldSkOnly = skdf1k %>% filter(TimeSinceFire %in% c("0", "6-10", "11-20", "21-31"))%>% filter(ECO_NAME %in% c("Mid-Canada Boreal Plains forests", "Midwest Canadian Shield forests"))
```

# Model two variable

```{r}
model1 = glm(albedo ~ burned + burned:TimeSinceFire, data = skdf1krecentskonly, family = tweedie(link.power=0, var.power=2))

pro.tweediemodel1 = tweedie.profile( model1, p.vec = seq(1.2, 1.8, by = 0.1) )
pro.tweediemodel1$p.max
pro.tweediemodel1


model2 = glm(albedo ~ burned + burned:TimeSinceFire, data = skdf1krecentskonly, family = tweedie(link.power=0,var.power=pro.tweediemodel1$p.max))

summary(model2)
```

### Residuals: 

```{r}
skdf1krecentskonly$model2.res = resid(model2, type = "pearson")
skdf1krecentskonly$model2.fit = fitted(model2)

sum(residuals(model2, type="pearson")^2)/model2$df.res

qplot(fitted(model2), resid(model2)) + theme_bw()

```

### Drop-in-deviance Test: Percent uncharred (BOM)

In order to conduct a drop in deviance test for to assess the significance of the interaction term, we need to generate a model without the interaction term using the tweedie distribution.
```{r}
model3 = glm(albedo ~ burned + TimeSinceFire, data = skdf1krecentskonly, 
             family = tweedie(link.power=0, var.power=2))

pro.tweediemodel3 = tweedie.profile( model3, p.vec = seq(1.2, 1.8, by = 0.1) )
pro.tweediemodel3$p.max
pro.tweediemodel3$ci


model4 = glm(albedo ~ burned + TimeSinceFire, data = skdf1krecentskonly,
                family = tweedie(link.power=0, var.power=pro.tweediemodel3$p.max))
```

Drop in deviance test for the interaction term. For tweedie distribution we must specify the type of test as "chisq" for a chi-squared test.
```{r}

anova(model4, model2, test = "Chisq")
```

### Estimates & CI: Percent uncharred (BOM)

Back transform to the original scale (tweedie link function = log). The Tweedie package does not have a function to extract CI. We will manually calculate estimated CI based on SE. 


First, extract estimate and exponentiate back to the original scale
```{r}
( model2.est = summary(model2)$coefficients[, 1] )
```

Inverse link back to the original scale of the data with plogis() function. Plogis() function provides the probabilities for the estimate.
```{r}
( model2.estorg = exp(model2.est) )
```

Extract SE 
```{r}
( model2.se = summary(model2)$coefficients[, 2] )
```

Reverse transform SE back to original scale
```{r}
( model2.se.org = exp(model2.se))
```

Calculate CI for the model then inverse link back to the original scale
```{r}
( model2.upper = exp(model2.est + (model2.se * 1.96)) )

( model2.lower = exp(model2.est - (model2.se * 1.96)) )
```


# Model one combined variable for All Data

```{r}
( plains = skdf1krecentskonly %>% filter(shldPln=="Plains"))
( shield = skdf1krecentskonly %>% filter(shldPln=="Shield"))
```

```{r}
fit1 = glm(albedo ~ bTSF, data = skdf1krecentskonly, family = tweedie(link.power=0, var.power=2))

pro.tweediefit1 = tweedie.profile( fit1, p.vec = seq(1.2, 1.8, by = 0.1) )
pro.tweediefit1$p.max
pro.tweediefit1


fit2 = glm(albedo ~ bTSF, data = skdf1krecentskonly, family = tweedie(link.power=0,var.power=pro.tweediefit1$p.max))

summary(fit2)
```
```{r}
emm1 = emmeans(fit2, specs = ~ bTSF)
emm1
```
```{r}
dfemm1 = as.data.frame(emm1)

dfemm1.1 = dfemm1 %>% mutate(meanResponseScale=plogis(emmean),
                  SE=plogis(SE),
                  LCL=plogis(asymp.LCL),
                  UCL=plogis(asymp.UCL),
                  bTSF = factor(bTSF, levels = c("Unburned _ 0", 'One _ 1', 'One _ 2-5',  'Two _ 1', 'Two _ 2-5', 'Three _ 1','Three _ 2-5')))

dfemm1.1
```
```{r}
allDataPlot = ggplot(data = dfemm1.1) + 
  geom_point(aes(x=bTSF, y=meanResponseScale, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#1b9e77','#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c("unburned", "one year", "2-5 years")) +
  labs(x="", y="Albedo") +
  scale_x_discrete(labels= c("Unburned", "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

allDataPlot
```

```{r}
ggsave("../figures/skContrastAll.png", plot = allDataPlot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```

```{r}
One_1 = c(1, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0, 0)
Three_1 = c(0, 0, 1, 0, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 1, 0, 0, 0)
Two_1 = c(0, 0, 0, 0, 1, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1, 0)
Unburned_0 = c(0, 0, 0, 0, 0, 0, 1) 
```

```{r}
compUnburn = contrast(emm1, method = list("Unburned_0 - One_1" = Unburned_0 - One_1,
                             "Unburned_0 - One_2_5" = Unburned_0 - One_2_5,
                             "Unburned_0 - Two_1" = Unburned_0 - Two_1,
                             "Unburned_0 - Two_2_5" = Unburned_0 - Two_2_5,
                             "Unburned_0 - Three_1" = Unburned_0 - Three_1,
                             "Unburned_0 - Three_2_5" = Unburned_0 - Three_2_5),
         adjust = "mvt") %>%
     confint()

compUnburn
```

```{r}
contrast(emm1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))
```
```{r}
compTB = contrast(emm1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()

compTB
```

### Residuals: 

```{r}
plains$fit2.res = resid(fit2, type = "pearson")
plains$fit2.fit = fitted(fit2)

sum(residuals(fit2, type="pearson")^2)/fit2$df.res

qplot(fitted(fit2), resid(fit2)) + theme_bw()

```

### Drop-in-deviance Test: Percent uncharred (BOM)

In order to conduct a drop in deviance test for to assess the significance of the interaction term, we need to generate a model without the interaction term using the tweedie distribution.
```{r }
fit3 = glm(albedo ~ bTSF + shldPln, data = skdf1krecentskonly, 
             family = tweedie(link.power=0, var.power=2))

pro.tweediefit3 = tweedie.profile( fit3, p.vec = seq(1.2, 1.8, by = 0.1) )
pro.tweediefit3$p.max
pro.tweediefit3$ci


fit4 = glm(albedo ~ bTSF + shldPln, data = skdf1krecentskonly,
                family = tweedie(link.power=0, var.power=pro.tweediefit3$p.max))
```

Drop in deviance test for the interaction term. For tweedie distribution we must specify the type of test as "chisq" for a chi-squared test.
```{r}

anova(fit4, fit2, test = "Chisq")
```

### Estimates & CI: Percent uncharred (BOM)

Back transform to the original scale (tweedie link function = log). The Tweedie package does not have a function to extract CI. We will manually calculate estimated CI based on SE. 


First, extract estimate and exponentiate back to the original scale
```{r}
( fit2.est = as.data.frame(summary(fit2)$coefficients[, 1] ))
```

Inverse link back to the original scale of the data with plogis() function. Plogis() function provides the probabilities for the estimate.
```{r}
( fit2.estorg = exp(fit2.est) )
```

Extract SE 
```{r}
( fit2.se = summary(fit2)$coefficients[, 2] )
```

Reverse transform SE back to original scale
```{r}
( fit2.se.org = exp(fit2.se))
```

Calculate CI for the model then inverse link back to the original scale
```{r}
( fit2.upper = exp(fit2.est + (fit2.se * 1.96)) )

( fit2.lower = exp(fit2.est - (fit2.se * 1.96)) )
```
albedo ~ bTSF, data = skdf1krecentskonly, 
```{r}

fit1_t <- aov(albedo ~ bTSF, data = skdf1krecentskonly)

#view model output
summary(fit1_t)
```

```{r}
pairwise.t.test(skdf1krecentskonly$albedo, skdf1krecentskonly$bTSF, p.adjust.method="bonferroni")
```

# Model one combined variable for Plains
```{r}
( plains = skdf1krecentskonly %>% filter(shldPln=="Plains"))
( shield = skdf1krecentskonly %>% filter(shldPln=="Shield"))
```
```{r}
plains1 = glm(albedo ~ bTSF, data = plains, family = tweedie(link.power=0, var.power=2))

pro.tweedieplains1 = tweedie.profile( plains1, p.vec = seq(1.2, 1.8, by = 0.1) )
pro.tweedieplains1$p.max
pro.tweedieplains1


plains2 = glm(albedo ~ bTSF, data = plains, family = tweedie(link.power=0,var.power=pro.tweedieplains1$p.max))

summary(plains2)
```
```{r}
plainsemm1 = emmeans(plains2, specs = ~ bTSF)
plainsemm1
```
```{r}
dfplainsemm1 = as.data.frame(plainsemm1)

dfplainsemm1.1 = dfplainsemm1 %>% mutate(meanResponseScale=plogis(emmean),
                  SE=plogis(SE),
                  LCL=plogis(asymp.LCL),
                  UCL=plogis(asymp.UCL),
                  bTSF = factor(bTSF, levels = c("Unburned _ 0", 'One _ 1', 'One _ 2-5',  'Two _ 1', 'Two _ 2-5', 'Three _ 1','Three _ 2-5')))

dfemm1.1
```
```{r}
plainsDataPlot = ggplot(data = dfplainsemm1.1) + 
  geom_point(aes(x=bTSF, y=meanResponseScale, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#1b9e77','#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c("unburned", "one year", "2-5 years")) +
  labs(x="", y="Albedo") +
  scale_x_discrete(labels= c("Unburned", "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() #+
  #theme(legend.position="none")

plainsDataPlot
```

```{r}
ggsave("../figures/skContrastPlains.png", plot = plainsDataPlot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```

```{r}
One_1 = c(1, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0, 0)
Three_1 = c(0, 0, 1, 0, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 1, 0, 0, 0)
Two_1 = c(0, 0, 0, 0, 1, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1, 0)
Unburned_0 = c(0, 0, 0, 0, 0, 0, 1) 
```

```{r}
compUnburn = contrast(plainsemm1, method = list("Unburned_0 - One_1" = Unburned_0 - One_1,
                             "Unburned_0 - One_2_5" = Unburned_0 - One_2_5,
                             "Unburned_0 - Two_1" = Unburned_0 - Two_1,
                             "Unburned_0 - Two_2_5" = Unburned_0 - Two_2_5,
                             "Unburned_0 - Three_1" = Unburned_0 - Three_1,
                             "Unburned_0 - Three_2_5" = Unburned_0 - Three_2_5),
         adjust = "mvt") %>%
     confint()

compUnburn
```

```{r}
contrast(plainsemm1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))
```
```{r}
compTB = contrast(plainsemm1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()

compTB
```

### Residuals: 

```{r}
plains$fit2.res = resid(fit2, type = "pearson")
plains$fit2.fit = fitted(fit2)

sum(residuals(fit2, type="pearson")^2)/fit2$df.res

qplot(fitted(fit2), resid(fit2)) + theme_bw()

```

### Drop-in-deviance Test: Percent uncharred (BOM)

In order to conduct a drop in deviance test for to assess the significance of the interaction term, we need to generate a model without the interaction term using the tweedie distribution.
```{r }
fit3 = glm(albedo ~ bTSF + shldPln, data = skdf1krecentskonly, 
             family = tweedie(link.power=0, var.power=2))

pro.tweediefit3 = tweedie.profile( fit3, p.vec = seq(1.2, 1.8, by = 0.1) )
pro.tweediefit3$p.max
pro.tweediefit3$ci


fit4 = glm(albedo ~ bTSF + shldPln, data = skdf1krecentskonly,
                family = tweedie(link.power=0, var.power=pro.tweediefit3$p.max))
```

Drop in deviance test for the interaction term. For tweedie distribution we must specify the type of test as "chisq" for a chi-squared test.
```{r}

anova(fit4, fit2, test = "Chisq")
```

### Estimates & CI: Percent uncharred (BOM)

Back transform to the original scale (tweedie link function = log). The Tweedie package does not have a function to extract CI. We will manually calculate estimated CI based on SE. 


First, extract estimate and exponentiate back to the original scale
```{r}
( fit2.est = as.data.frame(summary(fit2)$coefficients[, 1] ))
```

Inverse link back to the original scale of the data with plogis() function. Plogis() function provides the probabilities for the estimate.
```{r}
( fit2.estorg = exp(fit2.est) )
```

Extract SE 
```{r}
( fit2.se = summary(fit2)$coefficients[, 2] )
```

Reverse transform SE back to original scale
```{r}
( fit2.se.org = exp(fit2.se))
```

Calculate CI for the model then inverse link back to the original scale
```{r}
( fit2.upper = exp(fit2.est + (fit2.se * 1.96)) )

( fit2.lower = exp(fit2.est - (fit2.se * 1.96)) )
```
albedo ~ bTSF, data = skdf1krecentskonly, 
```{r}

fit1_t <- aov(albedo ~ bTSF, data = skdf1krecentskonly)

#view model output
summary(fit1_t)
```

```{r}
pairwise.t.test(skdf1krecentskonly$albedo, skdf1krecentskonly$bTSF, p.adjust.method="bonferroni")
```


# Model one combined variable for Shield

```{r}
( plains = skdf1krecentskonly %>% filter(shldPln=="Plains"))
( shield = skdf1krecentskonly %>% filter(shldPln=="Shield"))
```

```{r}
shield1 = glm(albedo ~ bTSF, data = shield, family = tweedie(link.power=0, var.power=2))

pro.tweedieshield1 = tweedie.profile( shield1, p.vec = seq(1.2, 1.8, by = 0.1) )
pro.tweedieshield1$p.max
pro.tweedieshield1


shield2 = glm(albedo ~ bTSF, data = shield, family = tweedie(link.power=0,var.power=pro.tweedieshield1$p.max))

summary(shield2)
```
```{r}
shieldemm1 = emmeans(shield2, specs = ~ bTSF)
shieldemm1
```
```{r}
dfshieldemm1 = as.data.frame(shieldemm1)

dfshieldemm1.1 = dfshieldemm1 %>% mutate(meanResponseScale=plogis(emmean),
                  SE=plogis(SE),
                  LCL=plogis(asymp.LCL),
                  UCL=plogis(asymp.UCL),
                  bTSF = factor(bTSF, levels = c("Unburned _ 0", 'One _ 1', 'One _ 2-5',  'Two _ 1', 'Two _ 2-5', 'Three _ 1',  'Three _ 2-5')),
                  burned= c("One", "One", "Two", "Two", "Three", "Three", "Unburned"))

dfshieldemm1.1
```
```{r}
(shieldDataPlot = ggplot(data = dfshieldemm1.1) + 
  geom_point(aes(x=bTSF, y=meanResponseScale, color = bTSF)) +
  geom_errorbar(aes(x=bTSF, ymin = LCL, ymax = UCL, color = bTSF)) +
  coord_flip() +
  scale_color_manual(name= "Time Since Fire", values=c('#1b9e77','#d95f02','#7570b3',  '#d95f02','#7570b3', '#d95f02',  '#7570b3'), labels=c("unburned", "one year", "2-5 years")) +
  labs(x="Times burned", y="Albedo") +
  scale_x_discrete(labels= c("Unburned", "One", "One", "Two", "Two", "Three", "Three")) +
    theme_bw() 
)
```

```{r}
ggsave("../figures/skContrastShield.png", plot = shieldDataPlot, width = 8, height =6, units = c("in"), dpi=600, bg = "white" )
```

```{r}
One_1 = c(1, 0, 0, 0, 0, 0, 0)
One_2_5 = c(0, 1, 0, 0, 0, 0, 0)
Three_1 = c(0, 0, 1, 0, 0, 0, 0)
Three_2_5 = c(0, 0, 0, 1, 0, 0, 0)
Two_1 = c(0, 0, 0, 0, 1, 0, 0)
Two_2_5 = c(0, 0, 0, 0, 0, 1, 0)
Unburned_0 = c(0, 0, 0, 0, 0, 0, 1) 
```

```{r}
compUnburn = contrast(shieldemm1, method = list("Unburned_0 - One_1" = Unburned_0 - One_1,
                             "Unburned_0 - One_2_5" = Unburned_0 - One_2_5,
                             "Unburned_0 - Two_1" = Unburned_0 - Two_1,
                             "Unburned_0 - Two_2_5" = Unburned_0 - Two_2_5,
                             "Unburned_0 - Three_1" = Unburned_0 - Three_1,
                             "Unburned_0 - Three_2_5" = Unburned_0 - Three_2_5),
         adjust = "mvt") %>%
     confint()

compUnburn
```

```{r}
contrast(shieldemm1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5))
```
```{r}
compTB = contrast(shieldemm1, method = list("One_1 - Two_1" = One_1 - Two_1,
                             "One_1 - Three_1" = One_1 - Three_1,
                             "One_2_5 - Two_2_5" = One_2_5 - Two_2_5,
                             "One_2_5 - Three_2_5" = One_2_5 - Three_2_5) ,
         adjust = "mvt") %>%
     confint()

compTB
```

### Residuals: 

```{r}
plains$fit2.res = resid(fit2, type = "pearson")
plains$fit2.fit = fitted(fit2)

sum(residuals(fit2, type="pearson")^2)/fit2$df.res

qplot(fitted(fit2), resid(fit2)) + theme_bw()

```

### Drop-in-deviance Test: Percent uncharred (BOM)

In order to conduct a drop in deviance test for to assess the significance of the interaction term, we need to generate a model without the interaction term using the tweedie distribution.
```{r }
fit3 = glm(albedo ~ bTSF + shldPln, data = skdf1krecentskonly, 
             family = tweedie(link.power=0, var.power=2))

pro.tweediefit3 = tweedie.profile( fit3, p.vec = seq(1.2, 1.8, by = 0.1) )
pro.tweediefit3$p.max
pro.tweediefit3$ci


fit4 = glm(albedo ~ bTSF + shldPln, data = skdf1krecentskonly,
                family = tweedie(link.power=0, var.power=pro.tweediefit3$p.max))
```

Drop in deviance test for the interaction term. For tweedie distribution we must specify the type of test as "chisq" for a chi-squared test.
```{r}

anova(fit4, fit2, test = "Chisq")
```

### Estimates & CI: Percent uncharred (BOM)

Back transform to the original scale (tweedie link function = log). The Tweedie package does not have a function to extract CI. We will manually calculate estimated CI based on SE. 


First, extract estimate and exponentiate back to the original scale
```{r}
( fit2.est = as.data.frame(summary(fit2)$coefficients[, 1] ))
```

Inverse link back to the original scale of the data with plogis() function. Plogis() function provides the probabilities for the estimate.
```{r}
( fit2.estorg = exp(fit2.est) )
```

Extract SE 
```{r}
( fit2.se = summary(fit2)$coefficients[, 2] )
```

Reverse transform SE back to original scale
```{r}
( fit2.se.org = exp(fit2.se))
```

Calculate CI for the model then inverse link back to the original scale
```{r}
( fit2.upper = exp(fit2.est + (fit2.se * 1.96)) )

( fit2.lower = exp(fit2.est - (fit2.se * 1.96)) )
```
albedo ~ bTSF, data = skdf1krecentskonly, 
```{r}

fit1_t <- aov(albedo ~ bTSF, data = skdf1krecentskonly)

#view model output
summary(fit1_t)
```

```{r}
pairwise.t.test(skdf1krecentskonly$albedo, skdf1krecentskonly$bTSF, p.adjust.method="bonferroni")
```




